---
title: "SEFL21G4d_Analysis"
author: "Zach"
output:
  html_document: 
    fig_height: 4.5
    fig_width: 4.5
    highlight: kate
    number_sections: yes
    theme: journal
    toc: yes
    toc_depth: 6
editor_options: 
  chunk_output_type: console
---


# Set Global Options
```{r global_options}
knitr::opts_chunk$set(fig.width=4.5, fig.height=4.5, echo=TRUE, warning=FALSE,message=FALSE)
```



# Load Packages 
```{r}

library(ggplot2)
library(ez)
library(knitr)
library(tidyr)
library(car)
library(dplyr)
library(purrr)
library(lme4)
library(lmerTest)
library(ggh4x)

```



# Import Data 
Load  files and factorize as necessary.\
```{r}

#load  files
data_fz = read.csv('SEFL21G4d_fzdata.csv',header=TRUE)
groups = read.csv('SEFL21G4d_groups.csv',header=TRUE)

#factorize as necessary
groups$Animal = factor(groups$Animal)
data_fz$Animal = factor(data_fz$Animal)
groups$Virus = recode_factor(groups$Virus, 'mCherry'='mCherry','stGtACR1'='GtACR')

# #add group info to data files
data_fz = inner_join(x=groups,y=data_fz,by="Animal")

```



# Drop subjects
See `groups` dataframe for exclusion rationale.
```{r}

data_fz = data_fz[data_fz$Exclude!=1,]

```


# Subset data
Separate phases (trauma, trauma recall, etc) are placed in unique dataframes for ease of analysis.
```{r}

#trauma
#freezing
trauma.freezing = data_fz[data_fz$phase=='trauma' & data_fz$subphase!="shock",]

#trauma shock reactivity
trauma.shockreact = summarise_at(
  .tbl = group_by(
    .data = data_fz[data_fz$phase=='trauma' & data_fz$subphase!="postshock",],
    Animal, subphase, Virus, Sex),
  .funs = list(motion=mean),
  .vars = c('motion')
)


#trauma recall
#freezing
traumatest = data_fz[data_fz$phase=='traumatest',]
traumatest$bin = as.numeric(traumatest$bin)
traumatest$lightpair = as.factor((traumatest$bin+1) %/% 2)
traumatest$light = as.numeric((traumatest$bin+1) %% 2)

#collapse traumatest data by light off/on periods
traumatest.collapsed = summarise_at(
  .tbl = group_by(
    .data = traumatest,
    Animal, Virus, Sex, phase, light),
  .funs = list(mean),
  .vars = c('freezing','motion'))



#nov stressor data
##prepost startle stimulus
novstressor.prepost = data_fz[data_fz$phase=='mildstressor' & data_fz$subphase=='prepost',]
novstressor.prepost$bin = factor(novstressor.prepost$bin, levels=c('bl','post'))

##startle response
novstressor.startle = data_fz[data_fz$phase=='mildstressor' & data_fz$subphase=='startle',]
novstressor.startle$bin = factor(
  novstressor.startle$bin,
  levels=c('noise_pre','noise_noise','noise_post'))



#novel stressor (sefl) test data
sefltest = data_fz[data_fz$phase=='sefltest' & data_fz$subphase=='avg',]

```


# Get summary statistics for plotting
Calculate relevant means and std. errors.
```{r}

#trauma session
#freezing
trauma.freezing.stats = summarise_at(
  .tbl = group_by(
    .data = trauma.freezing,
    bin, Virus),
  .funs = list(mean=~mean(.), sem=~sd(.)/sqrt(n())),
  .vars = c('motion','freezing')
)
trauma.freezing.stats$subphase = 'postshock'
trauma.freezing.stats$subphase[trauma.freezing.stats$bin=='baseline'] = 'baseline'

#shock reactivity
trauma.shockreact.stats = summarise_at(
  .tbl = group_by(
    .data = trauma.shockreact,
    subphase, Virus),
  .funs = list(mean=~mean(.), sem=~sd(.)/sqrt(n())),
  .vars = c('motion')
)



#trauma recall
traumatest.stats = summarise_at(
  .tbl = group_by(
    .data = traumatest.collapsed,
    Virus, light),
  .funs = list(mean=~mean(.), sem=~sd(.)/sqrt(n())),
  .vars = c('freezing','motion')
)
traumatest_bybin.stats = summarise_at(
  .tbl = group_by(
    .data = traumatest,
    Virus, bin),
  .funs = list(mean=~mean(.), sem=~sd(.)/sqrt(n())),
  .vars = c('freezing','motion')
)



#novel stressor
#novel stressor pre/post
novstressor.prepost.stats = summarise_at(
  .tbl = group_by(
    .data = novstressor.prepost,
    bin, Virus),
  .funs = list(mean=~mean(.), sem=~sd(.)/sqrt(n())),
  .vars = c('freezing','motion')
)

#novel stressor startle
novstressor.startle.stats = summarise_at(
  .tbl = group_by(
    .data = novstressor.startle,
    bin, Virus),
  .funs = list(mean=~mean(.), sem=~sd(.)/sqrt(n())),
  .vars = c('freezing','motion')
)



#sefl test
sefltest.stats = summarise_at(
  .tbl = group_by(
    .data = sefltest,
    Virus),
  .funs = list(mean=~mean(.), sem=~sd(.)/sqrt(n())),
  .vars = c('freezing','motion')
)

```


# Plots
## Global Plot Parameters
```{r}

#set figure file prefix
f_pfx = '21g4d'

p = list()

#ggsave
p$save = FALSE
p$savetype = '.svg'

##Plot dimensions
p$dim = list(
  width = 4.5, #Plot width (in)
  height = 4.5, #Plot height (in)
  panel_w = unit(1, 'in'),
  panel_h = unit(2, 'in'),
  dpi = 500 #Plot dpi
)

#Plot text
p$textsize = list(
  title = 20,
  legend = 14,
  axis_title = 16,
  axis_text = 14
)

#Plot line dimensions
p$linesize = list(
  data_grp = 1,
  data_ind = .6,
  axis = .5,
  bar = .7,
  error =.75
)
p$linealpha =list(
  data_ind = .2
)

#Plot Widths
p$width = list(
  bar = .8,
  jitter = .3,
  error = .3
)

#Point/dot sizes
p$points = list(
  grp = list(
    shape = 21,
    size = 4,
    stroke = 1,
    alpha = 1
  ),
  ind = list(
    shape =  21,
    size = 2.5,
    alpha = .4
  )
)

#theme
p$theme = list(
  base = theme_bw(),
  specs = theme(
    panel.grid = element_blank(), 
    panel.border = element_rect(colour="white"),
    axis.line = element_line(
      linewidth=p$linesize$axis,
      color="black"
    ),
    plot.title = element_text(
      size=p$textsize$title, 
      hjust=.5, 
      margin=margin(b=15,t=10), 
      face="bold"
    ),
    axis.title = element_text(
      size=p$textsize$axis_title, 
      face="plain"
    ),
    axis.text=element_text(
      size=p$textsize$axis_text, 
      colour="black"
    ),
    legend.position = c(.3,.85),
    legend.title = element_blank(),
    legend.text = element_text(
      size=p$textsize$legend, 
      face="plain"
    ),
    legend.key = element_rect(colour=NA, fill=NA),
    legend.background = element_rect(fill=NA),
    plot.background = element_rect(fill = "transparent", colour = NA),
  ) 
)

#Plot colors
p$cols = list(
  color = c("white","blue4"),
  fill = c("white","blue4"),
  rec = 'slategray1'
)

#predefine partial function for saving plots
saveplot = partial(
  ggsave,
  width = p$dim$width,
  height = p$dim$height,
  units = "in", 
  dpi = p$dim$dpi,
  bg = 'transparent'
)

```


## Make Plots
### Trauma
#### Freezing
```{r, fig.width=4.5, fig.height=4.5}

#Create base layer from group level data
P.Trauma.Freezing = ggplot(
  data = trauma.freezing.stats,
  aes(x = bin,
      y = freezing_mean,
      fill = Virus,
      shape = Virus,
      group = interaction(subphase,Virus)))

#Define plot specific parameters
p.ymin = 0
p.ymax = 100
p.yint = 25
p.breaks = seq(p.ymin,p.ymax,p.yint)

#Create plot
P.Trauma.Freezing  +
  
  geom_errorbar(
    stat = "identity",
    linewidth = p$linesize$error,
    colour = "black",
    aes(ymin = freezing_mean - freezing_sem,
        ymax = freezing_mean + freezing_sem,
        width = p$width$error
        )) +
  
  geom_line(
    stat = "identity",
    linewidth = p$linesize$data_grp
  ) +
  
  geom_point(
    stat = "identity",
    size = p$points$grp$size,
    stroke = p$points$grp$stroke,
    shape = p$points$grp$shape,
    show.legend = TRUE
  ) +
  
  scale_fill_manual(values=p$cols$fill) +
  coord_cartesian(ylim=c(p.ymin,p.ymax)) +
  scale_y_continuous(breaks=p.breaks, expand = c(0, 0)) +
  scale_x_discrete(
    labels=c('P','1','2','3','4','5','6','7','8','9','10'),
    expand = c(.1,.1)
  ) +
  
  labs(
    x = "Shock",
    y = "% Freezing"
  ) +
  
  p$theme$base +
  p$theme$specs +
  
  theme(
    axis.text.x = element_text(size=p$textsize$axis_text-4),
    #axis.title.x = element_text(margin=margin(t=-5)),
    legend.title = element_blank(),
    legend.position = c(.45, 1.02),
    legend.box.margin = margin(t=-0)
  ) +
  
  force_panelsizes(
      rows = p$dim$panel_h,
      cols = unit(1.35, "in")
  ) 

if (p$save == TRUE) {
  saveplot(filename = paste(f_pfx,'_',"Trauma", p$savetype, sep = ''))
}

```


#### Shock Reactivity
```{r, fig.width=4.5, fig.height=4.5}

#Create base layer from group level data
P.Trauma.Shock = ggplot(
  data = trauma.shockreact.stats,
  aes(x = subphase,
      y = mean,
      fill = Virus,
      group = Virus)) 

#Define plot specific parameters
p.ymin = 0
p.ymax = 3000
p.yint = 1000
p.breaks = seq(p.ymin,p.ymax,p.yint)

#Create plot
P.Trauma.Shock  +
  
  geom_errorbar(
    stat = "identity",
    linewidth = p$linesize$error,
    colour = "black",
    aes(ymin = mean - sem,
        ymax = mean + sem,
        width = p$width$error)) +
  
  geom_line(
    stat="identity",
    linewidth = p$linesize$data_grp
  ) +
  
  geom_point(
    stat = "identity",
    size = p$points$grp$size,
    shape = p$points$grp$shape,
    stroke = p$points$grp$stroke,
    show.legend = FALSE
  ) +
  
  scale_fill_manual(values=p$cols$fill) +
  coord_cartesian(ylim=c(p.ymin,p.ymax)) +
  scale_y_continuous(breaks=p.breaks, expand = c(0, 0)) +
  scale_x_discrete(labels=c('bl','shock')) +
  
  labs(
    x = "",
    y = "Motion (au)"
  ) +
  
  p$theme$base +
  p$theme$specs +

  force_panelsizes(
      rows = p$dim$panel_h,
      cols = p$dim$panel_w
  ) 

# if (p$save == TRUE) {
#   saveplot(filename = paste(f_pfx,'_',"Trauma_ShockReact", p$savetype, sep = ''))
# }

```


### Trauma Recall
#### Freezing
##### Collapsed
```{r, fig.width=4.5, fig.height=4.5}


#Create base layer from group level data
P.TR = ggplot(
  data = traumatest.stats,
  aes(x = light,
      y = freezing_mean,
      fill = Virus,
      group = Virus))

#Define plot specific parameters
p.ymin = 0
p.ymax = 100
p.yint = 25
p.breaks = seq(p.ymin,p.ymax,p.yint)

#Create plot
P.TR  +
  
  geom_rect(
    aes(xmin = .5, xmax =1.5, ymin = -Inf,  ymax=Inf),
    fill = p$cols$rec
  ) +
  
  geom_errorbar(
    stat = "identity",
    linewidth = p$linesize$error,
    colour = "black",
    aes(ymin = freezing_mean - freezing_sem,
        ymax = freezing_mean + freezing_sem,
        width = p$width$error
        )) +
  
  geom_line(
    stat = "identity",
    linewidth = p$linesize$data_grp
  ) +
  
  geom_point(
    stat = "identity",
    size = p$points$grp$size,
    shape = p$points$grp$shape,
    stroke = p$points$grp$stroke,
    show.legend = FALSE
  ) +
  
  scale_fill_manual(values=p$cols$fill) +
  coord_cartesian(xlim=c(-.5,1.5), ylim=c(p.ymin,p.ymax)) +
  scale_x_continuous(breaks=c(0,1),labels=c('off','on')) +
  scale_y_continuous(breaks=p.breaks, labels = p.breaks, expand = c(0, 0)) +
  
  labs(
    x = 'Laser',
    y = "% Freezing"
  ) +
  
  p$theme$base +
  p$theme$specs +
  
  theme(
    axis.text.x = element_text(angle=-45, hjust=0, vjust=1),
  ) +
  
  force_panelsizes(
      rows = p$dim$panel_h,
      cols = p$dim$panel_w
  ) 


if (p$save == TRUE) {
  saveplot(filename = paste(f_pfx,'_',"TraumaRecall_fz", p$savetype, sep = ''))
}

```


##### By bin
```{r, fig.width=4.5, fig.height=4.5}


#Create base layer from group level data
P.TR_bybin = ggplot(
  data = traumatest_bybin.stats,
  aes(x = bin,
      y = freezing_mean,
      fill = Virus,
      group = Virus))

#Define plot specific parameters
p.ymin = 0
p.ymax = 100
p.yint = 25
p.breaks = seq(p.ymin,p.ymax,p.yint)

#Create plot
P.TR_bybin  +
  
  geom_rect(
    aes(xmin = 1.5, xmax =2.5, ymin = -Inf,  ymax=Inf),
    fill = p$cols$rec
  ) +
  
  geom_rect(
    aes(xmin = 3.5, xmax =4.5, ymin = -Inf,  ymax=Inf),
    fill = p$cols$rec
  ) +
  
  geom_errorbar(
    stat = "identity",
    linewidth = p$linesize$error,
    colour = "black",
    aes(ymin = freezing_mean - freezing_sem,
        ymax = freezing_mean + freezing_sem,
        width = p$width$error
        )) +
  
  geom_line(
    stat = "identity",
    linewidth = p$linesize$data_grp
  ) +
  
  geom_point(
    stat = "identity",
    size = p$points$grp$size,
    shape = p$points$grp$shape,
    stroke = p$points$grp$stroke,
    show.legend = FALSE
  ) +
  
  scale_fill_manual(values=p$cols$fill) +
  coord_cartesian(xlim=c(.5,4.5), ylim=c(p.ymin,p.ymax)) +
  scale_x_continuous(
    breaks=1:4,
    labels=c('Off','On','Off','On')
  ) +
  scale_y_continuous(breaks=p.breaks, labels = p.breaks, expand = c(0, 0)) +
  
  labs(
    x = 'Laser',
    y = "% Freezing"
  ) +
  
  p$theme$base +
  p$theme$specs +
  
  theme(
    axis.text.x = element_text(angle=-45, hjust=0, vjust=1),
  ) +
  
  force_panelsizes(
      rows = p$dim$panel_h,
      cols = p$dim$panel_w*1.2
  ) 


if (p$save == TRUE) {
  saveplot(filename = paste(f_pfx,'_',"TraumaRecall_bybin_fz", p$savetype, sep = ''))
}

```


### Novel Stressor
#### Pre-Post
```{r, fig.width=4.5, fig.height=4.5}


#Create base layer from group level data
P.NovSts.prepost = ggplot(
  data = novstressor.prepost.stats,
  aes(x = bin,
      y = freezing_mean,
      fill = Virus,
      group = Virus))

#Define plot specific parameters
p.ymin = 0
p.ymax = 100
p.yint = 25
p.breaks = seq(p.ymin,p.ymax,p.yint)

#Create plot
P.NovSts.prepost  +
  
  geom_errorbar(
    stat = "identity",
    linewidth = p$linesize$error,
    colour = "black",
    aes(ymin = freezing_mean - freezing_sem,
        ymax = freezing_mean + freezing_sem,
        width = p$width$error
        )) +
  
  geom_line(
    stat="identity",
    linewidth = p$linesize$data_grp
  ) +
  
  geom_point(
    stat = "identity",
    size = p$points$grp$size,
    shape = p$points$grp$shape,
    stroke = p$points$grp$stroke,
    show.legend = FALSE
  ) +
  
  scale_fill_manual(values=p$cols$fill) +
  coord_cartesian(ylim=c(p.ymin,p.ymax)) +
  scale_x_discrete(labels=c('pre','post')) +
  scale_y_continuous(breaks=p.breaks, labels = p.breaks, expand = c(0, 0)) +
  
  labs(
    x = '',
    y = "% Freezing"
  ) +
  
  p$theme$base +
  p$theme$specs +
  
  theme(
    panel.background = element_rect(fill=p$cols$rec),
    axis.text.x = element_text(angle=-45, hjust=0, vjust=1),
  ) +
  
  force_panelsizes(
      rows = p$dim$panel_h,
      cols = p$dim$panel_w
  ) 

if (p$save == TRUE) {
  saveplot(filename = paste(f_pfx,'_',"NovSts_PrePost", p$savetype, sep = ''))
}


```


#### Startle
```{r, fig.width=4.5, fig.height=4.5}


#Create base layer from group level data
P.NovSts.startle = ggplot(
  data = novstressor.startle.stats,
  aes(x = bin,
      y = motion_mean,
      fill = Virus,
      group = Virus))

#Define plot specific parameters
p.ymin = 0
p.ymax = 2500
p.yint = 500
p.breaks = seq(p.ymin,p.ymax,p.yint)

#Create plot
P.NovSts.startle  +
  
  geom_errorbar(
    stat = "identity",
    linewidth = p$linesize$error,
    colour = "black",
    aes(ymin = motion_mean - motion_sem,
        ymax = motion_mean + motion_sem,
        width = p$width$error
        )) +
  
  geom_line(
    stat="identity",
    linewidth = p$linesize$data_grp
  ) +
  
  geom_point(
    stat = "identity",
    size = p$points$grp$size,
    shape = p$points$grp$shape,
    stroke = p$points$grp$stroke,
    show.legend = FALSE
  ) +
  
  scale_fill_manual(values=p$cols$fill) +
  coord_cartesian(ylim=c(p.ymin,p.ymax)) +
  scale_x_discrete(labels=c('Pre','Noise','Post')) +
  scale_y_continuous(breaks=p.breaks, labels = p.breaks, expand = c(0, 0)) +
  
  labs(
    x = '',
    y = "Motion (au)"
  ) +
  
  p$theme$base +
  p$theme$specs +
  
  theme(
    axis.text.x= element_text(angle=-90, hjust=0),
  ) +
  
  force_panelsizes(
      rows = p$dim$panel_h,
      cols = p$dim$panel_w
  ) 

# if (p$save == TRUE) {
#   saveplot(filename = paste(f_pfx,'_',"NovSts_Startle", p$savetype, sep = ''))
# }

```

### SEFL Test
```{r, fig.width=4.5, fig.height=4.5}

#Create base layer from group level data
P.sefltest = ggplot(
  data = sefltest.stats,
  aes(x = Virus,
      y = freezing_mean,
      fill = Virus,
      group = Virus)) 

#Define plot specific parameters
p.ymin = 0
p.ymax = 40
p.yint = 10
p.breaks = seq(p.ymin,p.ymax,p.yint)

#Create plot
P.sefltest  +
  
  geom_crossbar(
    stat = "identity",
    linewidth = p$linesize$error,
    colour = "black",
    aes(ymin = freezing_mean - freezing_sem,
        ymax = freezing_mean + freezing_sem),
    width = p$width$bar,
    show.legend = FALSE) +
  
  geom_jitter(
    data = sefltest,
    aes(x = Virus,
        y = freezing),
    width = p$width$bar * p$width$jitter,
    height = 0,
    shape = p$points$ind$shape,
    size = p$points$ind$size,
    colour = 'black',
    fill = NA,
    alpha = p$points$ind$alpha,
    show.legend=FALSE
  ) +
  
  scale_fill_manual(values=p$cols$fill) +
  coord_cartesian(ylim=c(p.ymin,p.ymax)) +
  scale_y_continuous(breaks=p.breaks, labels = p.breaks, expand = c(0, 0)) +
  
  labs(
    x = "",
    y = "% Freezing"
  ) +
  
  p$theme$base +
  p$theme$specs +
  
  theme(
    axis.text.x = element_text(angle=-45, hjust=0, vjust=1)
  ) +
  
  force_panelsizes(
      rows = p$dim$panel_h,
      cols = p$dim$panel_w, #unit(1.75, 'in')
  ) 

if (p$save == TRUE) {
  saveplot(filename = paste(f_pfx,'_',"SEFLtest", p$savetype, sep = ''))
}

```



## Stats
### trauma
```{r}

trauma.anova =ezANOVA(
  data = trauma.freezing[trauma.freezing$bin!='baseline',],
  wid = Animal,
  dv = freezing,
  within = bin,
  between = Virus,
  type = 3,
  white.adjust = TRUE
)
trauma.anova

```

### trauma test
#### freezing
```{r}

# fit using anova
traumatest$light = as.character(traumatest$light)
traumatest$lightpair = as.character(traumatest$lightpair)
traumatest.bybin.anova = ezANOVA(
  data = traumatest,
  dv = freezing,
  wid = Animal,
  within = c(lightpair,light),
  white.adjust = TRUE,
  between = Virus,
  type=3
)
traumatest.bybin.anova

traumatest.collapsed$light = as.character(traumatest.collapsed$light)
traumatest.anova = ezANOVA(
  data = traumatest.collapsed,
  dv = freezing,
  wid = Animal,
  within = light,
  white.adjust = TRUE,
  between = Virus,
  type=3
)
traumatest.anova

traumatest.nolight.anova = ezANOVA(
  data = traumatest.collapsed[traumatest.collapsed$light=='0',],
  dv = freezing,
  wid = Animal,
  between = Virus,
  white.adjust = TRUE,
  type=3
)
traumatest.nolight.anova

traumatest.light.anova = ezANOVA(
  data = traumatest.collapsed[traumatest.collapsed$light=='1',],
  dv = freezing,
  wid = Animal,
  between = Virus,
  white.adjust = TRUE,
  type=3
)
traumatest.light.anova


```



### novel stressor
```{r}

novstressor.prepost.anova = ezANOVA(
  data = novstressor.prepost,
  dv = freezing,
  wid = Animal,
  within = bin,
  between = Virus,
  type = 3,
  white.adjust = TRUE
)
novstressor.prepost.anova

```


### sefltest
```{r}

sefltest.anova = ezANOVA(
  data = sefltest,
  dv = freezing,
  wid = Animal,
  between = Virus,
  type = 3,
  white.adjust = TRUE
)
sefltest.anova

```


# Export figure data for publication
```{r}

export = FALSE

#export S1 data
S1_export = trauma.freezing[,c('Animal','Sex','Virus','bin','freezing')]

#export S1 recall data
S1recall_export = traumatest[,c('Animal','Sex','Virus','bin','freezing')]

#export S2 data
S2_export =  novstressor.prepost[,c('Animal','Sex','Virus','bin','freezing')]
S2_export$bin = recode_factor(S2_export$bin, 'bl'='pre')

#export S2 recall data
S2recall_export = sefltest[,c('Animal','Sex','Virus','freezing')]

if (export){
  write.csv(S1_export, 'Fig3i.csv', row.names=FALSE)
  write.csv(S1recall_export, 'Fig3j.csv', row.names=FALSE)
  write.csv(S2_export, 'Fig3k.csv', row.names=FALSE)
  write.csv(S2recall_export, 'Fig3l.csv', row.names=FALSE)
}

```












