---
title: "SEFL21J_Analysis"
author: "Zach"
output:
  html_document: 
    fig_height: 4.5
    fig_width: 4.5
    highlight: kate
    number_sections: yes
    theme: journal
    toc: yes
    toc_depth: 6
editor_options: 
  chunk_output_type: console
---


# Set Global Options
```{r global_options}
knitr::opts_chunk$set(fig.width=4.5, fig.height=4.5, echo=TRUE, warning=FALSE,message=FALSE)
```



# Load Packages 
```{r}
library(ggplot2)
library(ez)
library(knitr)
library(tidyr)
library(lmtest)
library(car)
library(dplyr)
library(purrr)
library(ggh4x)
```



# Import Data 
Load  files and factorize as necessary.
```{r}

#load  files
data_vf = read.csv('SEFL21J_vfdata.csv',header=TRUE)
data_ld = read.csv('SEFL21J_lddata.csv',header=TRUE)
groups = read.csv('SEFL21J_groups.csv',header=TRUE)


#factorize as necessary
groups$Group = factor(groups$Group, levels = c("mCherry","HM4D: BLA", "HM4D: vHC","HM4D: LS"))
groups$Animal = factor(groups$Animal)
data_vf$Animal = factor(data_vf$Animal)
data_ld$Animal = factor(data_ld$Animal)
data_ld$testnum = factor(data_ld$testnum)


#get dataframe for drug order
drug_order = pivot_longer(
  data = groups[,c('Animal','Trec1','Trec2','LD1','LD2','LD3','LD4')],
  cols = c('Trec1','Trec2','LD1','LD2','LD3','LD4'),
  names_to = 'TestName',
  values_to = 'Drug'
)
drug_order$testnum = sapply(drug_order$TestName, function (x) {l = nchar(x); return(as.numeric(substr(x,l,l)))})
drug_order$testnum = as.factor(drug_order$testnum)
drug_order$test = sapply(drug_order$TestName,function (x) {l = nchar(x); return(substr(x,1,l-1))})
drug_order$Drug = factor(drug_order$Drug, levels = c('veh','cno'))


#add group info to data files
data_vf = inner_join(x=groups,y=data_vf,by="Animal")
data_ld = inner_join(x=groups,y=data_ld,by="Animal")
data_vf = left_join(x=data_vf, y=drug_order[drug_order$test=='Trec',], by=c('Animal'='Animal','subphase'='testnum'))
data_ld = left_join(x=data_ld, y=drug_order[drug_order$test=='LD',], by=c('Animal'='Animal','testnum'='testnum'))


```


# Drop subjects
See `groups` dataframe for exclusion rationale.
Note that there was a batch of animals that got HM4D/mCherry in the lateral septum. While very interesting (LS inhibiton increased freezing), animals with HM4Di in LS were not included in final manuscript for sake of simplicity. 
```{r}

data_vf = data_vf[data_vf$Exclude!=1,]
data_ld = data_ld[data_ld$Exclude!=1,]

data_vf = data_vf[
  !(data_vf$Virus=='HM4D' & data_vf$Source=="LS"),
]
data_ld = data_ld[
  !(data_ld$Virus=='HM4D' & data_ld$Source=="LS"),
]

```


# Subset data
Separate phases (trauma, trauma recall, etc) are placed in unique dataframes for ease of analysis.
```{r}

#trauma data
trauma.freezing = data_vf[data_vf$phase=='trauma' & data_vf$subphase!="shock",]


#trauma shock reactivity
trauma.shockreact = summarise_at(
  .tbl = group_by(
    .data = data_vf[data_vf$phase=='trauma' & data_vf$subphase!="postshock",],
    Animal,subphase, Group),
  .funs = list(motion=mean),
  .vars = c('motion')
)


#Collapse ld data
data_ld = data_ld[data_ld$bin<5,]
ld.collapsed = summarize_at(
  .tbl = group_by(data_ld, Animal, Group, Sex, Drug, testnum),
  .funs = list(mean),
  .vars = c('dark','Distance_in'))
ld.collapsed$post = as.numeric(ld.collapsed$testnum)>2

#trauma test data
traumatest = data_vf[data_vf$phase=='trec' & data_vf$component=='avg',]


#nov stressor data
##prepost startle stimulus
novstressor.prepost = data_vf[data_vf$phase=='mildstressor' & data_vf$subphase=='prepost',]
novstressor.prepost$component = factor(novstressor.prepost$component, levels=c('bl','post'))

##startle response
novstressor.startle = data_vf[data_vf$phase=='mildstressor' & data_vf$subphase=='startle',]
novstressor.startle$component = factor(
  novstressor.startle$component,
  levels=c('startle_pre','startle_startle','startle_post'))


#novel stressor (sefl) test data
sefltest = data_vf[data_vf$phase=='sefltest' & data_vf$subphase=='avg',]

```



# Get summary statistics for plotting
Calculate relevant means and std. errors.
```{r}

#trauma session freezing
trauma.freezing.stats = summarise_at(
  .tbl = group_by(
    .data = trauma.freezing,
    component, Group),
  .funs = list(mean=~mean(.), sem=~sd(.)/sqrt(n())),
  .vars = c('motion','freezing')
)
trauma.freezing.stats$subphase = 'postshock'
trauma.freezing.stats$subphase[trauma.freezing.stats$component=='baseline'] = 'baseline'

#trauma shock reactivity
trauma.shockreact.stats = summarise_at(
  .tbl = group_by(
    .data = trauma.shockreact,
    subphase, Group),
  .funs = list(mean=~mean(.), sem=~sd(.)/sqrt(n())),
  .vars = c('motion')
)


#trauma test
traumatest.stats = summarise_at(
  .tbl = group_by(
    .data = traumatest,
    Group, Drug),
  .funs = list(mean=~mean(.), sem=~sd(.)/sqrt(n())),
  .vars = c('freezing','motion')
)


#novel stressor
#novel stressor pre/post
novstressor.prepost.stats = summarise_at(
  .tbl = group_by(
    .data = novstressor.prepost,
    component, Group),
  .funs = list(mean=~mean(.), sem=~sd(.)/sqrt(n())),
  .vars = c('freezing','motion')
)

#novel stressor startle
novstressor.startle.stats = summarise_at(
  .tbl = group_by(
    .data = novstressor.startle,
    component,Group),
  .funs = list(mean=~mean(.), sem=~sd(.)/sqrt(n())),
  .vars = c('freezing','motion')
)


#sefl test
sefltest.stats = summarise_at(
  .tbl = group_by(
    .data = sefltest,
    Virus, Group),
  .funs = list(mean=~mean(.), sem=~sd(.)/sqrt(n())),
  .vars = c('freezing','motion')
)


#ld stats
ld.stats = summarise_at(
  .tbl = group_by(
    .data = ld.collapsed,
    Group, Drug, post),
  .funs = list(mean=~mean(.), sem=~sd(.)/sqrt(n())),
  .vars = c('dark','Distance_in'))


```



# Plots
## Global Plot Parameters
```{r}

#set figure file prefix
f_pfx = '21J'

#list of plotting parameters
p = list()

#ggsave
p$save = FALSE
p$savetype = '.svg'

##Plot dimensions
p$dim = list(
  width = 4.5, #Plot width (in)
  height = 4.5, #Plot height (in)
  panel_w = unit(.75, 'in'),
  panel_h = unit(2, 'in'),
  dpi = 500 #Plot dpi
)

#Plot text
p$textsize = list(
  title = 20,
  legend = 16,
  axis_title = 16,
  axis_text = 14
)

#Plot line dimensions
p$linesize = list(
  data_grp = 1,
  data_ind = .6,
  axis = .5,
  bar = .7,
  error =.75
)
p$linealpha =list(
  data_ind = .2
)

#Plot Widths
p$width = list(
  bar = .8,
  jitter = .3,
  error = .5
)

#Point/dot sizes
p$points = list(
  grp = list(
    shape = 21,
    size = 4,
    stroke = 1,
    alpha = 1
  ),
  ind = list(
    shape =  21,
    size = 2.5,
    alpha = .4
  )
)

#Plot colors
p$cols = list(
  fill = c('white','blue4','red3')
)

#theme
p$theme = list(
  base = theme_bw(),
  specs = theme(
    panel.grid = element_blank(), 
    panel.border = element_rect(colour="white"),
    axis.line = element_line(
      linewidth=p$linesize$axis,
      color="black"),
    plot.title = element_text(
      size=p$textsize$title, 
      hjust=.5, 
      margin=margin(b=15,t=10), 
      face="bold"),
    axis.title = element_text(
      size=p$textsize$axis_title, 
      face="plain"),
    axis.text=element_text(
      size=p$textsize$axis_text, 
      colour="black",
      face = 'plain'),
    axis.text.x = element_text(angle=-90, hjust=0),
    strip.background = element_blank(),
    strip.text = element_text(
      #angle = -90,
      size=p$textsize$axis_title, 
      colour="black",
      face = 'plain'),
    legend.position = c(.1,.9),
    legend.title = element_blank(),
    legend.text = element_text(size=p$textsize$legend, face="plain"),
    legend.key = element_rect(colour=NA, fill=NA),
    legend.background = element_rect(fill=NA),
    plot.background = element_rect(fill = "transparent", colour = NA),
  ) 
)

saveplot = partial(
  ggsave,
  width = p$dim$width,
  height = p$dim$height,
  units = "in",
  dpi = p$dim$dpi,
  bg = 'transparent'
)


```




## Make Plots

### Trauma
```{r, fig.width=4.5, fig.height=4.5}

#Create base layer from group level data
P.Trauma.Freezing = ggplot(
  data = trauma.freezing.stats,
  aes(x = component,
      y = freezing_mean,
      fill = Group,
      shape = Group,
      group = interaction(subphase,Group)))

#Define plot specific parameters
p.ymin = 0
p.ymax = 100
p.yint = 25
p.breaks = seq(p.ymin,p.ymax,p.yint)

#Create plot
P.Trauma.Freezing  +
  
  geom_errorbar(
    stat = "identity",
    linewidth = p$linesize$error,
    colour = "black",
    aes(ymin = freezing_mean - freezing_sem,
        ymax = freezing_mean + freezing_sem,
        width = p$width$error
        )) +
  
  geom_line(
    stat = "identity",
    linewidth = p$linesize$data_grp
  ) +
  
  geom_point(
    stat = "identity",
    size = p$points$grp$size,
    stroke = p$points$grp$stroke,
    shape = p$points$grp$shape,
    show.legend = TRUE
  ) +
  
  scale_fill_manual(values=p$cols$fill) +
  coord_cartesian(ylim=c(p.ymin,p.ymax)) +
  scale_y_continuous(breaks=p.breaks, expand = c(0, 0)) +
  scale_x_discrete(
    labels=c('P','1','2','3','4','5','6','7','8','9','10'),
    expand = c(.1,.1)
  ) +
  
  labs(
    x = "Shock",
    y = "% Freezing"
  ) +
  
  p$theme$base +
  p$theme$specs +
  
  theme(
    axis.text.x = element_text(
      size=p$textsize$axis_text-4,
      hjust = .5,
      angle = 0
    ),
    legend.title = element_blank(),
    legend.position = c(1, .2),
    legend.box.margin = margin(t=-0)
  ) +
  
  force_panelsizes(
      rows = p$dim$panel_h,
      cols = unit(1.35, "in")
  ) 

if (p$save == TRUE) {saveplot(filename = paste(f_pfx,'_',"Trauma", p$savetype, sep = ''))}

```



### Trauma Recall
```{r, fig.width=4.5, fig.height=4.5}

#Create base layer from group level data
P.Traumatest = ggplot(
  data = traumatest.stats,
  aes(x = Drug,
      y = freezing_mean,
      fill = Group,
      group = Group))

#Define plot specific parameters
p.ymin = 0
p.ymax = 100
p.yint = 25
p.breaks = seq(p.ymin,p.ymax,p.yint)

#Create plot
P.Traumatest  +
  
  geom_errorbar(
    stat = "identity",
    linewidth = p$linesize$error,
    colour = "black",
    aes(ymin = freezing_mean - freezing_sem,
        ymax = freezing_mean + freezing_sem,
        width = p$width$error
        )) +
  
  geom_line(
    stat = "identity",
    linewidth = p$linesize$data_grp
  ) +
  
  geom_point(
    stat = "identity",
    size = p$points$grp$size,
    shape = p$points$grp$shape,
    stroke = p$points$grp$stroke,
    show.legend = FALSE
  ) +

  scale_fill_manual(values=p$cols$fill) +
  coord_cartesian(ylim=c(p.ymin,p.ymax)) +
  scale_y_continuous(breaks=p.breaks, labels = p.breaks, expand = c(0, 0)) +
  scale_x_discrete(labels=c('Veh','CNO')) +
  
  labs(
    x = '',
    y = "% Freezing"
  ) +
  
  p$theme$base +
  p$theme$specs +
  theme(
    axis.text.x = element_text(vjust=.5)
  ) +
  
  force_panelsizes(
      rows = p$dim$panel_h,
      cols = p$dim$panel_w #unit(.75, 'in')
  ) 


if (p$save == TRUE) {saveplot(filename = paste(f_pfx,'_',"TraumaRecall", p$savetype, sep = ''))}

```



### Novel Stressor
#### Pre-Post
```{r, fig.width=4.5, fig.height=4.5}


#Create base layer from group level data
P.NovSts.prepost = ggplot(
  data = novstressor.prepost.stats[novstressor.prepost.stats$Group!='HM4D: LS',],
  aes(x = component,
      y = freezing_mean,
      fill = Group,
      group = Group))

#Define plot specific parameters
p.ymin = 0
p.ymax = 60
p.yint = 10
p.breaks = seq(p.ymin,p.ymax,p.yint)

#Create plot
P.NovSts.prepost  +
  
  geom_errorbar(
    stat = "identity",
    linewidth = p$linesize$error,
    colour = "black",
    aes(ymin = freezing_mean - freezing_sem,
        ymax = freezing_mean + freezing_sem,
        width = p$width$error
        )) +
  
  geom_line(
    stat="identity",
    linewidth = p$linesize$data_grp
  ) +
  
  geom_point(
    stat = "identity",
    size = p$points$grp$size,
    shape = p$points$grp$shape,
    stroke = p$points$grp$stroke,
    show.legend = FALSE
  ) +
  
  scale_fill_manual(values=p$cols$fill) +
  coord_cartesian(ylim=c(p.ymin,p.ymax)) +
  scale_x_discrete(labels=c('Pre','Post')) +
  scale_y_continuous(breaks=p.breaks, labels = p.breaks, expand = c(0, 0)) +
  
  labs(
    x = '',
    y = "% Freezing"
  ) +
  
  p$theme$base +
  p$theme$specs +
  theme(
    axis.text.x = element_text(vjust=.5),
  ) +

  force_panelsizes(
      rows = p$dim$panel_h,
      cols = p$dim$panel_w 
  ) 


if (p$save == TRUE) {saveplot(filename = paste(f_pfx,'_',"NovSts_PrePost", p$savetype, sep = ''))}


```


#### Startle
```{r, fig.width=4.5, fig.height=4.5}


#Create base layer from group level data
P.NovSts.startle = ggplot(
  data = novstressor.startle.stats,
  aes(x = component,
      y = motion_mean,
      fill = Group,
      group = Group))

#Define plot specific parameters
p.ymin = 0
p.ymax = 2500
p.yint = 500
p.breaks = seq(p.ymin,p.ymax,p.yint)

#Create plot
P.NovSts.startle  +
  
  geom_errorbar(
    stat = "identity",
    linewidth = p$linesize$error,
    colour = "black",
    aes(ymin = motion_mean - motion_sem,
        ymax = motion_mean + motion_sem,
        width = p$width$error
        )) +
  
  geom_line(
    stat="identity",
    linewidth = p$linesize$data_grp
  ) +
  
  geom_point(
    stat = "identity",
    size = p$points$grp$size,
    shape = p$points$grp$shape,
    stroke = p$points$grp$stroke,
    show.legend = FALSE
  ) +
  
  scale_fill_manual(values=p$cols$fill) +
  coord_cartesian(ylim=c(p.ymin,p.ymax)) +
  scale_x_discrete(labels=c('Pre','Noise','Post')) +
  scale_y_continuous(breaks=p.breaks, labels = p.breaks, expand = c(0, 0)) +
  
  labs(
    x = '',
    y = "Motion (au)"
  ) +
  
  p$theme$base +
  p$theme$specs +
  
  theme(
    axis.text.x= element_text(angle=-90, hjust=0),
  )

#if (p$save == TRUE) {saveplot(filename = paste(f_pfx, "NovSts_Startle.tiff", sep = '_'))}

```

### SEFL Test
```{r, fig.width=4.5, fig.height=4.5}


 p$margins[[4]] = unit(c(2, 3.6, -.4, 3.6), 'cm')

#Create base layer from group level data
P.sefltest = ggplot(
  data = sefltest.stats[sefltest.stats$Group!='HM4D: LS',],
  aes(x = Group,
      y = freezing_mean,
      fill = Group,
      group = Group)) 

#Define plot specific parameters
p.ymin = 0
p.ymax = 40
p.yint = 10
p.breaks = seq(p.ymin,p.ymax,p.yint)

#Create plot
P.sefltest  +
  
  geom_crossbar(
    stat = "identity",
    linewidth = p$linesize$error,
    colour = "black",
    aes(ymin = freezing_mean - freezing_sem,
        ymax = freezing_mean + freezing_sem),
    width = p$width$bar,
    show.legend = FALSE) +
  
  
  geom_jitter(
    data = sefltest[sefltest$Group!='HM4D: LS',],
    aes(x = Group,
        y = freezing),
    width = p$width$bar * p$width$jitter,
    height = 0,
    shape = p$points$ind$shape,
    size = p$points$ind$size,
    colour = 'black',
    fill = NA,
    alpha = p$points$ind$alpha,
    show.legend=FALSE
  ) +
  
  scale_fill_manual(values=p$cols$fill) +
  coord_cartesian(ylim=c(p.ymin,p.ymax)) +
  scale_y_continuous(breaks=p.breaks, labels = p.breaks, expand = c(0, 0)) +
  
  labs(
    x = "",
    y = "% Freezing"
  ) +
  
  p$theme$base +
  p$theme$specs +
  
  force_panelsizes(
      rows = p$dim$panel_h,
      cols = p$dim$panel_w*1.3 #unit(.75, 'in')
  ) 


if (p$save == TRUE) {saveplot(filename = paste(f_pfx,'_',"SEFLtest", p$savetype, sep = ''))}

```


### Light Dark
```{r, fig.width=4.5, fig.height=4.5}


#Create base layer from group level data
P.LD.dark = ggplot(
  data = ld.stats[ld.stats$post==TRUE & ld.stats$Group!='HM4D: LS',],
  aes(x = Drug,
      y = dark_mean,
      fill = Group,
      group = Group))

#Define plot specific parameters
p.ymin = .4
p.ymax = 1
p.yint = .2
p.breaks = seq(p.ymin,p.ymax,p.yint)

#Create plot
P.LD.dark  +

  geom_hline(yintercept=.5, linetype='dashed') +
  
  geom_errorbar(
    stat = "identity",
    linewidth = p$linesize$error,
    colour = "black",
    aes(ymin = dark_mean - dark_sem,
        ymax = dark_mean + dark_sem,
        width = p$width$error
        )) +
  
  geom_line(
    stat = "identity",
    linewidth = p$linesize$data_grp
  ) +
  
  geom_point(
    stat = "identity",
    size = p$points$grp$size,
    shape = p$points$grp$shape,
    stroke = p$points$grp$stroke,
    show.legend = FALSE
  ) +
  
  scale_fill_manual(values=p$cols$fill) +
  coord_cartesian(ylim=c(p.ymin,p.ymax)) +
  scale_x_discrete(labels=c("Veh","CNO")) +
  scale_y_continuous(breaks=p.breaks, labels = p.breaks*100, expand = c(0, 0)) +
  facet_grid(cols=vars(Group)) +
  
  labs(
    x = '',
    y = "% Time in Dark"
  ) +
  
  p$theme$base +
  p$theme$specs +
  theme(
    strip.text = element_blank(),
    axis.text.x = element_text(vjust=.5)
  ) +
  
  force_panelsizes(
      rows = p$dim$panel_h,
      cols = p$dim$panel_w*.5 #unit(.75, 'in')
  ) 


if (p$save == TRUE) {saveplot(filename = paste(f_pfx,'_',"LD_Dark", p$savetype, sep = ''))}

```



# Stats

## trauma
```{r}

trauma.fz.anova = ezANOVA(
  data = trauma.freezing[trauma.freezing$component!='baseline',],
  dv = freezing,
  wid = Animal,
  within = component,
  white.adjust = TRUE,
  between = Group,
  type = 3,
  detailed = TRUE
)
trauma.fz.anova

```

## Trauma Test Stats
```{r}

traumatest.anova = ezANOVA(
  data = traumatest,
  dv = freezing,
  wid = Animal,
  between = Group,
  within = Drug,
  white.adjust = TRUE,
  type = 3,
  detailed = TRUE
)
traumatest.anova

```


## novel stressor
```{r}

novstressor.anova = ezANOVA(
  data = novstressor.prepost,
  dv = freezing,
  wid = Animal,
  within = component,
  white.adjust = TRUE,
  between = Group,
  type = 3,
  detailed = TRUE
)
novstressor.anova


```


## SEFL Test Stats
```{r}

sefl.anova = ezANOVA(
  data = sefltest,
  dv = freezing,
  wid = Animal,
  between = Group,
  white.adjust = TRUE,
  type = 3,
  detailed = TRUE
)
sefl.anova


sefl.posthoc = list()
for (group in levels(droplevels(sefltest$Group))){
  if (group != 'mCherry') {
    sefl.posthoc[[group]] = ezANOVA(
      data = sefltest[sefltest$Group==group | sefltest$Group=='mCherry',],
      dv = freezing,
      wid = Animal,
      between = Group,
      white.adjust = TRUE,
      type = 3,
      detailed = TRUE
    )
  }
}
sefl.posthoc

```


## Light Dark Stats
```{r}

ld.dark.anova = ezANOVA(
  data = ld.collapsed[ld.collapsed$post,],
  dv = dark,
  wid = Animal,
  within = Drug,
  between = Group,
  white.adjust = FALSE,
  type = 3,
  detailed = TRUE
)
ld.dark.anova
#interaction significant when 
#HM4D: LS retained.

ld.posthoc = list()
for (group in levels(droplevels(ld.collapsed$Group))){
  ld.posthoc[[group]] = ezANOVA(
    data = ld.collapsed[ld.collapsed$post & ld.collapsed$Group==group,],
    dv = dark,
    wid = Animal,
    within = Drug,
    white.adjust = TRUE,
    type = 3,
    detailed = TRUE
  )
}
ld.posthoc

```


# Export figure data for publication
```{r}

export = FALSE

#export S1 data
S1_export = trauma.freezing[,c('Animal','Sex','Group','component','freezing')]

#export S1 recall data
S1recall_export = traumatest[,c('Animal','Sex','Group','Drug','freezing')]

#export S2 data
S2_export =  novstressor.prepost[,c('Animal','Sex','Group','component','freezing')]
S2_export$component = recode_factor(S2_export$component, 'bl'='pre')

#export S2 recall data
S2recall_export = sefltest[,c('Animal','Sex','Group','component','freezing')]

#export light-dark data
ld_export = ld.collapsed[ld.collapsed$post,c('Animal','Sex','Group','Drug','dark','Distance_in')]
ld_export$dark = ld_export$dark*100

if (export){
  write.csv(S1_export, 'ExtFig6d.csv', row.names=FALSE)
  write.csv(S1recall_export, 'ExtFig6e.csv', row.names=FALSE)
  write.csv(S2_export, 'ExtFig6f.csv', row.names=FALSE)
  write.csv(S2recall_export, 'ExtFig6g.csv', row.names=FALSE)
  write.csv(ld_export, 'ExtFig6h.csv', row.names=FALSE)
}

```






