---
title: "SEFL17_BehavAnalysis"
author: "Zach"
output:
  html_document: 
    fig_height: 4.5
    fig_width: 4.5
    highlight: kate
    number_sections: yes
    theme: journal
    toc: yes
    toc_depth: 6
editor_options: 
  chunk_output_type: console
---


# Set Global Options
```{r global_options}
knitr::opts_chunk$set(fig.width=4.5, fig.height=4.5, echo=TRUE, warning=FALSE,message=FALSE)
```



# Load Packages 
```{r}
library(ggplot2)
library(ez)
library(knitr)
library(tidyr)
library(car)
library(dplyr)
library(purrr)
library(ez)
library(lme4)
library(lmerTest)

```



# Import Data 
Load  files and factorize as necessary.
```{r}

#load  files
data_fz = read.csv('SEFL17_vfData.csv',header=TRUE)
data_ms = read.csv('SEFL17_MildStressor.csv',header=TRUE)
groups = read.csv('../SEFL17_Groups.csv',header=TRUE)

#factorize as necessary
groups$Animal = factor(groups$Animal)
data_fz$Animal = factor(data_fz$Animal)
data_ms$Animal = factor(data_ms$Animal)
groups$group = recode_factor(
  groups$group,
  'No Stim' = 'No Stim',
  'No Trauma' = 'NS',
  'Trauma' = 'S',
)

#Exclude no stim animals for behavioral analysis, since most sleeping.
groups = groups[groups$group!='No Stim',]

# #add group info to data files
data_fz = inner_join(x=groups,y=data_fz,by="Animal")
data_ms = inner_join(x=groups,y=data_ms,by="Animal")
data_ms$t = round(data_ms$t, 1)

```



# Subset data
Separate phases (trauma, trauma recall, etc) are placed in unique dataframes for ease of analysis.
```{r}

#trauma
#freezing
trauma.freezing = data_fz[data_fz$phase=='trauma' & data_fz$subphase!="shock",]

#trauma recall
#freezing
traumatest = data_fz[data_fz$phase=='traumatest' & data_fz$subphase=='avg',]

#novel stressor
#freezing
ms.fz = data_ms[data_ms$phase=='min',]
#freezing
ms.stl = data_ms[data_ms$phase=='noise',]


```



# Get summary statistics for plotting
Calculate relevant means and std. errors.
```{r}


#trauma session
#freezing
trauma.freezing.stats = summarise_at(
  .tbl = group_by(
    .data = trauma.freezing,
    component, group),
  .funs = list(mean=~mean(.), sem=~sd(.)/sqrt(n())),
  .vars = c('motion','freezing')
)
trauma.freezing.stats$subphase = 'postshock'
trauma.freezing.stats$subphase[trauma.freezing.stats$component=='baseline'] = 'baseline'


#trauma recall
traumatest.stats = summarise_at(
  .tbl = group_by(
    .data = traumatest,
    group),
  .funs = list(mean=~mean(.), sem=~sd(.)/sqrt(n())),
  .vars = c('freezing','motion')
)


#novel stressor
#freezing
prepostmin = 3 #minutes before and after startle onset
ms.fz.collapsed = summarise_at(
  .tbl = group_by(
    .data = ms.fz[ms.fz$t>=-prepostmin & ms.fz$t<prepostmin,],
    Animal, group, subphase),
  .funs = list(Freezing=~mean(.)),
  .vars = c('Freezing')
)
ms.fz.stats = summarise_at(
  .tbl = group_by(
    .data = ms.fz.collapsed,
    subphase, group),
  .funs = list(mean=~mean(.), sem=~sd(.)/sqrt(n())),
  .vars = c('Freezing')
)

#startle
prepostsec = 5 #seconds before and after startle onset
ms.stl.stats = summarise_at(
  .tbl = group_by(
    .data = ms.stl[ms.stl$t>=-prepostsec & ms.stl$t<prepostsec,],
    subphase,t, group),
  .funs = list(mean=~mean(.), sem=~sd(.)/sqrt(n())),
  .vars = c('Motion', 'Freezing')
)

```



# Plots
## Global Plot Parameters
```{r}

#set figure file prefix
f_pfx = 'SEFL17'

p = list()

#ggsave
p$save = FALSE
p$savetype = '.svg'

##Plot dimensions
p$dim = list(
  width = 4.5, #Plot width (in)
  height = 4.5, #Plot height (in)
  dpi = 500 #Plot dpi
)

#Plot text
p$textsize = list(
  title = 20,
  legend = 16,
  axis_title = 16,
  axis_text = 14
)

#Plot line dimensions
p$linesize = list(
  data_grp = 1,
  data_ind = .6,
  axis = .5,
  bar = .7,
  error =.75
)
p$linealpha =list(
  data_ind = .2
)

#Plot Widths
p$width = list(
  bar = .8,
  jitter = .3,
  error = .5
)

#Point/dot sizes
p$points = list(
  grp = list(
    shape = 21,
    size = 5,
    stroke = 1,
    alpha = 1
  ),
  ind = list(
    shape =  21,
    size = 2.5,
    alpha = .4
  )
)

#theme
p$theme = list(
  base = theme_bw(),
  specs = theme(
    panel.grid = element_blank(), 
    panel.border = element_rect(colour="white"),
    axis.line = element_line(
      linewidth=p$linesize$axis,
      color="black"),
    axis.title = element_text(size=p$textsize$axis_title, face="plain"),
    axis.text=element_text(
      size=p$textsize$axis_text, 
      colour="black"),
    axis.text.x= element_text(
      size=p$textsize$axis_title,
      face='plain',
      angle = 0,
      hjust = .5,
      vjust=.5,
      margin=margin(b=10,t=5)),
    legend.position = c(.3,.85),
    legend.title = element_blank(),
    legend.text = element_text(size=p$textsize$legend, face="plain"),
    legend.key = element_rect(colour=NA, fill=NA),
    legend.background = element_rect(fill=NA),
    plot.background = element_rect(fill = "transparent", colour = NA),
  ) 
)


#
# Plot features unique to experiment
#

#Plot colors
p$cols = list(
  color = c('gray80','red4'),
  fill = c('gray80','red4'),
  rec = 'gray80'
)

#Plot margins
p$margins = list(
  #unit(c(2, 3.75, 2, 3.75), 'cm'), # between groups
  unit(c(2, 3.75, 3.72-.5, 3.75), 'cm'), # between groups
  unit(c(2, 3.2, 2.28, 3.2), "cm") #for trauma
)
p$theme$specs = p$theme$specs + theme(plot.margin = p$margins[[1]]) #sets default margin

#predefine partial function for saving plots
saveplot = partial(
  ggsave,
  width = p$dim$width,
  height = p$dim$height,
  units = "in",
  dpi = p$dim$dpi,
  bg = 'transparent'
)

```


## Make Plots

### Trauma
```{r, fig.width=4.5, fig.height=4.5}


p$margins[[1]] = unit(c(2, 2.75, 3.5, 2.75), "cm")

#Create base layer from group level data
P.Trauma.Freezing = ggplot(
  data = trauma.freezing.stats[trauma.freezing.stats$group!='No Stim',],
  aes(x = component,
      y = freezing_mean,
      fill = group,
      group = group))

#Define plot specific parameters
p.ymin = 0
p.ymax = 100
p.yint = 25
p.breaks = seq(p.ymin,p.ymax,p.yint)

#Create plot
P.Trauma.Freezing  +
  
  # geom_ribbon(
  #   linewidth = p$linesize$error,
  #   colour = "black",
  #   alpha = .9,
  #   aes(ymin = freezing_mean - freezing_sem,
  #       ymax = freezing_mean + freezing_sem
  #       )) +
  
  geom_errorbar(
    stat = "identity",
    linewidth = p$linesize$error,
    colour = "black",
    aes(ymin = freezing_mean - freezing_sem,
        ymax = freezing_mean + freezing_sem,
        width = p$width$error
        )) +
  
  geom_line(
    stat="identity",
    linewidth = p$linesize$data_grp 
  ) +
  
  geom_point(
    stat = "identity",
    size = p$points$grp$size,
    shape = p$points$grp$shape,
    stroke = p$points$grp$stroke
  ) +
  
  scale_fill_manual(values=p$cols$fill) +
  coord_cartesian(ylim=c(p.ymin,p.ymax)) +
  scale_y_continuous(breaks=p.breaks, expand = c(0, 0)) +
  scale_x_discrete(labels=c('P','1','2','3','4','5','6','7','8','9','10'), expand=c(0,.75)) +
  
  labs(
    x = "Shock",
    y = "% Freezing"
  ) +
  
  p$theme$base +
  p$theme$specs +
  
  theme(
    axis.title.x = element_text(margin=margin(t=-3)),
    axis.text.x= element_text(
      size = p$textsize$axis_title-2,
      face='plain'),
    legend.position = c(.2,.85),
    plot.margin = p$margins[[1]]
  )

if (p$save == TRUE) {saveplot(filename = paste(f_pfx,'_',"Trauma", p$savetype, sep = ''))}

```


### Trauma Test
```{r, fig.width=4.5, fig.height=4.5}

#Create base layer from group level data
P.traumatest = ggplot(
  data = traumatest.stats[traumatest.stats$group!='No Stim',],
  aes(x = group,
      y = freezing_mean,
      fill = group,
      group = group)) 

#Define plot specific parameters
p.ymin = 0
p.ymax = 100
p.yint = 25
p.breaks = seq(p.ymin,p.ymax,p.yint)

#Create plot
P.traumatest  +
  
  geom_crossbar(
    stat = "identity",
    linewidth = p$linesize$error,
    colour = "black",
    aes(ymin = freezing_mean - freezing_sem,
        ymax = freezing_mean + freezing_sem),
    width = p$width$bar,
    show.legend = FALSE) +
  
  
  geom_jitter(
    data = traumatest[traumatest$group!='No Stim',],
    aes(x = group,
        y = freezing),
    width = p$width$bar * p$width$jitter,
    height = 0,
    shape = p$points$ind$shape,
    size = p$points$ind$size,
    colour = 'black',
    fill = NA,
    alpha = p$points$ind$alpha,
    show.legend=FALSE
  ) +
  
  scale_fill_manual(values=p$cols$fill) +
  coord_cartesian(ylim=c(p.ymin,p.ymax)) +
  scale_y_continuous(breaks=p.breaks, labels = p.breaks, expand = c(0, 0)) +
  
  labs(
    x = "",
    y = "% Freezing"
  ) +
  
  p$theme$base +
  p$theme$specs 

#if (p$save == TRUE) {saveplot(filename = paste(f_pfx,'_',"TraumaTest", p$savetype, sep = ''))}


```


### Nov Stressor Post
```{r, fig.width=4.5, fig.height=4.5}

#Create base layer from group level data
P.ms = ggplot(
  data = ms.fz.stats[ms.fz.stats$group!='No Stim' & ms.fz.stats$subphase=='post',],
  aes(x = group,
      y = mean,
      fill = group,
      group = group)) 

#Define plot specific parameters
p.ymin = 0
p.ymax = 100
p.yint = 25
p.breaks = seq(p.ymin,p.ymax,p.yint)

#Create plot
P.ms +
  
  geom_crossbar(
    stat = "identity",
    linewidth = p$linesize$error,
    colour = "black",
    aes(ymin = mean - sem,
        ymax = mean + sem),
    width = p$width$bar,
    show.legend = FALSE) +
  
  
  geom_jitter(
    data = ms.fz.collapsed[ms.fz.collapsed$group!='No Stim' & ms.fz.collapsed$subphase=='post',],
    aes(x = group,
        y = Freezing),
    width = p$width$bar * p$width$jitter,
    height = 0,
    shape = p$points$ind$shape,
    size = p$points$ind$size,
    colour = 'black',
    fill = NA,
    alpha = p$points$ind$alpha,
    show.legend=FALSE
  ) +
  
  scale_fill_manual(values=p$cols$fill) +
  coord_cartesian(ylim=c(p.ymin,p.ymax)) +
  scale_y_continuous(breaks=p.breaks, labels = p.breaks, expand = c(0, 0)) +
  
  labs(
    x = "",
    y = "% Freezing"
  ) +
  
  p$theme$base +
  p$theme$specs 

if (p$save == TRUE) {saveplot(filename = paste(f_pfx,'_',"NovStsPost", p$savetype, sep = ''))}

```


### Startle
```{r, fig.width=4.5, fig.height=4.5}

#Create base layer from group level data
P.ms.stl = ggplot(
  data = ms.stl.stats[ms.stl.stats$group!='No Stim',],
  aes(x = t,
      y = Motion_mean,
      fill = group,
      group = group)) 

#Define plot specific parameters
p.xmin = -1
p.xmax = 4
p.xint = 1
p.ymin = 0
p.ymax = 2500
p.yint = 500
p.ybreaks = seq(p.ymin,p.ymax,p.yint)
p.xbreaks = seq(p.xmin,p.xmax,p.xint)

#Create plot
P.ms.stl +
  
  geom_tile(
    aes(x=1.5),
    fill = 'gray80',
    width = 3,
    height = Inf
  ) +
  
  geom_ribbon(
    stat = "identity",
    linewidth = p$linesize$error,
    colour = "black",
    aes(ymin = Motion_mean - Motion_sem,
        ymax = Motion_mean + Motion_sem),
    alpha = .9,
    show.legend = TRUE) +
  
  scale_fill_manual(values=p$cols$fill) +
  coord_cartesian(ylim=c(p.ymin,p.ymax), xlim=c(p.xmin,p.xmax)) +
  scale_y_continuous(breaks=p.ybreaks, labels = p.ybreaks, expand = c(0, 0)) +
  scale_x_continuous(breaks=p.xbreaks, labels = p.xbreaks, expand = c(0, 0)) +
  
  labs(
    x = "Second",
    y = "Motion (au)"
  ) +
  
  p$theme$base +
  p$theme$specs +
  
  theme(
    legend.position = c(1.1,.9),
    plot.margin = unit(c(2, 3.25, 2, 3.25), 'cm') #between
  )

#if (p$save == TRUE) {saveplot(filename = paste(f_pfx,'_',"NovStsStl", p$savetype, sep = ''))}


```


# Stats

## Trauma 
```{r}

trauma.anova = ezANOVA(
  data = trauma.freezing[trauma.freezing$component!='baseline',],
  dv = freezing,
  wid = Animal,
  within = component,
  between = group
)
trauma.anova

```

## Trauma Test
```{r}

traumatest.anova = ezANOVA(
  data = traumatest,
  dv = freezing,
  wid = Animal,
  between = group
)
traumatest.anova

```

## Nov Stress
### Post freezing
```{r}

ms.fz.anova = ezANOVA(
  data = ms.fz.collapsed[ms.fz.collapsed$subphase=='post',],
  dv = Freezing,
  wid = Animal,
  between = group
)
ms.fz.anova

```

### Startle
```{r}

ms.stl.anova = ezANOVA(
  data = ms.stl[ms.stl$t>=0 & ms.stl$t<=3,],
  dv = Freezing,
  wid = Animal,
  within = t,
  between = group
)
ms.stl.anova

```



# Export figure data for publication
```{r}

export = FALSE

#export S1 data
S1_export = trauma.freezing[,c('Animal','Sex','group','component','freezing')]

#export S2 data
S2_export = ms.fz.collapsed[ms.fz.collapsed$subphase=='post',c('Animal','Sex','group','Freezing')]

if (export){
  write.csv(S1_export, 'Fig1b_Stressor1.csv', row.names=FALSE)
  write.csv(S2_export, 'Fig1c_Stressor2.csv', row.names=FALSE)
}

```









