---
title: "SEFL25b_NovStress"
author: "Zach"
output:
  html_document: 
    fig_height: 4.5
    fig_width: 4.5
    highlight: kate
    number_sections: yes
    theme: journal
    toc: yes
    toc_depth: 6
editor_options: 
  chunk_output_type: console
---


# Set Global Options
```{r global_options}
knitr::opts_chunk$set(fig.width=4.5, fig.height=4.5, echo=TRUE, warning=FALSE,message=FALSE)
```



# Load Packages 
```{r}

library(arrow)
library(ggplot2)
library(ggrepel)
library(ez)
library(knitr)
library(dplyr)
library(purrr)
library(scales)
library(tidyr)
library(ggh4x)

```



# Import Data 
Load  files and factorize as necessary.
```{r}

#load group info
groups = read.csv('../SEFL25b_groups.csv',header=TRUE)

#load average cell response
dfiles = dir('UnitClassification//')
data = list()
for (f in dfiles) {data[[f]] = read_feather(paste('UnitClassification/',f,sep=''))}
units = do.call('rbind', data)

#load average cell response
dfiles = dir('BinnedUnitResponses//')
data = list()
for (f in dfiles) {data[[f]] = read_feather(paste('BinnedUnitResponses/',f,sep=''))}
data = do.call('rbind', data)

#factorize as necessary
groups$Animal = factor(groups$Animal)
units$Animal = factor(as.numeric(units$Animal))
data$Animal = factor(as.numeric(data$Animal))
groups$Group = recode_factor(groups$Group, 'NT'='NS', 'T'='S')

#give each unit a unique id
units$uniqID = paste(units$Animal, units$unit_id, sep="_")
data$uniqID = paste(data$Animal, data$unit_id, sep="_")

#add group info to data files
units = droplevels(inner_join(x=groups, y=units, by="Animal"))
data = droplevels(inner_join(x=groups, y=data, by="Animal"))

```


# Add unit categories and prepost
```{r}

#here cells are classified only if they display an increase at p=.05, one-tailed
units = mutate(
  units, 
  class = ifelse(rand_pct>=95, 'responsive', 'nonresponsive')
)
units$class = factor(units$class, levels=c('nonresponsive','responsive'))

#add classification to data
data = inner_join(data, units[,c('uniqID','class')], by='uniqID')

```



# Exclude animals
```{r}

#had to re-run mouse 16 re-ran because cable had gotten stuck in door.
#first round, animal only received single noise
data = droplevels(data[data$Exclude!=1,])
units = droplevels(units[units$Exclude!=1,])

```


# Tabulate Unit
```{r}

#Table of all cells
UnitClassTbl = table(droplevels(units[,c('Group','class')])) 
UnitClassProp = as.data.frame(prop.table(UnitClassTbl, margin = 1))

```



# Get mean activity response, separated by group and cell classification
```{r}

sem = function(x){sd(x)/sqrt(length(x[!is.na(x)]))}

#get average of each cell across trials
data.unit.collapsed = summarise_at(
  .tbl = group_by(
    .data = data,
    Animal, uniqID, class, Group, bin),
  .funs = list(sig=mean),
  .vars = 'sig'
)

#
#units as subjects stats
#

#response by classification and time
data.unit.stats = summarise_at(
  .tbl = group_by(
    .data = data.unit.collapsed,
    bin, Group),
  .funs = list(mean=mean, sem=sem),
  .vars = c('sig'))

#response by classification and time
data.unit.byclass.stats = summarise_at(
  .tbl = group_by(
    .data = data.unit.collapsed,
    bin, class, Group),
  .funs = list(mean=mean, sem=sem),
  .vars = c('sig'))


```


# Plots
## Global Plot Parameters
```{r}

#set figure file prefix
f_pfx = '25b'

p = list()

#list of plotting parameters
p = list()

#ggsave
p$save = FALSE
p$savetype = '.svg'

##Plot dimensions
p$dim = list(
  width = 4.5, #Plot width (in)
  height = 4.5, #Plot height (in)
  panel_w = unit(1.5, 'in'),
  panel_h = unit(2, 'in'),
  dpi = 500 #Plot dpi
)

#Plot text
p$textsize = list(
  title = 20,
  legend = 16,
  axis_title = 16,
  axis_text = 14
)

#Plot line dimensions
p$linesize = list(
  data_grp = 1,
  data_ind = .6,
  axis = .5,
  bar = .7,
  error =.75
)
p$linealpha =list(
  data_ind = .2
)

#Plot Widths
p$width = list(
  bar = .8,
  jitter = .3,
  error = .5
)

#Point/dot sizes
p$points = list(
  grp = list(
    shape = 21,
    size = 4,
    stroke = 1,
    alpha = 1
  ),
  ind = list(
    shape =  21,
    size = 2.5,
    alpha = .4
  )
)


#theme
p$theme = list(
  base = theme_bw(),
  specs = theme(
    panel.grid = element_blank(), 
    panel.border = element_rect(colour="white"),
    axis.line = element_line(
      linewidth=p$linesize$axis,
      color="black"),
    plot.title = element_text(
      size=p$textsize$title, 
      hjust=.5, 
      margin=margin(b=15,t=10), 
      face="bold"),
    axis.title = element_text(size=p$textsize$axis_title, face="plain"),
    axis.text=element_text(
      size=p$textsize$axis_text, 
      colour="black",
      face = 'plain'),
    legend.position = c(.1,.9),
    legend.title = element_text(size=p$textsize$axis_title, face="plain"),
    legend.text = element_text(size=p$textsize$legend, face="plain"),
    legend.key = element_rect(colour=NA, fill=NA),
    legend.background = element_rect(fill=NA),
    plot.background = element_rect(fill = "transparent", colour = NA),
  ) 
)


#
# Plot features unique to experiment
#

#Plot colors
p$cols = list(
  color = c('royalblue3','red3'),
  fill = c('gray80','red4')
)

#predefine partial function for saving plots
saveplot = partial(
  ggsave,
  width = p$dim$width,
  height = p$dim$height,
  units = "in",
  dpi = p$dim$dpi,
  bg = 'transparent'
)

```


# Make Plots

## Unit Response

### All cells
```{r, fig.width=4.5, fig.height=4.5}

#Create base layer from group level data
P.unit.all.NTvsT = ggplot(
  data = data.unit.stats,
  aes(x = bin,
      y = mean,
      fill = Group,
      group = Group))


#Define plot specific parameters
p.xmin = -2000
p.xmax = 10000
p.xint = 5000
p.xbreaks = seq(0,p.xmax,p.xint)
p.ymin = -.1
p.ymax = .8
p.yint = .2
p.ybreaks = seq(0,p.ymax,p.yint)

#Create plot
P.unit.all.NTvsT  +
  
  geom_vline(xintercept = c(0,3000), linetype='dashed', color='gray50') +
  
  geom_ribbon(
    stat = "identity",
    linewidth = p$linesize$data_ind*.9,
    colour = 'black',
    show.legend = TRUE,
    aes(ymin = mean - sem,
        ymax = mean + sem)) +
  
  scale_fill_manual(values=p$cols$fill) +
  scale_y_continuous(breaks=p.ybreaks, labels = p.ybreaks, expand = c(0, 0)) +
  scale_x_continuous(breaks=p.xbreaks, labels=p.xbreaks/1000) +
  coord_cartesian(ylim=c(p.ymin,p.ymax), xlim=c(-1000,p.xmax)) +
  
  labs(
    x = 't (sec)',
    y = "calcium (z)"
  ) +
  
  p$theme$base +
  p$theme$specs +
  theme(
    legend.title = element_blank(),
    legend.position = c(.935, .9),
    legend.key.size = unit(.5, 'cm'),
    axis.text.x = element_text(margin = margin(t=5,b=5)),
    axis.title.x = element_text(margin = margin(t=0, b=0))
  ) +
  
  force_panelsizes(
    rows = p$dim$panel_h,
    cols = unit(1.25, "in")
  ) 

if (p$save == TRUE) {saveplot(filename = paste(f_pfx,'_',"UnitsAsSubjs_All_NSvsS", p$savetype, sep = ''))}

```



### Responsive cells
```{r, fig.width=4.5, fig.height=4.5}

#Create base layer from group level data
P.unit.responsive.NTvsT = ggplot(
  data = data.unit.byclass.stats[data.unit.byclass.stats$class=='responsive',],
  aes(x = bin,
      y = mean,
      fill = Group,
      group = Group))


#Define plot specific parameters
p.xmin = -2000
p.xmax = 10000
p.xint = 5000
p.xbreaks = seq(0,p.xmax,p.xint)
p.ymin = -.1
p.ymax = 2
p.yint = .5
p.ybreaks = seq(0,p.ymax,p.yint)

#Create plot
P.unit.responsive.NTvsT  +
  
  geom_vline(xintercept = c(0,3000), linetype='dashed', color='gray50') +
  
  geom_ribbon(
    stat = "identity",
    linewidth = p$linesize$data_ind*.9,
    colour = 'black',
    show.legend = TRUE,
    aes(ymin = mean - sem,
        ymax = mean + sem)) +
  
  scale_fill_manual(values=p$cols$fill) +
  scale_y_continuous(breaks=p.ybreaks, labels = p.ybreaks, expand = c(0, 0)) +
  scale_x_continuous(breaks=p.xbreaks, labels=p.xbreaks/1000) +
  coord_cartesian(ylim=c(p.ymin,p.ymax), xlim=c(-1000,p.xmax)) +
  
  labs(
    x = 't (sec)',
    y = "calcium (z)"
  ) +
  
  p$theme$base +
  p$theme$specs +
  theme(
    legend.title = element_blank(),
    legend.position = c(.935, .9),
    legend.key.size = unit(.5, 'cm'),
    axis.text.x = element_text(margin = margin(t=5,b=5)),
    axis.title.x = element_text(margin = margin(t=0, b=0))
  ) +

  force_panelsizes(
      rows = p$dim$panel_h,
      cols = unit(1.25, "in")
    ) 
  
  if (p$save == TRUE) {saveplot(filename = paste(f_pfx,'_',"UnitsAsSubjs_Responsive_NSvsS", p$savetype, sep = ''))}

```




## Pie Charts
### NS
```{r}

propdf = UnitClassProp[UnitClassProp$Group=='NS',]

brks = cumsum(rev(propdf$Freq)) - rev(propdf$Freq)/2
propdf$labels = as.factor(sapply(round(propdf$Freq*100, 1), function(x){paste(as.character(x), '%', sep=' ')}))

P.pie.ns = ggplot(
  propdf, 
  aes(x = 1, y = Freq, fill = class)
) 

P.pie.ns +
  
  geom_col(
    width = .1,
    colour = 'gray10',
    linewidth = 1.5,
    show.legend = TRUE,
    linejoin = 'bevel'
  ) + 
  
  scale_y_continuous(
    breaks = c()
  ) +
  
  geom_text(
    aes(label = labels, x=1.095),
    position = position_stack(vjust = 0.5),
    size = 7.5,
    fontface = 'plain',
    vjust = .5
  ) +
  
  guides(fill=guide_legend(title="Class.", nrow=2)) +
  
  scale_x_continuous(breaks=c()) +
  scale_fill_manual(values=c('gray90','blue3')) +
  coord_polar(theta = "y", direction=1, clip='off') +
  
  theme_void() +
  theme(
    axis.line = element_blank(),
    axis.title = element_blank(),
    axis.text.x = element_text(size=20, margin = margin(r=1) ),
    legend.position = c(.5, .1),
    legend.title = element_blank(),
    legend.key.size = unit(.85, 'cm'),
    legend.text = element_text(size=20, face='plain'),
    plot.background = element_rect(fill = "transparent", colour = NA),
    plot.margin = unit(c(-1, 0, 2, 0), "cm")
  ) +

  force_panelsizes(
    rows = unit(4, "in"),
    cols = unit(4, "in")
  )
  
  if (p$save == TRUE) {saveplot(filename = paste(f_pfx,'_',"Stressor2_Pie_NS", p$savetype, sep = ''))}

```

### S
```{r}

propdf = UnitClassProp[UnitClassProp$Group=='S',]

brks = cumsum(rev(propdf$Freq)) - rev(propdf$Freq)/2
propdf$labels = as.factor(sapply(round(propdf$Freq*100, 1), function(x){paste(as.character(x), '%', sep=' ')}))

P.pie.ns = ggplot(
  propdf, 
  aes(x = 1, y = Freq, fill = class)
) 

P.pie.ns +
  
  geom_col(
    width = .1,
    colour = 'gray10',
    linewidth = 1.5,
    show.legend = TRUE,
    linejoin = 'bevel'
  ) + 
  
  scale_y_continuous(
    breaks = c()
  ) +
  
  geom_text(
    aes(label = labels, x=1.095),
    position = position_stack(vjust = 0.5),
    size = 7.5,
    fontface = 'plain',
    vjust = .5
  ) +
  
  guides(fill=guide_legend(title="Class.", nrow=2)) +
  
  scale_x_continuous(breaks=c()) +
  scale_fill_manual(values=c('gray90','blue3')) +
  coord_polar(theta = "y", direction=1, clip='off') +
  
  theme_void() +
  theme(
    axis.line = element_blank(),
    axis.title = element_blank(),
    axis.text.x = element_text(size=20, margin = margin(r=1) ),
    legend.position = c(.5, .1),
    legend.title = element_blank(),
    legend.key.size = unit(.85, 'cm'),
    legend.text = element_text(size=20, face='plain'),
    plot.background = element_rect(fill = "transparent", colour = NA),
    plot.margin = unit(c(-1, 0, 2, 0), "cm")
  )
  
  if (p$save == TRUE) {saveplot(filename = paste(f_pfx,'_',"Stressor2_Pie_S", p$savetype, sep = ''))}

```


# ANOVA
## Sound response (unit level stats)
```{r}

#repeated measures for first 5 seconds for all neurons
adata = data.unit.collapsed[data.unit.collapsed$bin>0 & data.unit.collapsed$bin<=5000,]
adata$bin = factor(adata$bin)
rmanova.unit.sound = ezANOVA(
  data = adata,
  wid = uniqID,
  dv = sig,
  within = bin,
  between = Group,
  type = 3,
  white.adjust = TRUE
)
rmanova.unit.sound

#repeated measures for first 5 seconds for responsive neurons
adata = data.unit.collapsed[
  data.unit.collapsed$class=='responsive' & data.unit.collapsed$bin>0 & data.unit.collapsed$bin<=5000
  ,]
adata$bin = factor(adata$bin)
rmanova.responsiveunit.sound = ezANOVA(
  data = adata,
  wid = uniqID,
  dv = sig,
  within = bin,
  between = Group,
  type = 3,
  white.adjust = TRUE
)
rmanova.responsiveunit.sound


```



# Export figure data for publication
```{r}

export = FALSE

# export unit classification info
unit_export = units[,
  c('Group','uniqID','Animal','Sex','pre','post','dif','rand_pct','class')
]
names(unit_export)[names(unit_export)=='uniqID'] = 'unit'

# export mean unit responses
unitresponse_export = data.unit.collapsed
names(unitresponse_export)[names(unitresponse_export)=='uniqID'] = 'unit'
names(unitresponse_export)[names(unitresponse_export)=='sig'] = 'signal'
unitresponse_export$second = unitresponse_export$bin/1000
unitresponse_export = left_join(
  unitresponse_export,
  unit_export[,c('unit','Sex')],
  by = 'unit'
)
unitresponse_export = unitresponse_export[,c('unit','Animal','Sex','Group','class','second','signal')]


if (export){
  write.csv(unit_export, 'Fig2c_units.csv', row.names=FALSE)
  write.csv(unitresponse_export, 'Fig2de_unitresponse.csv', row.names=FALSE)
}

```






