---
title: 'SEFL17_HierarchicalClusterAnalysis'
author: 'Zach'
output:
  html_document: 
    fig_height: 4.5
    fig_width: 4.5
    highlight: kate
    number_sections: yes
    theme: readable
    toc: yes
    toc_depth: 6
editor_options: 
  chunk_output_type: console
---


#Set Global Options
```{r global_options}
knitr::opts_chunk$set(fig.width=4.5, fig.height=4.5, echo=TRUE, warning=FALSE,message=FALSE)
```


#Load Packages 
```{r}

library(ggplot2)
library(ggsci)
library(ggrepel)
library(factoextra)
library(ez)
library(knitr)
library(dplyr)
library(tidyr)
library(tibble)
library(stats)
library(MASS)
library(reshape2)
library(multcomp)
library(Hmisc)
library(dendextend)
library(circlize)
library(psych)
library(purrr)
library(cocor)
library(ggh4x)

#function to return top off correlation matrix, excluding off diagonal
cormat_returntop = function(cmat) {
  rvec = vector()
  cmatL = nrow(cmat)
  for (x in 1:(cmatL-1)) {rvec = c(rvec, unlist(cmat[x, (x+1):cmatL]))}
  return(rvec)
}

```


# Set plotting functions
```{r}

#set figure file prefix
f_pfx = 'SEFL17'

#ggsave
p.save = FALSE
p.savetype = '.svg'
p.savetype.corr = '.tiff'
# note that all correlation matrices tiffs. too cumbersome as svg.

#plot dims
p.dim.width = 5
p.dim.height = 5
p.dim.dpi = 500

#predefine partial function for saving plots
saveplot = partial(
  ggsave,
  width = p.dim.width,
  height = p.dim.height,
  units = "in",
  dpi = p.dim.dpi,
  bg = 'transparent'
)

```


# Initial Processing

## Import Data 
```{r}

#load counts of all regions at the lowest order of Allen Brain Atlas
counts = read.csv('countsummary_cleaned.csv')
counts = counts[counts$parcellation_group != 'cerebellum',]
counts$graph_order = factor(counts$graph_order)

#load all label info
labels = read.csv('allgraylabels.csv')
labels$parcellation_group = factor(
  labels$parcellation_group, levels = c(
    'cortical_plate', 'cortical_subplate','cerebral_nuclei','diencephalon','midbrain','hindbrain'
  )
)
labels$annotation_graph_order = factor(labels$annotation_graph_order)
labels = labels[order(labels$annotation_graph_order),]

```



## Restructure counts
```{r}

#pivot data to wide
counts_wide = pivot_wider(
  data = counts[order(counts$graph_order),],
  id_cols = c(animal,group),
  names_from = annotation_abrv,
  values_from = count
)
counts_wide = column_to_rownames(counts_wide, var = 'animal')

#z score all counts by region
counts_wide[,-1] = scale(counts_wide[,-1])

#Select each group's counts and re-center 
counts_wide.ctrl = counts_wide[counts_wide$group=='Control', -1]
counts_wide.t = counts_wide[counts_wide$group=='Trauma', -1]
counts_wide.nt = counts_wide[counts_wide$group=='No Trauma', -1]
counts_wide.ctrl[] = scale(counts_wide.ctrl)
counts_wide.t[] = scale(counts_wide.t)
counts_wide.nt[] = scale(counts_wide.nt)

```


## Get Correlation Matrix, Convert to Adjacency Matrix, and Define Distance
```{r}

rtype = 'pearson'

#control
cmat.ctrl = rcorr(as.matrix(counts_wide.ctrl), type=rtype)
cor.ctrl.orig = as.data.frame(cmat.ctrl$r)
cor.ctrl.orig.p = as.data.frame(cmat.ctrl$P)
cor.ctrl.adj = 1 - cor.ctrl.orig
cor.ctrl.dist = as.dist(cor.ctrl.adj)
distances.ctrl = as.vector(cor.ctrl.dist)

#trauma
cmat.t = rcorr(as.matrix(counts_wide.t), type=rtype)
cor.t.orig = as.data.frame(cmat.t$r)
cor.t.orig.p = as.data.frame(cmat.t$P)
cor.t.adj = 1 - cor.t.orig
cor.t.dist = as.dist(cor.t.adj)
distances.t = as.vector(cor.t.dist)

#no trauma
cmat.nt = rcorr(as.matrix(counts_wide.nt), type=rtype)
cor.nt.orig = as.data.frame(cmat.nt$r)
cor.nt.orig.p = as.data.frame(cmat.nt$P)
cor.nt.adj = 1 - cor.nt.orig
cor.nt.dist = as.dist(cor.nt.adj)
distances.nt = as.vector(cor.nt.dist)

```

## Define AHN correlations
```{r}

#isolate ahn correlations and sort factors by r
ahn_idx =  which(names(cor.t.orig)=='AHN')

#create df for control with region, r, and p
cor.ctrl.ahn = cor.ctrl.orig[-ahn_idx,'AHN', drop=FALSE]
cor.ctrl.ahn$p = cor.ctrl.orig.p[-ahn_idx,'AHN']
cor.ctrl.ahn = cbind(rownames(cor.ctrl.ahn), cor.ctrl.ahn)
cor.ctrl.ahn$p_fdr = p.adjust(cor.ctrl.ahn$p, 'fdr')
rownames(cor.ctrl.ahn) = NULL
names(cor.ctrl.ahn) = c('abrv','r','p','p_fdr')
cor.ctrl.ahn$group = 'Control'

#create df for trauma with region, r, and p
cor.t.ahn = cor.t.orig[-ahn_idx,'AHN', drop=FALSE]
cor.t.ahn$p = cor.t.orig.p[-ahn_idx,'AHN']
cor.t.ahn = cbind(rownames(cor.t.ahn), cor.t.ahn)
cor.t.ahn$p_fdr = p.adjust(cor.t.ahn$p, 'fdr')
rownames(cor.t.ahn) = NULL
names(cor.t.ahn) = c('abrv','r','p','p_fdr')
cor.t.ahn$group = 'Trauma'

#create df for no trauma with region, r, and p
cor.nt.ahn = cor.nt.orig[-ahn_idx,'AHN', drop=FALSE]
cor.nt.ahn$p = cor.nt.orig.p[-ahn_idx,'AHN']
cor.nt.ahn = cbind(rownames(cor.nt.ahn), cor.nt.ahn)
cor.nt.ahn$p_fdr = p.adjust(cor.nt.ahn$p, 'fdr')
rownames(cor.nt.ahn) = NULL
names(cor.nt.ahn) = c('abrv','r','p','p_fdr')
cor.nt.ahn$group = 'No Trauma'

#define abrv factor and order by group corr values
cor.ctrl.ahn$abrv = factor(
  cor.ctrl.ahn$abrv, 
  levels=cor.ctrl.ahn$abrv[order(cor.ctrl.ahn$r, decreasing=FALSE)])
cor.ctrl.ahn$sig = cor.ctrl.ahn$p<.05
cor.ctrl.ahn$sig_fdr = cor.ctrl.ahn$p_fdr<.05

cor.t.ahn$abrv = factor(
  cor.t.ahn$abrv, 
  levels=cor.t.ahn$abrv[order(cor.t.ahn$r, decreasing=FALSE)])
cor.t.ahn$sig = cor.t.ahn$p<.05
cor.t.ahn$sig_fdr = cor.t.ahn$p_fdr<.05

cor.nt.ahn$abrv = factor(
  cor.nt.ahn$abrv, 
  levels=cor.nt.ahn$abrv[order(cor.nt.ahn$r, decreasing=FALSE)])
cor.nt.ahn$sig = cor.nt.ahn$p<.05
cor.nt.ahn$sig_fdr = cor.nt.ahn$p_fdr<.05

```


## Compare AHN corrs
```{r}


#compare proportion of sig corrs using fisher's exact test
cor.ahn.ntVSt.fish = fisher.test(
  rbind(
    table(cor.t.ahn$sig & cor.t.ahn$r>0),
    table(cor.nt.ahn$sig & cor.nt.ahn$r>0)
  )
)
cor.ahn.ntVSt.fish

cor.ahn.ctrlVSnt.fish = fisher.test(
  rbind(
    table(cor.ctrl.ahn$sig & cor.ctrl.ahn$r>0),
    table(cor.nt.ahn$sig & cor.nt.ahn$r>0)
  )
)
cor.ahn.ctrlVSnt.fish

#compare proportion of sig corrs using fisher's exact on fdr corrected corrs
cor.ahn.ntVSt.fish.fdr = fisher.test(
  rbind(
    table(cor.t.ahn$sig_fdr & cor.t.ahn$r>0),
    table(cor.nt.ahn$sig_fdr & cor.nt.ahn$r>0)
  )
)
cor.ahn.ntVSt.fish.fdr

```

## plot AHN corrs
### set abrvs to display
```{r}

#define which regions to show
corahn_abrv_display = c(
      'LHA',
    #'CA3',
    'CA1',
    #'SUB',
    #'ENTm',
    #'ACAd2/3',
    #'DG-po',
    #'MEA',
    #'PAG',
    'BLAv',
    #'sAMY',
    'BLAa',
    #'BMA',
    #'PB',
    'ILA2/3'
    #'PB'
    #'DG-mo'
    #'ACAv2/3'
)

cor.ctrl.ahn$abrv_dis = cor.ctrl.ahn$abrv %in% c(corahn_abrv_display)
cor.t.ahn$abrv_dis = cor.t.ahn$abrv %in% c(corahn_abrv_display)
cor.nt.ahn$abrv_dis = cor.nt.ahn$abrv %in% c(corahn_abrv_display)


```


#### plot CTRL
```{r}

labeldata = cor.ctrl.ahn[cor.ctrl.ahn$abrv_dis,]

cor.ctrl.ahn.plot = ggplot(
  data = cor.ctrl.ahn,
  aes(x=abrv, y=r, color=sig)
)

cor.ctrl.ahn.plot +
  
  geom_bar(
    stat='identity',
    position = position_dodge(width=1),
    show.legend = FALSE
  ) +
  
  geom_point(
    data = labeldata,
    size = 1.5,
    color = 'black',
    show.legend = FALSE
  ) +
  
  geom_text_repel(
    data = labeldata,
    aes(label=abrv),
    force_pull = 4,
    min.segment.length = 0.5,
    fontface = 'plain',
    color = 'black',
    size = 6,
  ) +
  
  geom_hline(yintercept=0, linetype='dashed') +

  scale_color_manual(values=c('gray75','red2')) +
  
  coord_cartesian(ylim=c(-1,1)) +
  scale_y_continuous(breaks=seq(-1,1,1), labels=seq(-1,1,1), expand = c(0, .2)) +
  scale_x_discrete(expand = c(.05, .1)) +
  
  labs(
    x = 'Brain Region',
    y = 'R (AHN x Region)',
    title = 'CTRL'
  ) +
  
  theme_classic() +

  theme(
    plot.title = element_text(size=40,face='bold', hjust=.5),
    axis.title = element_text(size=20,face='plain', hjust=.5),
    axis.title.x = element_text(margin = margin(t=10)),
    axis.text.y = element_text(size=16, face='plain', color='black'),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    plot.margin = margin(c(1.75, 1, 1.75, 1), unit='cm'),
    plot.background = element_rect(fill = 'transparent',colour = NA),
  ) 
  
if (p.save == TRUE) {saveplot(filename = paste(f_pfx,'_',"AHN_corrs_CTRL", p.savetype, sep = ''))}

```


#### plot T
```{r}

labeldata = cor.t.ahn[cor.t.ahn$abrv_dis,]

cor.t.ahn.plot = ggplot(
  data = cor.t.ahn,
  aes(x=abrv, y=r, color=sig)
)

cor.t.ahn.plot +
  
  geom_bar(
    stat='identity',
    position = position_dodge(width=1),
    show.legend = FALSE
  ) +
  
  geom_point(
    data = labeldata,
    size = 1.5,
    color = 'black',
    show.legend = FALSE
  ) +
  
  geom_text_repel(
    data = labeldata,
    aes(label=abrv),
    force_pull = 4,
    min.segment.length = 0.5,
    fontface = 'plain',
    color = 'black',
    size = 6,
  ) +
  
  geom_hline(yintercept=0, linetype='dashed') +

  scale_color_manual(values=c('gray75','red2')) +
  
  coord_cartesian(ylim=c(-1,1)) +
  scale_y_continuous(breaks=seq(-1,1,1), labels=seq(-1,1,1), expand = c(0, .2)) +
  scale_x_discrete(expand = c(.05, .1)) +
  
  labs(
    x = 'Brain Region',
    y = 'R (AHN x Region)',
    title = 'S'
  ) +
  
  theme_classic() +

  theme(
    plot.title = element_text(size=40,face='bold', hjust=.5),
    axis.title = element_text(size=20,face='plain', hjust=.5),
    axis.title.x = element_text(margin = margin(t=10)),
    axis.text.y = element_text(size=16, face='plain', color='black'),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    plot.margin = margin(c(1.75, 1, 1.75, 1), unit='cm'),
    plot.background = element_rect(fill = 'transparent',colour = NA),
  ) 
  
if (p.save == TRUE) {saveplot(filename = paste(f_pfx,'_',"AHN_corrs_T", p.savetype, sep = ''))}

```


#### plot NT
```{r}

labeldata = cor.nt.ahn[cor.nt.ahn$abrv_dis,]

cor.nt.ahn.plot = ggplot(
  data = cor.nt.ahn,
  aes(x=abrv, y=r, color=sig)
)

cor.nt.ahn.plot +
  
  geom_bar(
    stat='identity',
    position = position_dodge(width=1),
    show.legend = FALSE
  ) +
  
  geom_point(
    data = labeldata,
    size = 1.5,
    color = 'black',
    show.legend = FALSE
  ) +
  
  geom_text_repel(
    data = labeldata,
    aes(label=abrv),
    force_pull = 4,
    min.segment.length = 0.5,
    fontface = 'plain',
    color = 'black',
    size = 6,
  ) +
  
  geom_hline(yintercept=0, linetype='dashed') +

  scale_color_manual(values=c('gray75','red2')) +
  
  coord_cartesian(ylim=c(-1,1)) +
  scale_y_continuous(breaks=seq(-1,1,1), labels=seq(-1,1,1), expand = c(0, .2)) +
  scale_x_discrete(expand = c(.05, .1)) +
  
  labs(
    x = 'Brain Region',
    y = 'R (AHN x Region)',
    title = 'NS'
  ) +
  
  theme_classic() +

  theme(
    plot.title = element_text(size=40,face='bold', hjust=.5),
    axis.title = element_text(size=20,face='plain', hjust=.5),
    axis.title.x = element_text(margin = margin(t=10)),
    axis.text.y = element_text(size=16, face='plain', color='black'),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    plot.margin = margin(c(1.75, 1, 1.75, 1), unit='cm'),
    plot.background = element_rect(fill = 'transparent',colour = NA),
  ) 
  
if (p.save == TRUE) {saveplot(filename = paste(f_pfx,'_',"AHN_corrs_NT", p.savetype, sep = ''))}

```



## Define ID info for used regions
```{r}

id_labels = as.data.frame(list('annotation_abrv' = names(counts_wide.t)))

id_labels = left_join(
  id_labels,
  labels[,c('annotation_abrv','annotation_graph_order','parcellation_group')],
  by = 'annotation_abrv'
)

#Set colors for parcellation groups
id_labels$color = recode_factor(
  id_labels$parcellation_group,
  cortical_plate = 'red',
  cortical_subplate = 'blue',
  cerebral_nuclei = 'yellow',
  diencephalon = 'green4',
  midbrain = 'steelblue3',
  hindbrain = 'gray'
)

```


# Clustering
## Perform initial clustering and view by higher-order parcellation
```{r}

hclust = hclust(cor.t.dist, method = 'average')
dend = as.dendrogram(hclust)

par(mar = c(10,5,5,5))

dend %>% 
  set('labels','') %>% 
  set('branches_lwd', 1.2) %>%
  set('branches_col', 'black') %>%
  hang.dendrogram(hang=1) %>%
  plot(
    main='Heirarchical Clustering of Regions',
    cex.main=1.5,
    ylim=c(0,1.1), 
    yaxt='n')

mtext(
  text='Subdivision',
  line = 1,
  side=1,
  padj = 2,
  cex=1.5, col='black')

colored_bars(
  id_labels$color, dend,
  y_scale=.2,
  y_shift = 0,
  text_shift = -100)

```


## Examine impact of cluster height on number of clusters
```{r}

potential_cut = .5
cutoffs = seq(0, 1, .02)
num_groups = vector()

for (cut in cutoffs){
  Hcut <- cutree(hclust, h = 1 - cut)
  num_groups = append(num_groups, length(unique(Hcut)))
}

cutoff_impact = data.frame(cutoffs,num_groups)

ggplot(cutoff_impact, aes(x=cutoffs, y=num_groups)) + 
  geom_line() +
  coord_cartesian(expand = FALSE) +
  scale_x_continuous(breaks = seq(0,1,.1)) +
  geom_vline(xintercept = potential_cut, color = 'red') +
  labs(
    title = 'Impact of clustering cutoff on number of clusters',
    x = 'R cutoff',
    y = '# clusters'
  )

```


## Examine cluster membership with 'elbow method' 
```{r}

testclusts = 150

wss = fviz_nbclust(cor.t.adj, FUNcluster=hcut, method='wss', k.max = testclusts)
wss = wss$data

ggplot(wss, aes(x=clusters, y=y, group=1)) + 
  geom_line(size=1.5) +
  geom_vline(xintercept = 26, color='red') +
  scale_x_discrete(breaks = seq(0, testclusts, 25)) +
  scale_y_continuous(breaks = seq(0, 22000, 5000), limits=c(0,22500)) +
  labs(
    x = '# of Clusters',
    y = 'Within-Cluster\nVariance',
    title = 'Cluster Selection'
  ) +
  
  theme_classic() +
  
  theme(
    plot.title = element_text(size=28,face='bold', hjust=.5),
    axis.title = element_text(size=18,face='bold', hjust=.5),
    axis.text.x = element_text(size=12, face='bold', color='black'),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    plot.margin = margin(c(2, 2, 2, 2), unit='cm'),
    plot.background = element_rect(fill = 'transparent',colour = NA),
  ) 
  
if (p.save == TRUE) {saveplot(filename = paste(f_pfx,'_',"ClusterElbow", p.savetype, sep = ''))}

```


## Define cutoff for clusters
```{r}

cor_cutoff = .5

```


## Cut cluster tree and visualize
```{r}

dist_cut = 1 - cor_cutoff

hclust = hclust(cor.t.dist, method = 'average')
Hcut <- cutree(hclust, h = dist_cut)
id_labels$cluster = Hcut
dend = as.dendrogram(hclust)

num_regions = nrow(cor.t.orig)
num_clust = length(unique(Hcut))
reduction = ((num_regions-num_clust)/num_regions)*100
to_print = c(
  paste('r cutoff: ', cor_cutoff),
  paste('number of regions reduced from ', num_regions, ' to ', num_clust),
  paste('percent reduction: ', reduction)
)

par(mar = c(5,5,5,5))
dend %>% 
  hang.dendrogram(hang=.01) %>%
  set('labels','') %>% 
  plot(
    main='Clustering Dendrogram\n Combine Clusters Below Line',
    ylim=c(0,1.1),yaxt='n')
mtext(
  text='Subdivision', 
  line = .5,
  side=1,
  cex=1, col='black')
abline(h = dist_cut, lwd=2, col = 'red')

print(to_print)

```


## Create Circular dendrogram
manually export at 5 by 5 inch
```{r}

#implement clustering and cut tree
hc = hclust(cor.t.dist, method = 'average')
ct = cutree(hc, h=1-cor_cutoff)
dend = as.dendrogram(hc)
dend_height = attr(dend, 'height')
lbls = labels(dend)
n = length(lbls)

#define plot colors
set.seed = NULL
set.seed(1)
cols = sample(
  #Earth, Cividis, 
  attr(colorRamp2(breaks=seq(1:max(ct)), hcl_palette = 'OrRd'),'colors'),
  #attr(colorRamp2(breaks=seq(1:max(ct)), hcl_palette = 'cividis'),'colors'),
  size=max(ct), 
  replace=FALSE
)

#define cluster bounds
start = vector()
stop = vector()
ct.df = left_join(
  as.data.frame(list('annotation_abrv'=names(ct),'cluster'=ct)), 
  as.data.frame(list('annotation_abrv'=lbls, 'i' = 1:length(lbls))),
  by='annotation_abrv'
)
for (c in seq(max(ct))){
  start[c] = min(ct.df$i[ct.df$cluster==c])
  stop[c] = max(ct.df$i[ct.df$cluster==c])
}
clstr_bnds = as.data.frame(list('start'=start,'stop'=stop))
clstr_bnds$cluster = rownames(clstr_bnds)

#give each cluster color based upon start
clstr_bnds = clstr_bnds[order(clstr_bnds$start),]
clstr_bnds$cols = cols

#set labels to highlight
tlbls = rep('', length(lbls))
cluster3_highlit_regions = c(
  'CA1','CA3','ENTm','SUB',#'ENTl',
  'BLAa','BLAp','BMA','CEA','MEA',
  'COAp', 'COAa',
  'MEPO','AHN','LHA',
  #'SNc','SNr',
  'PB',
  'PAG'#,'PCG
  )
lbls_cluster3_highlit = sapply(lbls, FUN=function(x){x %in% cluster3_highlit_regions})
tlbls[lbls_cluster3_highlit] = lbls[lbls_cluster3_highlit]

#initialize
par(mar = c(0,0,0,0))
circos.par(cell.padding = c(0, 0, 0, 0))
circos.initialize('a', xlim = c(0, n)) # only one sector

#labels
circos.labels(
  niceFacing = FALSE,
  sectors = 'a',
  x = which(tlbls!=''),
  labels = tlbls[which(tlbls!='')],
  side = 'outside',
  cex = 1,
  font = 2
)

#rectangular track
circos.track(
    ylim=c(0,1), 
    bg.border=NA, 
    track.height=.15,
    panel.fun = function(x,y) {
      for (i in 1:nrow(clstr_bnds)){
        circos.rect(
          clstr_bnds$start[clstr_bnds$cluster==i], 0, 
          clstr_bnds$stop[clstr_bnds$cluster==i], 1,
          col=clstr_bnds$cols[clstr_bnds$cluster==i])
      }
    }
)

#dendrogram
dend = color_branches(
  dend %>% set('branches_lwd', 1.1),
  h=1-cor_cutoff, 
  col=cols
)
circos.track(
  #ylim = c(.3, dend_height), 
  ylim = c(0, dend_height), 
  bg.border = NA, 
  track.height = .45, 
  panel.fun = function(x, y) {
    circos.dendrogram(dend)
  }
)

circos.clear()

```


# Extract AHN cluster Correlations
```{r}

ahn_clust = 3 #note that cluster number is arbitrary and renamed to 1 for manuscript.
ahn_clust_regions = id_labels$annotation_abrv[id_labels$cluster==3]

cor.ctrl.clusterAHN = cor.ctrl.orig[ahn_clust_regions, ahn_clust_regions]
cor.t.clusterAHN = cor.t.orig[ahn_clust_regions, ahn_clust_regions]
cor.nt.clusterAHN = cor.nt.orig[ahn_clust_regions, ahn_clust_regions]

cor.ctrl.clusterAHN.long = pivot_longer(
  data = cbind(list('Reg1'=rownames(cor.ctrl.clusterAHN)), cor.ctrl.clusterAHN),
  cols = names(cor.ctrl.clusterAHN),
  names_to = 'Reg2',
  values_to = 'R'
)
cor.t.clusterAHN.long = pivot_longer(
  data = cbind(list('Reg1'=rownames(cor.t.clusterAHN)), cor.t.clusterAHN),
  cols = names(cor.t.clusterAHN),
  names_to = 'Reg2',
  values_to = 'R'
)
cor.nt.clusterAHN.long = pivot_longer(
  data = cbind(list('Reg1'=rownames(cor.nt.clusterAHN)), cor.nt.clusterAHN),
  cols = names(cor.nt.clusterAHN),
  names_to = 'Reg2',
  values_to = 'R'
)

## add p values so they can be thresholded and plotted
## get p values
cor.ctrl.clusterAHN.p = cor.ctrl.orig.p[ahn_clust_regions, ahn_clust_regions]
cor.t.clusterAHN.p = cor.t.orig.p[ahn_clust_regions, ahn_clust_regions]
cor.nt.clusterAHN.p = cor.nt.orig.p[ahn_clust_regions, ahn_clust_regions]
##
cor.ctrl.clusterAHN.p.long = pivot_longer(
  data = cbind(list('Reg1'=rownames(cor.ctrl.clusterAHN.p)), cor.ctrl.clusterAHN.p),
  cols = names(cor.ctrl.clusterAHN.p),
  names_to = 'Reg2',
  values_to = 'p'
)
cor.t.clusterAHN.p.long = pivot_longer(
  data = cbind(list('Reg1'=rownames(cor.t.clusterAHN.p)), cor.t.clusterAHN.p),
  cols = names(cor.t.clusterAHN.p),
  names_to = 'Reg2',
  values_to = 'p'
)
cor.nt.clusterAHN.p.long = pivot_longer(
  data = cbind(list('Reg1'=rownames(cor.nt.clusterAHN.p)), cor.nt.clusterAHN.p),
  cols = names(cor.nt.clusterAHN.p),
  names_to = 'Reg2',
  values_to = 'p'
)

#combine p with r values and define color dummy strategy
cor.ctrl.clusterAHN.long = left_join(
  cor.ctrl.clusterAHN.long, 
  cor.ctrl.clusterAHN.p.long, by=c('Reg1','Reg2')
)
cor.ctrl.clusterAHN.long$p[is.na(cor.ctrl.clusterAHN.long$p)]=0
cor.ctrl.clusterAHN.long$sig = cor.ctrl.clusterAHN.long$p<0.05
cor.ctrl.clusterAHN.long = mutate(
  cor.ctrl.clusterAHN.long, sigcode = case_when(
    sig == TRUE & R>0 ~ 1,
    sig == TRUE & R<0 ~ -1,
    sig == FALSE ~ 0,
  )
)

cor.t.clusterAHN.long = left_join(
  cor.t.clusterAHN.long, 
  cor.t.clusterAHN.p.long, by=c('Reg1','Reg2')
)
cor.t.clusterAHN.long$sig = cor.t.clusterAHN.long$p<0.05
cor.t.clusterAHN.long$p[is.na(cor.t.clusterAHN.long$p)]=0
cor.t.clusterAHN.long$sig = cor.t.clusterAHN.long$p<0.05
cor.t.clusterAHN.long = mutate(
  cor.t.clusterAHN.long, sigcode = case_when(
    sig == TRUE & R>0 ~ 1,
    sig == TRUE & R<0 ~ -1,
    sig == FALSE ~ 0,
  )
)

cor.nt.clusterAHN.long = left_join(
  cor.nt.clusterAHN.long, 
  cor.nt.clusterAHN.p.long, by=c('Reg1','Reg2')
)
cor.nt.clusterAHN.long$sig = cor.nt.clusterAHN.long$p<0.05
cor.nt.clusterAHN.long$p[is.na(cor.nt.clusterAHN.long$p)]=0
cor.nt.clusterAHN.long$sig = cor.nt.clusterAHN.long$p<0.05
cor.nt.clusterAHN.long = mutate(
  cor.nt.clusterAHN.long, sigcode = case_when(
    sig == TRUE & R>0 ~ 1,
    sig == TRUE & R<0 ~ -1,
    sig == FALSE ~ 0,
  )
)


#this is used to sort original correlation matrices by ahn cluster membership
id_labels$clusterAHN = id_labels$cluster == ahn_clust
id_labels$annotation_abrv = factor(
  id_labels$annotation_abrv,
  levels = id_labels$annotation_abrv[order(
    id_labels$clusterAHN,
    id_labels$cluster,
    id_labels$annotation_abrv
  )]
)
id_labels = id_labels[order(id_labels$annotation_abrv),]


#factor regions
cor.ctrl.clusterAHN.long$Reg1 = factor(
  cor.ctrl.clusterAHN.long$Reg1,
  levels = id_labels$annotation_abrv[id_labels$clusterAHN]
)
cor.ctrl.clusterAHN.long$Reg2 = factor(
  cor.ctrl.clusterAHN.long$Reg2,
  levels = id_labels$annotation_abrv[id_labels$clusterAHN]
)

```


## Plot AHN cluster correlations
### Control
#### R
```{r}

p.clusterAHNcor.ctrl = ggplot(
  data = cor.ctrl.clusterAHN.long,
  aes(x=Reg1, y=Reg2, fill=R)
)

p.clusterAHNcor.ctrl +
  geom_tile() +
  scale_fill_gradient2(
    low='blue2', high='red3', midpoint=0, breaks=seq(-1,1,1),
    limits=c(-1,1)
  ) +
  labs(
    title = 'CTRL',
    x = 'C1 Region',
    y = 'C1 Region',
  )+

  theme(
    plot.title = element_text(size=40,face='bold', hjust=.5),
    axis.title = element_text(size=24,face='bold', hjust=.5),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    legend.text = element_text(size = 12, face = 'bold'),
    legend.title = element_text(size = 24, face = 'bold'),
    line = element_line(size=5),
    plot.background = element_rect(fill = 'transparent',colour = NA),
    panel.border = element_rect(size=1, fill='transparent')
  ) +
  force_panelsizes(
    cols = unit(3, "in"),
    rows = unit(3, "in")
  )

#if (p.save == TRUE) {saveplot(filename = paste(f_pfx,'_',"AHNclstr_corrs_CTRL", p.savetype.corr, sep = ''))}

```


#### Sig
```{r}

cor.ctrl.clusterAHN.long$sigcodeF = recode_factor(
  factor(cor.ctrl.clusterAHN.long$sigcode,levels = c(1,0,-1)),
  '1' = 'pos.',
  '0' = 'none',
  '-1' = 'neg.'
)

p.clusterAHNcorsig.ctrl = ggplot(
  data = cor.ctrl.clusterAHN.long,
  aes(x=Reg1, y=Reg2, fill=sigcodeF)
)

p.clusterAHNcorsig.ctrl +
  geom_tile(show.legend = FALSE) +
  scale_fill_manual(values=c('red3','white','blue2')) +
  labs(
    title = 'CTRL',
    x = 'C1 Region',
    y = 'C1 Region',
  )+
  theme(
    plot.title = element_text(size=40,face='bold', hjust=.5),
    axis.title = element_text(size=24,face='plain', hjust=.5),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    plot.background = element_rect(fill = 'transparent',colour = NA),
    panel.border = element_rect(size=1, fill='transparent')
  ) +
  force_panelsizes(
    cols = unit(3, "in"),
    rows = unit(3, "in")
  )

if (p.save == TRUE) {
  saveplot(filename = paste(f_pfx,'_',"AHNclstr_corrs_sig_CTRL", p.savetype.corr, sep = ''))
  #saveplot(filename = paste(f_pfx,'_',"AHNclstr_corrs_sig_CTRL", p.savetype.corr, sep = ''))
}

```

### Trauma
#### R
```{r}

p.clusterAHNcor.t = ggplot(
  data = cor.t.clusterAHN.long,
  aes(x=Reg1, y=Reg2, fill=R)
)

p.clusterAHNcor.t +
  geom_tile() +
  scale_fill_gradient2(
    low='blue2', high='red3', midpoint=0, breaks=seq(-1,1,1),
    limits=c(-1,1)
  ) +
  labs(
    title = 'S',
    x = 'C1 Region',
    y = 'C1 Region',
  )+
  theme(
    plot.title = element_text(size=40,face='bold', hjust=.5),
    axis.title = element_text(size=24,face='bold', hjust=.5),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    legend.text = element_text(size = 12, face = 'bold'),
    legend.title = element_text(size = 24, face = 'bold'),
    plot.background = element_rect(fill = 'transparent',colour = NA),
    panel.border = element_rect(size=1, fill='transparent')
  ) +
  force_panelsizes(
    cols = unit(3, "in"),
    rows = unit(3, "in")
  )

#if (p.save == TRUE) {saveplot(filename = paste(f_pfx,'_',"AHNclstr_corrs_T", p.savetype.corr, sep = ''))}

```

#### Sig
```{r}

cor.t.clusterAHN.long$sigcodeF = recode_factor(
  factor(cor.t.clusterAHN.long$sigcode,levels = c(1,0,-1)),
  '1' = 'pos.',
  '0' = 'none',
  '-1' = 'neg.'
)

p.clusterAHNcorsig.t = ggplot(
  data = cor.t.clusterAHN.long,
  aes(x=Reg1, y=Reg2, fill=sigcodeF)
)

p.clusterAHNcorsig.t +
  geom_tile(show.legend = FALSE) +
  scale_fill_manual(values=c('red3','white','blue2')) +
  labs(
    title = 'S',
    x = 'C1 Region',
    y = 'C1 Region',
  )+
  theme(
    plot.title = element_text(size=40,face='bold', hjust=.5),
    axis.title = element_text(size=24,face='plain', hjust=.5),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    plot.background = element_rect(fill = 'transparent',colour = NA),
    panel.border = element_rect(size=1, fill='transparent')
  ) +
  force_panelsizes(
    cols = unit(3, "in"),
    rows = unit(3, "in")
  )

if (p.save == TRUE) {
  saveplot(filename = paste(f_pfx,'_',"AHNclstr_corrs_sig_T", p.savetype.corr, sep = ''))
  saveplot(filename = paste(f_pfx,'_',"AHNclstr_corrs_sig_T", p.savetype, sep = ''))
}
#save both ways for plotting in fig1

```

### No Trauma
#### R
```{r}

p.clusterAHNcor.nt = ggplot(
  data = cor.nt.clusterAHN.long,
  aes(x=Reg1, y=Reg2, fill=R)
)

p.clusterAHNcor.nt +
  geom_tile() +
  scale_fill_gradient2(
    low='blue2', high='red3', midpoint=0, breaks=seq(-1,1,1),
    limits=c(-1,1)
  ) +
  labs(
    title = 'NS',
    x = 'C1 Region',
    y = 'C1 Region',
  )+
  theme(
    plot.title = element_text(size=40,face='bold', hjust=.5),
    axis.title = element_text(size=24,face='bold', hjust=.5),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    legend.text = element_text(size = 12, face = 'bold'),
    legend.title = element_text(size = 24, face = 'bold'),
    plot.background = element_rect(fill = 'transparent',colour = NA),
    panel.border = element_rect(size=1, fill='transparent')
  ) +
  force_panelsizes(
    cols = unit(3, "in"),
    rows = unit(3, "in")
  )

#if (p.save == TRUE) {saveplot(filename = paste(f_pfx,'_',"AHNclstr_corrs_NT", p.savetype.corr, sep = ''))}

```

#### Sig
```{r}

cor.nt.clusterAHN.long$sigcodeF = recode_factor(
  factor(cor.nt.clusterAHN.long$sigcode,levels = c(1,0,-1)),
  '1' = 'pos.',
  '0' = 'none',
  '-1' = 'neg.'
)

p.clusterAHNcorsig.nt = ggplot(
  data = cor.nt.clusterAHN.long,
  aes(x=Reg1, y=Reg2, fill=sigcodeF)
)

p.clusterAHNcorsig.nt +
  geom_tile(show.legend = FALSE) +
  scale_fill_manual(values=c('red3','white','blue2')) +
  labs(
    title = 'NS',
    x = 'C1 Region',
    y = 'C1 Region',
  )+
  theme(
    plot.title = element_text(size=40,face='bold', hjust=.5),
    axis.title = element_text(size=24,face='plain', hjust=.5),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    legend.text = element_text(size = 12, face = 'bold'),
    legend.title = element_blank(),
    plot.background = element_rect(fill = 'transparent',colour = NA),
    panel.border = element_rect(size=1, fill='transparent')
  ) +
  force_panelsizes(
    cols = unit(3, "in"),
    rows = unit(3, "in")
  )

if (p.save == TRUE) {
  saveplot(filename = paste(f_pfx,'_',"AHNclstr_corrs_sig_NT", p.savetype.corr, sep = ''))
  saveplot(filename = paste(f_pfx,'_',"AHNclstr_corrs_sig_NT", p.savetype, sep = ''))
}
#save both ways for plotting in fig1

```

## Plot all correlations
### Ctrl
#### Prep Data
```{r}

#pivot to long and add p values
cor.ctrl.orig.long = pivot_longer(
  data = cbind(list('Reg1'=rownames(cor.ctrl.orig)), cor.ctrl.orig),
  cols = names(cor.ctrl.orig),
  names_to = 'Reg2',
  values_to = 'R'
)
cor.ctrl.orig.p.long = pivot_longer(
  data = cbind(list('Reg1'=rownames(cor.ctrl.orig.p)), cor.ctrl.orig.p),
  cols = names(cor.ctrl.orig.p),
  names_to = 'Reg2',
  values_to = 'p'
)
cor.ctrl.orig.long = left_join(cor.ctrl.orig.long, cor.ctrl.orig.p.long, by=c('Reg1','Reg2'))

#define significance coding variable
cor.ctrl.orig.long$p[is.na(cor.ctrl.orig.long$p)]=0
cor.ctrl.orig.long$sig = cor.ctrl.orig.long$p<0.05
cor.ctrl.orig.long = mutate(
  cor.ctrl.orig.long, sigcode = case_when(
    sig == TRUE & R>0 ~ 1,
    sig == TRUE & R<0 ~ -1,
    sig == FALSE ~ 0,
  )
)

#sort Reg 1 and Reg2 by ahn cluster membership
cor.ctrl.orig.long$Reg1 = factor(cor.ctrl.orig.long$Reg1, levels=id_labels$annotation_abrv)
cor.ctrl.orig.long$Reg2 = factor(cor.ctrl.orig.long$Reg2, levels=id_labels$annotation_abrv)

```

#### R
```{r}

p.ALLcor.ctrl = ggplot(
  data = cor.ctrl.orig.long,
  aes(x=Reg1, y=Reg2, fill=R)
)

p.ALLcor.ctrl +
  geom_tile() +
  scale_fill_gradient2(
    low='blue2', high='red3', midpoint=0, breaks=seq(-1,1,1),
    limits=c(-1,1)
  ) +
  labs(
    title = 'CTRL',
    x = 'Brain Region',
    y = 'Brain Region',
  )+
  theme(
    plot.title = element_text(size=40,face='bold', hjust=.5),
    axis.title = element_text(size=24,face='bold', hjust=.5),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    legend.text = element_text(size = 12, face = 'bold'),
    legend.title = element_text(size = 24, face = 'bold'),
    plot.background = element_rect(fill = 'transparent',colour = NA),
    panel.border = element_rect(size=1, fill='transparent')
  ) +
  force_panelsizes(
    cols = unit(3, "in"),
    rows = unit(3, "in")
  )

#if (p.save == TRUE) {saveplot(filename = paste(f_pfx,'_',"Allcorrs_CTRL", p.savetype.corr, sep = ''))}

```
#### Sig
```{r}

cor.ctrl.orig.long$sigcodeF = recode_factor(
  factor(cor.ctrl.orig.long$sigcode,levels = c(1,0,-1)),
  '1' = 'pos.',
  '0' = 'none',
  '-1' = 'neg.'
)

p.ALLcorsig.ctrl = ggplot(
  data = cor.ctrl.orig.long,
  aes(x=Reg1, y=Reg2, fill=sigcodeF)
)

p.ALLcorsig.ctrl +
  geom_tile(show.legend = FALSE) +
  geom_rect(
    xmin = id_labels$annotation_abrv[id_labels$clusterAHN][1],
    xmax = id_labels$annotation_abrv[id_labels$clusterAHN][sum(id_labels$clusterAHN)],
    ymin = id_labels$annotation_abrv[id_labels$clusterAHN][1],
    ymax = id_labels$annotation_abrv[id_labels$clusterAHN][sum(id_labels$clusterAHN)],
    color = 'green3', fill = NA, linetype = 'solid', linewidth = 3
  ) +
  scale_fill_manual(values=c('red3','white','blue2')) +
  labs(
    title = 'CTRL',
    x = 'Brain Region',
    y = 'Brain Region',
  )+
  theme(
    plot.title = element_text(size=40,face='bold', hjust=.5),
    axis.title = element_text(size=24,face='bold', hjust=.5),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    legend.text = element_text(size = 12, face = 'bold'),
    legend.title = element_blank(),
    plot.background = element_rect(fill = 'transparent',colour = NA),
    panel.border = element_rect(size=1, fill='transparent')
  ) +
  force_panelsizes(
    cols = unit(3, "in"),
    rows = unit(3, "in")
  )

if (p.save == TRUE) {saveplot(filename = paste(f_pfx,'_',"Allcorrs_sig_CTRL", p.savetype.corr, sep = ''))}

```


### Trauma
#### Prep Data
```{r}

cor.t.orig.long = pivot_longer(
  data = cbind(list('Reg1'=rownames(cor.t.orig)), cor.t.orig),
  cols = names(cor.t.orig),
  names_to = 'Reg2',
  values_to = 'R'
)
cor.t.orig.p.long = pivot_longer(
  data = cbind(list('Reg1'=rownames(cor.t.orig.p)), cor.t.orig.p),
  cols = names(cor.t.orig.p),
  names_to = 'Reg2',
  values_to = 'p'
)
cor.t.orig.long = left_join(cor.t.orig.long, cor.t.orig.p.long, by=c('Reg1','Reg2'))

cor.t.orig.long$p[is.na(cor.t.orig.long$p)]=0
cor.t.orig.long$sig = cor.t.orig.long$p<0.05
cor.t.orig.long = mutate(
  cor.t.orig.long, sigcode = case_when(
    sig == TRUE & R>0 ~ 1,
    sig == TRUE & R<0 ~ -1,
    sig == FALSE ~ 0,
  )
)

#sort Reg 1 and Reg2 by ahn cluster membership
cor.t.orig.long$Reg1 = factor(cor.t.orig.long$Reg1, levels=id_labels$annotation_abrv)
cor.t.orig.long$Reg2 = factor(cor.t.orig.long$Reg2, levels=id_labels$annotation_abrv)

```
#### R
```{r}

p.ALLcor.t = ggplot(
  data = cor.t.orig.long,
  aes(x=Reg1, y=Reg2, fill=R)
)

p.ALLcor.t +
  geom_tile() +
  scale_fill_gradient2(
    low='blue2', high='red3', midpoint=0, breaks=seq(-1,1,1),
    limits=c(-1,1)
  ) +
  labs(
    title = 'S',
    x = 'Brain Region',
    y = 'Brain Region',
  )+
  theme(
    plot.title = element_text(size=40,face='bold', hjust=.5),
    axis.title = element_text(size=24,face='bold', hjust=.5),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    legend.text = element_text(size = 12, face = 'bold'),
    legend.title = element_text(size = 24, face = 'bold'),
    plot.background = element_rect(fill = 'transparent',colour = NA),
    panel.border = element_rect(size=1, fill='transparent')
  ) +
  force_panelsizes(
    cols = unit(3, "in"),
    rows = unit(3, "in")
  )

#if (p.save == TRUE) {saveplot(filename = paste(f_pfx,'_',"Allcorrs_T", p.savetype.corr, sep = ''))}

```

#### Sig
```{r}

cor.t.orig.long$sigcodeF = recode_factor(
  factor(cor.t.orig.long$sigcode,levels = c(1,0,-1)),
  '1' = 'pos.',
  '0' = 'none',
  '-1' = 'neg.'
)

p.ALLcorsig.t = ggplot(
  data = cor.t.orig.long,
  aes(x=Reg1, y=Reg2, fill=sigcodeF)
)

p.ALLcorsig.t +
  geom_tile(show.legend = FALSE) +
  geom_rect(
    xmin = id_labels$annotation_abrv[id_labels$clusterAHN][1],
    xmax = id_labels$annotation_abrv[id_labels$clusterAHN][sum(id_labels$clusterAHN)],
    ymin = id_labels$annotation_abrv[id_labels$clusterAHN][1],
    ymax = id_labels$annotation_abrv[id_labels$clusterAHN][sum(id_labels$clusterAHN)],
    color = 'green3', fill = NA, linetype = 'solid', linewidth = 3
  ) +
  scale_fill_manual(values=c('red3','white','blue2')) +
  labs(
    title = 'S',
    x = 'Brain Region',
    y = 'Brain Region',
  )+
  theme(
    plot.title = element_text(size=40,face='bold', hjust=.5),
    axis.title = element_text(size=24,face='bold', hjust=.5),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    legend.text = element_text(size = 12, face = 'bold'),
    legend.title = element_blank(),
    plot.background = element_rect(fill = 'transparent',colour = NA),
    panel.border = element_rect(size=1, fill='transparent')
  ) +
  force_panelsizes(
    cols = unit(3, "in"),
    rows = unit(3, "in")
  )

if (p.save == TRUE) {saveplot(filename = paste(f_pfx,'_',"Allcorrs_sig_T", p.savetype.corr, sep = ''))}

```

### No Trauma
#### Prep Data
```{r}

cor.nt.orig.long = pivot_longer(
  data = cbind(list('Reg1'=rownames(cor.nt.orig)), cor.nt.orig),
  cols = names(cor.nt.orig),
  names_to = 'Reg2',
  values_to = 'R'
)
cor.nt.orig.p.long = pivot_longer(
  data = cbind(list('Reg1'=rownames(cor.nt.orig.p)), cor.nt.orig.p),
  cols = names(cor.nt.orig.p),
  names_to = 'Reg2',
  values_to = 'p'
)
cor.nt.orig.long = left_join(cor.nt.orig.long, cor.nt.orig.p.long, by=c('Reg1','Reg2'))

cor.nt.orig.long$p[is.na(cor.nt.orig.long$p)]=0
cor.nt.orig.long$sig = cor.nt.orig.long$p<0.05
cor.nt.orig.long = mutate(
  cor.nt.orig.long, sigcode = case_when(
    sig == TRUE & R>0 ~ 1,
    sig == TRUE & R<0 ~ -1,
    sig == FALSE ~ 0,
  )
)

#sort Reg 1 and Reg2 by ahn cluster membership
cor.nt.orig.long$Reg1 = factor(cor.nt.orig.long$Reg1, levels=id_labels$annotation_abrv)
cor.nt.orig.long$Reg2 = factor(cor.nt.orig.long$Reg2, levels=id_labels$annotation_abrv)

```

#### R
```{r}

p.ALLcor.nt = ggplot(
  data = cor.nt.orig.long,
  aes(x=Reg1, y=Reg2, fill=R)
)

p.ALLcor.nt +
  geom_tile() +
  scale_fill_gradient2(
    low='blue2', high='red3', midpoint=0, breaks=seq(-1,1,1),
    limits=c(-1,1)
  ) +
  labs(
    title = 'NS',
    x = 'Brain Region',
    y = 'Brain Region',
  )+
  theme(
    plot.title = element_text(size=40,face='bold', hjust=.5),
    axis.title = element_text(size=24,face='bold', hjust=.5),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    legend.text = element_text(size = 12, face = 'bold'),
    legend.title = element_text(size = 24, face = 'bold'),
    plot.background = element_rect(fill = 'transparent',colour = NA),
    panel.border = element_rect(size=1, fill='transparent')
  ) +
  force_panelsizes(
    cols = unit(3, "in"),
    rows = unit(3, "in")
  )

#if (p.save == TRUE) {saveplot(filename = paste(f_pfx,'_',"Allcorrs_NT", p.savetype.corr, sep = ''))}

```
#### Sig
```{r}

cor.nt.orig.long$sigcodeF = recode_factor(
  factor(cor.nt.orig.long$sigcode,levels = c(1,0,-1)),
  '1' = 'pos.',
  '0' = 'none',
  '-1' = 'neg.'
)

p.ALLcorsig.nt = ggplot(
  data = cor.nt.orig.long,
  aes(x=Reg1, y=Reg2, fill=sigcodeF)
)

p.ALLcorsig.nt +
  geom_tile(show.legend = FALSE) +
  geom_rect(
    xmin = id_labels$annotation_abrv[id_labels$clusterAHN][1],
    xmax = id_labels$annotation_abrv[id_labels$clusterAHN][sum(id_labels$clusterAHN)],
    ymin = id_labels$annotation_abrv[id_labels$clusterAHN][1],
    ymax = id_labels$annotation_abrv[id_labels$clusterAHN][sum(id_labels$clusterAHN)],
    color = 'green3', fill = NA, linetype = 'solid', linewidth = 3
  ) +
  scale_fill_manual(values=c('red3','white','blue2')) +
  labs(
    title = 'NS',
    x = 'Brain Region',
    y = 'Brain Region',
  )+
  theme(
    plot.title = element_text(size=40,face='bold', hjust=.5),
    axis.title = element_text(size=24,face='bold', hjust=.5),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    legend.text = element_text(size = 12, face = 'bold'),
    legend.title = element_blank(),
    plot.background = element_rect(fill = 'transparent',colour = NA),
    panel.border = element_rect(size=1, fill='transparent')
  ) +
  force_panelsizes(
    cols = unit(3, "in"),
    rows = unit(3, "in")
  )

if (p.save == TRUE) {saveplot(filename = paste(f_pfx,'_',"Allcorrs_sig_NT", p.savetype.corr, sep = ''))}

```

## define mean correlation difference
```{r}

#obtain off diagonal correlations in vector, avoiding duplicates
cor.ctrl.clusterAHN.vec = cormat_returntop(cor.ctrl.clusterAHN)
cor.t.clusterAHN.vec = cormat_returntop(cor.t.clusterAHN)
cor.nt.clusterAHN.vec = cormat_returntop(cor.nt.clusterAHN)
cor.ctrl.orig.vec = cormat_returntop(cor.ctrl.orig)
cor.t.orig.vec = cormat_returntop(cor.t.orig)
cor.nt.orig.vec = cormat_returntop(cor.nt.orig)

print(paste('ctrl, mean intracluster correlation:', mean(cor.ctrl.clusterAHN.vec)))
print(paste('trauma, mean intracluster correlation:', mean(cor.t.clusterAHN.vec)))
print(paste('no trauma, mean intracluster correlation:', mean(cor.nt.clusterAHN.vec)))
print(paste('ctrl, mean correlation:', mean(cor.ctrl.orig.vec)))
print(paste('trauma, mean correlation:', mean(cor.t.orig.vec)))
print(paste('no trauma, mean correlation:', mean(cor.nt.orig.vec)))

```

# Test difference in cluster correlational strength between groups

## fisher exact of number of significant positive correltaions
```{r}

cor.ctrl.orig.p.vec = cormat_returntop(cor.ctrl.orig.p)
cor.nt.orig.p.vec = cormat_returntop(cor.nt.orig.p)
cor.t.orig.p.vec = cormat_returntop(cor.t.orig.p)

cor.ctrl.clusterAHN.p.vec = cormat_returntop(cor.ctrl.orig.p[ahn_clust_regions,ahn_clust_regions])
cor.nt.clusterAHN.p.vec = cormat_returntop(cor.nt.orig.p[ahn_clust_regions,ahn_clust_regions])
cor.t.clusterAHN.p.vec = cormat_returntop(cor.t.orig.p[ahn_clust_regions,ahn_clust_regions])

#
#compare proportion of sig pos corrs using chi square for ahn cluster
#
cor.ahncluster.ntVSt.fish = fisher.test(
  rbind(
    table(cor.nt.clusterAHN.p.vec<.05 & cor.nt.clusterAHN.vec>0),
    table(cor.t.clusterAHN.p.vec<.05 & cor.t.clusterAHN.vec>0)
  )
)
cor.ahncluster.ntVSt.fish

cor.ahncluster.ctrlVSnt.fish = fisher.test(
  rbind(
    table(cor.ctrl.clusterAHN.p.vec<.05 & cor.ctrl.clusterAHN.vec>0),
    table(cor.nt.clusterAHN.p.vec<.05 & cor.nt.clusterAHN.vec>0)
  )
)
cor.ahncluster.ctrlVSnt.fish

#
#compare proportion of sig corrs using chi square for all correlations
#
cor.all.ntVSt.fish = fisher.test(
  rbind(
    table(cor.nt.orig.p.vec<.05 & cor.nt.orig.vec>0),
    table(cor.t.orig.p.vec<.05 & cor.t.orig.vec>0)
  )
)
cor.all.ntVSt.fish

cor.all.ctrlVSnt.fish = fisher.test(
  rbind(
    table(cor.ctrl.orig.p.vec<.05 & cor.ctrl.orig.vec>0),
    table(cor.nt.orig.p.vec<.05 & cor.nt.orig.vec>0)
  )
)
cor.all.ctrlVSnt.fish


```

## Plot proportion of significant corrs 
### AHN cluster
```{r}

cor.clusterAHN.props = as.data.frame(prop.table(
  rbind(
      table(cor.ctrl.clusterAHN.p.vec <.05 & cor.ctrl.clusterAHN.vec >0),
      table(cor.nt.clusterAHN.p.vec <.05 & cor.nt.clusterAHN.vec >0),
      table(cor.t.clusterAHN.p.vec <.05 & cor.t.clusterAHN.vec >0)
  ),
  margin = 1
))
cor.clusterAHN.props$group = factor(c('Ctrl','NS','S'))
cor.clusterAHN.props = pivot_longer(
  cor.clusterAHN.props,
  cols = c('FALSE','TRUE'),
  names_to = 'sig'
)

```

#### No Ctrl
```{r}

P.bar = ggplot(
  cor.clusterAHN.props[cor.clusterAHN.props$group!='Ctrl',], 
  aes(x = group, y = value, fill = sig, label=sig)
) 

p.ymin = 0
p.ymax = 1
p.yint = .25
p.ybreaks = seq(p.ymin,p.ymax,p.yint)

P.bar +
  
  geom_col(
    colour = 'gray10',
    show.legend = FALSE
  ) + 
  
  force_panelsizes(
      rows = unit(2, 'in'),
      cols = unit(.6, "in")
  ) +

  labs(
    x = '',
    y = "% Significant R(+)"
  ) +

  scale_y_continuous(
    breaks = p.ybreaks, labels = p.ybreaks*100, expand = c(.005,.005),
    position = 'right'
  ) +

  scale_fill_manual(values=c('gray90','red3')) +
  theme_bw() +
  theme(
    panel.grid = element_blank(), 
    panel.border = element_rect(colour="white"),
    axis.text = element_text(size = 14, color='black'),
    axis.title = element_text(size = 14, color='black'),
    axis.line = element_line(linewidth=.5,color="black"),
    axis.line.x = element_blank(),
    plot.background = element_rect(fill = "transparent", colour = NA)
  )

if (p.save == TRUE) {saveplot(filename = paste(f_pfx,'_',"AHNclstr_ProportionR_noCtrl", p.savetype, sep = ''))}

```

#### with Ctrl
```{r}

P.bar = ggplot(
  cor.clusterAHN.props[,], 
  aes(x = group, y = value, fill = sig, label=sig)
) 

p.ymin = 0
p.ymax = 1
p.yint = .25
p.ybreaks = seq(p.ymin,p.ymax,p.yint)

P.bar +
  
  geom_col(
    colour = 'gray10',
    show.legend = FALSE
  ) + 
  
  force_panelsizes(
      rows = unit(2, 'in'),
      cols = unit(1, "in")
  ) +

  labs(
    x = '',
    y = "% Significant R(+)"
  ) +

  scale_y_continuous(
    breaks = p.ybreaks, labels = p.ybreaks*100, expand = c(.005,.005),
    position = 'right'
  ) +

  scale_fill_manual(values=c('gray90','red3')) +
  theme_bw() +
  theme(
    panel.grid = element_blank(), 
    panel.border = element_rect(colour="white"),
    axis.text = element_text(size = 14, color='black'),
    axis.title = element_text(size = 14, color='black'),
    axis.line = element_line(linewidth=.5,color="black"),
    axis.line.x = element_blank(),
    plot.background = element_rect(fill = "transparent", colour = NA)
  )

if (p.save == TRUE) {saveplot(filename = paste(f_pfx,'_',"AHNclstr_ProportionR_wCtrl", p.savetype, sep = ''))}

```


### All regions
```{r}

cor.orig.props = as.data.frame(prop.table(
  rbind(
      table(cor.ctrl.orig.p.vec <.05 & cor.ctrl.orig.vec >0),
      table(cor.nt.orig.p.vec <.05 & cor.nt.orig.vec >0),
      table(cor.t.orig.p.vec <.05 & cor.t.orig.vec >0)
  ),
  margin = 1
))
cor.orig.props$group = factor(c('Ctrl','NS','S'))
cor.orig.props = pivot_longer(
  cor.orig.props,
  cols = c('FALSE','TRUE'),
  names_to = 'sig'
)

```

#### No Ctrl
```{r}

P.bar = ggplot(
  cor.orig.props[cor.orig.props$group!='Ctrl',], 
  aes(x = group, y = value, fill = sig, label=sig)
) 

p.ymin = 0
p.ymax = 1
p.yint = .25
p.ybreaks = seq(p.ymin,p.ymax,p.yint)

P.bar +
  
  geom_col(
    colour = 'gray10',
    show.legend = FALSE
  ) + 
  
  force_panelsizes(
      rows = unit(2, 'in'),
      cols = unit(.6, "in")
  ) +

  labs(
    x = '',
    y = "% Significant R(+)"
  ) +

  scale_y_continuous(
    breaks = p.ybreaks, labels = p.ybreaks*100, expand = c(.005,.005),
    position = 'right'
  ) +

  scale_fill_manual(values=c('gray90','red3')) +
  theme_bw() +
  theme(
    panel.grid = element_blank(), 
    panel.border = element_rect(colour="white"),
    axis.text = element_text(size = 14, color='black'),
    axis.title = element_text(size = 14, color='black'),
    axis.line = element_line(linewidth=.5,color="black"),
    axis.line.x = element_blank(),
    plot.background = element_rect(fill = "transparent", colour = NA)
  )

if (p.save == TRUE) {saveplot(filename = paste(f_pfx,'_',"AllCor_ProportionR_noCtrl", p.savetype, sep = ''))}

```

#### with Ctrl
```{r}

P.bar = ggplot(
  cor.orig.props, 
  aes(x = group, y = value, fill = sig, label=sig)
) 

p.ymin = 0
p.ymax = 1
p.yint = .25
p.ybreaks = seq(p.ymin,p.ymax,p.yint)

P.bar +
  
  geom_col(
    colour = 'gray10',
    show.legend = FALSE
  ) + 
  
  force_panelsizes(
      rows = unit(2, 'in'),
      cols = unit(1, "in")
  ) +

  labs(
    x = '',
    y = "% Significant R(+)"
  ) +

  scale_y_continuous(
    breaks = p.ybreaks, labels = p.ybreaks*100, expand = c(.005,.005),
    position = 'right'
  ) +

  scale_fill_manual(values=c('gray90','red3')) +
  theme_bw() +
  theme(
    panel.grid = element_blank(), 
    panel.border = element_rect(colour="white"),
    axis.text = element_text(size = 14, color='black'),
    axis.title = element_text(size = 14, color='black'),
    axis.line = element_line(linewidth=.5,color="black"),
    axis.line.x = element_blank(),
    plot.background = element_rect(fill = "transparent", colour = NA)
  )

if (p.save == TRUE) {saveplot(filename = paste(f_pfx,'_',"AllCor_ProportionR_wCtrl", p.savetype, sep = ''))}

```




# Export figure data for publication
```{r}

export = FALSE

#export correlations between AHN and every other region
ahn_corrs_export = bind_rows(
  cor.ctrl.ahn, cor.nt.ahn, cor.t.ahn
)
ahn_corrs_export$group = recode_factor(
  factor(ahn_corrs_export$group),
  'Control' = 'Ctrl',
  'No Trauma' = 'NS',
  'Trauma' = 'S'
)
ahn_corrs_export = ahn_corrs_export[,c('group','abrv','r','p','p_fdr','sig','sig_fdr')]

#export cluster membership
cluster_membership_export = id_labels[,c('cluster','annotation_abrv','clusterAHN')]
cluster_membership_export$cluster = factor(cluster_membership_export$cluster)
cluster_membership_export$cluster = recode_factor(
  cluster_membership_export$cluster,
  '3'='1',
  '1'='3'
)

#export every correlation,
all_corrs_export = bind_rows(
  'Ctrl' = cor.ctrl.orig.long, 
  'NS' = cor.nt.orig.long, 
  'S' = cor.t.orig.long,
  .id = "group"
)
all_corrs_export = all_corrs_export[,c('group','Reg1','Reg2','R','p')]

#export cluster selection
wss_export = wss
names(wss_export) = c('# of clusters', 'within-cluster variance')

if (export){
  write.csv(ahn_corrs_export, 'Fig1e_AHNcorrs.csv', row.names=FALSE)
  write.csv(all_corrs_export, 'Fig1f_Allcorrs.csv', row.names=FALSE)
  write.csv(cluster_membership_export, 'Fig1f_ClusterMembership.csv', row.names=FALSE)
  write.csv(wss_export, 'ExtFig2d_ClusterSelection.csv', row.names=FALSE)
}

```