---
title: "SEFL17_CountAnalysis"
author: "Zach"
output:
  html_document: 
    fig_height: 4.5
    fig_width: 4.5
    highlight: kate
    number_sections: yes
    theme: readable
    toc: yes
    toc_depth: 6
editor_options: 
  chunk_output_type: console
---


# Set Global Options
```{r global_options}

knitr::opts_chunk$set(fig.width=4.5, fig.height=4.5, echo=TRUE, warning=FALSE,message=FALSE)

```


# Load Packages 
```{r}

library(ggplot2)
library(dplyr)
library(stats)
library(MASS)
library(multcomp)
library(ggh4x)

```


# Import Data 
```{r}

# load files
counts = read.csv('countsummary_cleaned.csv',header=TRUE)
group = read.csv('../SEFL17_Groups.csv', header=TRUE)
behav = read.csv('../Behavior/SEFL17_MildStressor.csv', header=TRUE)
labels = read.csv('allgraylabels.csv', header=TRUE)

# factorize as necessary
counts$group = factor(counts$group,levels=c("Control","No Trauma","Trauma"))
counts$graph_order = factor(counts$graph_order)
counts$annotation_abrv = factor(counts$annotation_abrv)

# get summary of behavior data for first 3 minutes after auditory stressor
behav = left_join(group, behav, by="Animal")
poststl = behav[behav$phase=='min' & behav$subphase=='post' & behav$t<3,]
poststl = summarise_at(
  .tbl = group_by(poststl, Animal, group),
  .funs = list(~mean(.)),
  .vars = c('Motion','Freezing')
)
poststl.stats = summarise_at(
  .tbl = group_by(poststl, group),
  .funs = list(mean=~mean(.), sem=~sd(.)/sqrt(n())),
  .vars = c('Motion','Freezing')
)

```


# Drop regions
Cerebellum excluded because several had cut in middle during extraction
```{r}

#drop cerebellum
counts = counts[counts$parcellation_group!='cerebellum',]
counts$annotation_abrv = droplevels(counts$annotation_abrv)

```



# Stats

## fit GLM and contrasts
```{r}

#Note that MLE failed to converge for ASO and PVi
region_info = list()
for (region in levels(counts$annotation_abrv)){
  
  data = counts[counts$annotation_abrv==region,]
  
  glm=glm.nb(
    formula = count~group,
    data = data
  )
  pmod = (anova(
    glm, 
    glm.nb(formula =  count ~ 1, data =  data)
    ))$`Pr(Chi)`[[2]]
  
  contrast = glht(glm, linfct = matrix(c(0,-1,1),1))
  
  coef_p = coef(summary(glm))[,'Pr(>|z|)']
  coef_z = coef(summary(glm))[,'z value']
  contrast_p = summary(contrast)$test$pvalues[1]
  contrast_t = summary(contrast)$test$tstat
  
  region_info[region] = list(c(
    unname(coef_z['groupNo Trauma']),
    unname(coef_p['groupNo Trauma']),
    unname(coef_z['groupTrauma']),
    unname(coef_p['groupTrauma']),
    contrast_t,
    contrast_p,
    pmod
  ))
}

glm_results = data.frame(do.call(rbind,region_info))
names(glm_results) = c(
  'ControlVnoTrauma_z',
  'ControlVnoTrauma_p',
  'ControlVTrauma_z',
  'ControlVTrauma_p',
  'TraumaVnoTrauma_t',
  'TraumaVnoTrauma_p',
  'Model_p'
)
glm_results = cbind(annotation_abrv = rownames(glm_results), glm_results)

```

## calculate fdr and write to file
```{r}

#assess overall effect of group
glm_results$Model_padj = p.adjust(glm_results$Model_p, 'fdr')
glm_results$Model_fdr = glm_results$Model_padj < .05

#assess comparison of NT and T groups
glm_results$NTvsT_padj = NA
glm_results$NTvsT_padj[glm_results$Model_fdr] = p.adjust(glm_results$TraumaVnoTrauma_p[glm_results$Model_fdr], 'fdr')
glm_results$NTvsT_fdr = glm_results$NTvsT_padj < .05

#caclulate count statistics
counts.stats = summarise_at(
  .tbl = group_by(
    .data = counts,
    annotation_abrv,group),
  .funs = list(mean=~mean(.), sem=~sd(.)/sqrt(n()), sd=~sd(.), var=~var(.)),
  .vars = c('count')
)
counts.stats = right_join(
  labels[,c('annotation_abrv','annotation_names')],
  counts.stats, 
  by = 'annotation_abrv'
)

#add parcellation group back to results df
glm_results = right_join(
  distinct(counts[c(
  'annotation_abrv','annotation_names','parcellation_group','graph_order'
  )]), 
  glm_results,
  by='annotation_abrv'
)

#write.csv(glm_results, 'STable1.csv', row.names=FALSE)
#write.csv(counts.stats, 'STable2.csv', row.names=FALSE)

```




# Plots
## Set Global Plot Parameters
```{r}

p.save = FALSE

```

## Plot T distribution
### Format data for plotting
```{r}

# select regions that display differential activation by group
pdata = glm_results[glm_results$Model_fdr,]
pdata$graph_order = droplevels(pdata$graph_order)
pdata$parcellation_group = factor(
  pdata$parcellation_group, 
  levels = c('cortical_plate','cortical_subplate','cerebral_nuclei','diencephalon','midbrain','hindbrain')
)

# add 'significance groups'
pdata$NTvsT_LiberalSig = pdata$TraumaVnoTrauma_p<.05
pdata$siggroup = 'NC' #initialize with all being 'no change'
pdata$siggroup[pdata$NTvsT_LiberalSig & pdata$TraumaVnoTrauma_t>0] = 'lib_up'
pdata$siggroup[pdata$NTvsT_LiberalSig & pdata$TraumaVnoTrauma_t<0] = 'lib_down'
pdata$siggroup[pdata$Model_fdr & pdata$NTvsT_fdr & pdata$TraumaVnoTrauma_t>0] = 'fdr_up'
pdata$siggroup[pdata$Model_fdr & pdata$NTvsT_fdr & pdata$TraumaVnoTrauma_t<0] = 'fdr_down'
pdata$siggroup = factor(pdata$siggroup, levels=c('NC','lib_down','fdr_down','lib_up','fdr_up'))

# create numeric numeric region ID to allow representation of region as continuous variable
pdata = pdata[order(pdata$parcellation_group, pdata$graph_order),]
pdata$graph_num = 1:nrow(pdata)

#define bounds of parcellation regions
parc_bounds = c()
parc_mids = c()
for (group in levels(pdata$parcellation_group)) {
  parc_mids = c(
    parc_mids, mean(pdata$graph_num[pdata$parcellation_group==group])
  )
  parc_bounds = c(
    parc_bounds, max(pdata$graph_num[pdata$parcellation_group==group]) + .5
  )
}

#define parcellation labels for graph
parc_labels = c('Cortex','Cortical Subplate','Cerebral Nuclei','Diencephalon','Midbrain','Hindbrain')

```


### Plot T values
```{r}

p.t = ggplot(
  data = pdata,
  aes(x=graph_num, y=TraumaVnoTrauma_t, fill=siggroup)
)

p.t + 
  
  geom_bar(stat='identity', show.legend = FALSE) +
  
  geom_vline(
   xintercept = parc_bounds[1:length(parc_bounds)-1],
   linetype = 'dotted',
   color = 'black',
   linewidth = .5,
   alpha = 1
  ) +
  
  scale_y_continuous(limits = c(-4,4)) +
  scale_x_continuous(breaks = parc_mids, labels = parc_labels) +
  
  coord_flip(expand = FALSE) +

  labs(
    title = 'T vs. NT cFos',
    x = element_blank(),
    y = 't (T-NT)'
  ) +
  
  scale_fill_manual(values = c('gray80','steelblue4','blue3','red3','red3')) +
  
  theme_classic() +
  theme(
    plot.title = element_text(
      hjust = .5, 
      size = 20,
      face = 'bold'),
    axis.title = element_text(
      color = 'black',
      face =  'bold',
      size = 12),
    axis.text.y = element_text(
      angle = 0,
      color = 'black',
      face = 'bold',
      size = 11),
    axis.text.x = element_text(
      color = 'black',
      size = 10,
      angle= 0,
      hjust = .5, 
      vjust=1),
    axis.ticks.x = element_blank()
  )

# if (p.save == TRUE) {
#   ggsave(
#     filename="SEFL17_Pvalues_singlecolor.tiff",
#     width=3.25, height=7, units="in",
#     dpi=300,
#     bg='transparent')
# }

```


## Plot means of selected regions
```{r}


regions = c('AHN', 'LA', 'BLAp', 'BMA', 'PL2/3', 'ILA2/3', 'LHA','VMH','MEA','PAG')
#'LA', 'BLAp', 'BMA', 'PL2/3', 'ILA2/3', 'LHA'

for (region in regions) {

  mx = max(counts$count[counts$annotation_abrv==region])*1.1
  mxbrk = round(mx, 1-nchar(as.character(round(mx))))
  brks = seq(0, mxbrk, mxbrk/4)
  
  p.region = ggplot(
    data = counts.stats[counts.stats$annotation_abrv==region,],
    aes(x=group, y=mean, fill=group)
  )
  
  p.region +
    
    geom_crossbar(
      stat="identity",
      linewidth = .75,
      colour = "black",
      aes(ymin = mean - sem,
          ymax = mean + sem),
      width = .7,
      show.legend = FALSE) +
    
    geom_jitter(
      data = counts[counts$annotation_abrv==region,],
      aes(x=group, y=count),
      fill = NA,
      shape =  21,
      size = 2.5,
      alpha = .4,
      width = .2
    ) +
    
    scale_fill_manual(values=c('white','yellow1','orange')) +
    coord_cartesian(ylim=c(0, mxbrk)) +
    scale_y_continuous(expand = c(0, 0), breaks = brks) +
    scale_x_discrete(labels = c('Ctrl','NS','S')) +
    
    labs(
      title = region,
      y = "Mean cFos Counts",
      x = ""
    ) +
    
    theme_bw() + 
    theme(
      panel.grid = element_blank(), 
      panel.border = element_rect(colour = "white"),
  
      axis.line = element_line(color="black", linewidth = .5),
      plot.title = element_text(
        size = 24, 
        hjust= .5, margin=margin(b=15,t=10), 
        face = "bold"),
      axis.title = element_text(size=16, face="bold"),
      axis.text= element_text(size=12, colour="black"),
      axis.text.x= element_text(
        size = 16,
        face='bold',
        angle = 0,
        hjust = .5,
        vjust=.5,
        margin=margin(b=10,t=5)),
      plot.background = element_rect(fill = "transparent",colour = NA),
      plot.margin = unit(c(2, 3.75, 2,3.75), "cm")
    ) 
  
  if (p.save) {
    ggsave(
      filename = paste('SEFL17_MeanCounts_', gsub('/','',region), ".svg", sep = ''),
      width = 5, height = 5, units = "in", dpi = 500, bg = 'transparent')
    }

}


```


## AHN-Behav Corr
```{r}

counts$Animal = sapply(
  counts$animal,
  function(x){as.numeric(strsplit(x,'_')[[1]][2])}
)

ahn_fz_cordata = left_join(
  poststl,
  counts[counts$annotation_abrv=='AHN',],
  by = join_by(Animal)
)
ahn_fz_cordata = ahn_fz_cordata[ahn_fz_cordata$group.x!='No Stim',]
ahn_fz_cordata$group.x = recode_factor(
  ahn_fz_cordata$group.x,
  'No Trauma' = 'NS',
  'Trauma' = 'S'
)

#predict freezing from group
ahn_fz_lm.g = lm(Freezing ~ group.x, data=ahn_fz_cordata)
anova(ahn_fz_lm.g)
summary(ahn_fz_lm.g)

#predict freezing from counts
ahn_fz_lm.c = lm(Freezing ~ count, data=ahn_fz_cordata)
anova(ahn_fz_lm.c)
summary(ahn_fz_lm.c)

#predict freezing from counts and group
ahn_fz_lm.null = lm(Freezing ~ 1, data=ahn_fz_cordata)
anova(ahn_fz_lm.null)
summary(ahn_fz_lm.null)


```

### Plot
```{r}

p.ahn_fz_cor = ggplot(
  data = ahn_fz_cordata, 
  aes(x=count, y=Freezing, fill=group.x, color=group.x)
)

#Define plot specific parameters
p.ymin = 0
p.ymax = 100
p.yint = 25
p.ybreaks = seq(p.ymin,p.ymax,p.yint)

p.xmin = 1000
p.xmax = 3200
p.xint = 1000
p.xbreaks = seq(p.xmin,p.xmax,p.xint)

p.ahn_fz_cor +
  
  # geom_abline(
  #   intercept = ahn_fz_lm.c$coefficients[[1]],
  #   slope = ahn_fz_lm.c$coefficients[[2]],
  #   linetype = 'dashed'
  # ) +
  
  geom_smooth(
    method = "lm", 
    se = FALSE,
    linetype = 'solid',
    linewidth = 1,
    fullrange = FALSE
  ) +
  
  geom_point(
    size = 3,
    shape = 21,
    color = 'black'
  ) +
  
  labs(
    x = 'AHN cFos Counts',
    y = '% Freezing'
  ) +
  
  scale_y_continuous(breaks=p.ybreaks, limits=c(p.ymin,p.ymax), expand=c(0,0)) +
  scale_x_continuous(breaks=p.xbreaks, limits=c(p.xmin,p.xmax), expand=c(.02,.02)) +
  
  scale_fill_manual(values=c('yellow1','orange')) +
  scale_color_manual(values=c('yellow3','orange')) +
  
  theme_classic() +
  theme(
    axis.text = element_text(size=12),
    axis.title = element_text(size=12, face='bold'),
    legend.title = element_blank(),
    legend.text = element_text(size=12, face='bold'),
    legend.position = 'bottom'
  ) +
  
  force_panelsizes(
    cols = unit(1.5, "in"),
    total_height = unit(1.5, "in")
  ) 

if (p.save) {
  ggsave(
    filename = paste('SEFL17_Corr_AHNbehav', ".svg", sep = ''),
    width = 4, height = 4, units = "in", dpi = 500, bg = 'transparent'
  )
}

#if (p$save == TRUE) {saveplot(filename = paste(f_pfx,'_',"Actxtest", p$savetype, sep = ''))}


```


# Export figure data for publication
```{r}

export = FALSE

#export count data
count_export = counts[,c('Animal','sex','group','annotation_abrv','annotation_names','count')]
count_export$group = recode_factor(
  count_export$group,
  'Control' = 'Ctrl',
  'No Trauma' = 'NS',
  'Trauma' = 'S'
)

ahn_cordata_export = ahn_fz_cordata[
  ,c('Animal','sex','group.x','annotation_abrv','annotation_names','count','Freezing')
]
names(ahn_cordata_export)[names(ahn_cordata_export)=='group.x'] = 'group'

if (export){
  write.csv(count_export, 'Fig1d_counts.csv', row.names=FALSE)
  write.csv(ahn_cordata_export, 'ExtFig1j_ahncountsANDfreezing.csv', row.names=FALSE)
}

```




