---
title: "SEFL21H2_Analysis"
author: "Zach"
output:
  html_document: 
    fig_height: 4.5
    fig_width: 4.5
    highlight: kate
    number_sections: yes
    theme: journal
    toc: yes
    toc_depth: 6
editor_options: 
  chunk_output_type: console
---


# Set Global Options
```{r global_options}
knitr::opts_chunk$set(fig.width=4.5, fig.height=4.5, echo=TRUE, warning=FALSE,message=FALSE)
```



# Load Packages 
```{r}

library(ggplot2)
library(ez)
library(knitr)
library(tidyr)
library(car)
library(dplyr)
library(purrr)
library(lmerTest)
library(ggh4x)

```



# Import Data 
Load  files and factorize as necessary.\
```{r}

#load  files
data_fz = read.csv('SEFL21H2_fzdata.csv',header=TRUE)
data_ld = read.csv('SEFL21H2_lddata.csv',header=TRUE)
groups = read.csv('SEFL21H2_groups.csv',header=TRUE)

#factorize as necessary
groups$Animal = factor(groups$Animal)
data_fz$Animal = factor(data_fz$Animal)
data_ld$Animal = factor(data_ld$Animal)
groups$Virus = recode_factor(groups$Virus, 'mCherry'='mCherry','stGtACR1'='GtACR')

# #add group info to data files
data_fz = inner_join(x=groups,y=data_fz,by="Animal")
data_ld = inner_join(x=groups,y=data_ld,by="Animal")

```



# Drop subjects
See `groups` dataframe for exclusion rationale.
```{r}

data_fz = data_fz[data_fz$Exclude!=1,]
data_ld = data_ld[data_ld$Exclude!=1,]

```


# Subset data
Separate phases (trauma, trauma recall, etc) are placed in unique dataframes for ease of analysis.
```{r}

#define bins for pre trauma opto
data_ld$lightbin = data_ld$bin %/% 2
data_ld$lightpair = ((data_ld$lightbin) %/% 2) + 1 
data_ld$laser = data_ld$lightbin %% 2 == 1 #no light=False, light=True

#collapse further into light off/on periods
ld.collapsed = summarise_at(
  .tbl = group_by(
    .data = data_ld,
    Animal, Sex, laser, Virus),
  .funs = list(mean),
  .vars = c('dark','Distance_in')
)

#trauma data
trauma.freezing = data_fz[data_fz$phase=='trauma' & data_fz$subphase!="shock",]
trauma.freezing$component=factor(
  trauma.freezing$component, 
  levels =  c(
    "bl",
    "laser",
    "postshock_1",
    "postshock_2",
    "postshock_3",
    "postshock_4",
    "postshock_5",
    "postshock_6",
    "postshock_7",
    "postshock_8",
    "postshock_9",
    "postshock_10"
  )
)


#trauma shock reactivity
trauma.shockreact = summarise_at(
  .tbl = group_by(
    .data = data_fz[data_fz$phase=='trauma' & data_fz$subphase!="postshock" & data_fz$component != 'laser',],
    Animal, subphase, Virus),
  .funs = list(motion=mean),
  .vars = c('motion')
)


#trauma recall data
traumatest = data_fz[data_fz$phase=='traumatest' & data_fz$component=='avg',]


#nov stressor data
##prepost startle stimulus
novstressor.prepost = data_fz[data_fz$phase=='mildstressor' & data_fz$subphase=='prepost',]
novstressor.prepost$component = factor(novstressor.prepost$component, levels=c('bl','post'))
##startle response
novstressor.startle = data_fz[data_fz$phase=='mildstressor' & data_fz$subphase=='startle',]
novstressor.startle$component = factor(
  novstressor.startle$component,
  levels=c('startle_pre','startle_startle','startle_post'))


#novel stressor (sefl) test data
sefltest = data_fz[data_fz$phase=='sefltest' & data_fz$subphase=='avg',]


```


# Get summary statistics for plotting
Calculate relevant means and std. errors.
```{r}

#ld stats
ld.stats = summarise_at(
  .tbl = group_by(
    .data = ld.collapsed,
    Virus, laser),
  .funs = list(mean=~mean(.), sem=~sd(.)/sqrt(n())),
  .vars = c('dark','Distance_in'))


#trauma session freezing
trauma.freezing.stats = summarise_at(
  .tbl = group_by(
    .data = trauma.freezing,
    component, Virus),
  .funs = list(mean=~mean(.), sem=~sd(.)/sqrt(n())),
  .vars = c('motion','freezing')
)
trauma.freezing.stats$subphase = 'postshock'
trauma.freezing.stats$subphase[trauma.freezing.stats$component=='bl' | trauma.freezing.stats$component=='laser'] = 'bl'

#trauma shock reactivity
trauma.shockreact.stats = summarise_at(
  .tbl = group_by(
    .data = trauma.shockreact,
    subphase, Virus),
  .funs = list(mean=~mean(.), sem=~sd(.)/sqrt(n())),
  .vars = c('motion')
)


#trauma recall
traumatest.stats = summarise_at(
  .tbl = group_by(
    .data = traumatest,
    Virus),
  .funs = list(mean=~mean(.), sem=~sd(.)/sqrt(n())),
  .vars = c('freezing','motion')
)


#novel stressor
#novel stressor pre/post
novstressor.prepost.stats = summarise_at(
  .tbl = group_by(
    .data = novstressor.prepost,
    component, Virus),
  .funs = list(mean=~mean(.), sem=~sd(.)/sqrt(n())),
  .vars = c('freezing','motion')
)

#novel stressor startle
novstressor.startle.stats = summarise_at(
  .tbl = group_by(
    .data = novstressor.startle,
    component, Virus),
  .funs = list(mean=~mean(.), sem=~sd(.)/sqrt(n())),
  .vars = c('freezing','motion')
)


#sefl test
sefltest.stats = summarise_at(
  .tbl = group_by(
    .data = sefltest,
    Virus),
  .funs = list(mean=~mean(.), sem=~sd(.)/sqrt(n())),
  .vars = c('freezing','motion')
)

```



# Plots
## Global Plot Parameters
```{r}

#set figure file prefix
f_pfx = '21H2'

p = list()

#ggsave
p$save = FALSE
p$savetype = '.svg'

##Plot dimensions
p$dim = list(
  width = 4.5, #Plot width (in)
  height = 4.5, #Plot height (in)
  panel_w = unit(1, 'in'),
  panel_h = unit(2, 'in'),
  dpi = 500 #Plot dpi
)

#Plot text
p$textsize = list(
  title = 20,
  legend = 14,
  axis_title = 16,
  axis_text = 14
)

#Plot line dimensions
p$linesize = list(
  data_grp = 1,
  data_ind = .6,
  axis = .5,
  bar = .7,
  error =.75
)
p$linealpha =list(
  data_ind = .2
)

#Plot Widths
p$width = list(
  bar = .8,
  jitter = .3,
  error = .3
)

#Point/dot sizes
p$points = list(
  grp = list(
    shape = 21,
    size = 4,
    stroke = 1,
    alpha = 1
  ),
  ind = list(
    shape =  21,
    size = 2.5,
    alpha = .4
  )
)

#theme
p$theme = list(
  base = theme_bw(),
  specs = theme(
    panel.grid = element_blank(), 
    panel.border = element_rect(colour="white"),
    axis.line = element_line(
      linewidth=p$linesize$axis,
      color="black"
    ),
    plot.title = element_text(
      size=p$textsize$title, 
      hjust=.5, 
      margin=margin(b=15,t=10), 
      face="bold"
    ),
    axis.title = element_text(
      size=p$textsize$axis_title, 
      face="plain"
    ),
    axis.text=element_text(
      size=p$textsize$axis_text, 
      colour="black"
    ),
    legend.position = c(.3,.85),
    legend.title = element_blank(),
    legend.text = element_text(
      size=p$textsize$legend, 
      face="plain"
    ),
    legend.key = element_rect(colour=NA, fill=NA),
    legend.background = element_rect(fill=NA),
    plot.background = element_rect(fill = "transparent", colour = NA),
  ) 
)

#Plot colors
p$cols = list(
  color = c("white","blue4"),
  fill = c("white","blue4"),
  rec = 'slategray1'
)

#predefine partial function for saving plots
saveplot = partial(
  ggsave,
  width = p$dim$width,
  height = p$dim$height,
  units = "in", 
  dpi = p$dim$dpi,
  bg = 'transparent'
)

```



## Make Plots

### Light Dark, Pre-trauma
#### Dark
```{r, fig.width=4.5, fig.height=4.5}

#Create base layer from group level data
P.LDpre.dark = ggplot(
  data = ld.stats,
  aes(x = laser,
      y = dark_mean,
      fill = Virus,
      group = Virus))

#Define plot specific parameters
p.ymin = .4
p.ymax = 1
p.yint = .2
p.breaks = seq(p.ymin,p.ymax,p.yint)

#Create plot
P.LDpre.dark  +

  geom_tile(
    aes(x = TRUE, y=0, width=1, height = Inf),
    fill = p$cols$rec
  ) +

  geom_hline(yintercept=.5, linetype='dashed') +
  
  geom_errorbar(
    stat = "identity",
    linewidth = p$linesize$error,
    colour = "black",
    aes(ymin = dark_mean - dark_sem,
        ymax = dark_mean + dark_sem,
        width = p$width$error
        )) +
  
  geom_line(
    stat = "identity",
    linewidth = p$linesize$data_grp
  ) +
  
  geom_point(
    stat = "identity",
    size = p$points$grp$size,
    shape = p$points$grp$shape,
    stroke = p$points$grp$stroke,
    show.legend = FALSE
  ) +
  
  scale_fill_manual(values=p$cols$fill) +
  coord_cartesian(ylim=c(p.ymin,p.ymax)) +
  scale_y_continuous(breaks=p.breaks, labels = p.breaks*100, expand = c(0, 0)) +
  scale_x_discrete(labels=c('off','on')) +
  
  labs(
    x = 'Laser',
    y = "% Time in Dark"
  ) +
  
  p$theme$base +
  p$theme$specs +
  
  force_panelsizes(
      rows = p$dim$panel_h,
      cols = p$dim$panel_w
  ) 


if (p$save == TRUE) {
  saveplot(filename = paste(f_pfx,'_',"LDpre_Dark", p$savetype, sep = ''))
}

```



#### Distance 
```{r, fig.width=4.5, fig.height=4.5}


#Create base layer from group level data
P.LDpre.dist = ggplot(
  data = ld.stats,
  aes(x = laser,
      y = Distance_in_mean,
      fill = Virus,
      group = Virus))

#Define plot specific parameters
p.ymin = 0
p.ymax = 200
p.yint = 50
p.breaks = seq(p.ymin,p.ymax,p.yint)

#Create plot
P.LDpre.dist  +

  geom_tile(
    aes(x = TRUE, y=0, width=1, height = Inf),
    fill = p$cols$rec
  ) +
  
  geom_errorbar(
    stat = "identity",
    linewidth = p$linesize$error,
    colour = "black",
    aes(ymin = Distance_in_mean - Distance_in_sem,
        ymax = Distance_in_mean + Distance_in_sem,
        width = p$width$error
        )) +
  
  geom_line(
    stat = "identity",
    linewidth = p$linesize$data_grp
  ) +
  
  geom_point(
    stat = "identity",
    size = p$points$grp$size,
    shape = p$points$grp$shape,
    stroke = p$points$grp$stroke,
    show.legend = FALSE
  ) +
  
  scale_fill_manual(values=p$cols$fill) +
  coord_cartesian(ylim=c(p.ymin,p.ymax)) +
  scale_y_continuous(breaks=p.breaks, labels = p.breaks, expand = c(0, 0)) +
  scale_x_discrete(labels=c('off','on')) +
  
  labs(
    x = 'Laser',
    y = "Distance (in)"
  ) +
  
  p$theme$base +
  p$theme$specs +
  
  force_panelsizes(
      rows = p$dim$panel_h,
      cols = p$dim$panel_w
  )

if (p$save == TRUE) {
  saveplot(filename = paste(f_pfx,'_',"LDpre_Dist", p$savetype, sep = ''))
}

```


### Trauma
#### Freezing
```{r, fig.width=4.5, fig.height=4.5}

#Create base layer from group level data
P.Trauma.Freezing = ggplot(
  data = trauma.freezing.stats[trauma.freezing.stats$component!='laser',],
  aes(x = component,
      y = freezing_mean,
      fill = Virus,
      shape = Virus,
      group = interaction(subphase,Virus)))

#Define plot specific parameters
p.ymin = 0
p.ymax = 100
p.yint = 25
p.breaks = seq(p.ymin,p.ymax,p.yint)

#Create plot
P.Trauma.Freezing  +
  
  geom_rect(
    aes(
      xmin = 1.5,
      xmax = Inf,
      ymin = 0,
      ymax = Inf
    ),
    fill = p$cols$rec
  ) +
  
  geom_errorbar(
    stat = "identity",
    linewidth = p$linesize$error,
    colour = "black",
    aes(ymin = freezing_mean - freezing_sem,
        ymax = freezing_mean + freezing_sem,
        width = p$width$error
        )) +
  
  geom_line(
    stat = "identity",
    linewidth = p$linesize$data_grp
  ) +
  
  geom_point(
    stat = "identity",
    size = p$points$grp$size,
    stroke = p$points$grp$stroke,
    shape = p$points$grp$shape,
    show.legend = TRUE
  ) +
  
  scale_fill_manual(values=p$cols$fill) +
  coord_cartesian(ylim=c(p.ymin,p.ymax)) +
  scale_y_continuous(breaks=p.breaks, expand = c(0, 0)) +
  scale_x_discrete(
    labels=c('P','1','2','3','4','5','6','7','8','9','10'),
    expand = c(.1,.1)
  ) +
  
  labs(
    x = "Shock",
    y = "% Freezing"
  ) +
  
  p$theme$base +
  p$theme$specs +
  
  theme(
    axis.text.x = element_text(size=p$textsize$axis_text-4),
    #axis.title.x = element_text(margin=margin(t=-5)),
    legend.title = element_blank(),
    legend.position = c(.45, 1.02),
    legend.box.margin = margin(t=-0)
  ) +
  
  force_panelsizes(
      rows = p$dim$panel_h,
      cols = unit(1.35, "in")
  ) 

if (p$save == TRUE) {
  saveplot(filename = paste(f_pfx,'_',"Trauma", p$savetype, sep = ''))
}

```



#### Shock Reactivity
```{r, fig.width=4.5, fig.height=4.5}

#Create base layer from group level data
P.Trauma.Shock = ggplot(
  data = trauma.shockreact.stats,
  aes(x = subphase,
      y = mean,
      fill = Virus,
      group = Virus)) 

#Define plot specific parameters
p.ymin = 0
p.ymax = 3000
p.yint = 1000
p.breaks = seq(p.ymin,p.ymax,p.yint)

#Create plot
P.Trauma.Shock  +
  
  geom_errorbar(
    stat = "identity",
    linewidth = p$linesize$error,
    colour = "black",
    aes(ymin = mean - sem,
        ymax = mean + sem,
        width = p$width$error)) +
  
  geom_line(
    stat="identity",
    linewidth = p$linesize$data_grp
  ) +
  
  geom_point(
    stat = "identity",
    size = p$points$grp$size,
    shape = p$points$grp$shape,
    stroke = p$points$grp$stroke,
    show.legend = FALSE
  ) +
  
  scale_fill_manual(values=p$cols$fill) +
  coord_cartesian(ylim=c(p.ymin,p.ymax)) +
  scale_y_continuous(breaks=p.breaks, expand = c(0, 0)) +
  scale_x_discrete(labels=c('bl','shock')) +
  
  labs(
    x = "",
    y = "Motion (au)"
  ) +
  
  p$theme$base +
  p$theme$specs +

  force_panelsizes(
      rows = p$dim$panel_h,
      cols = p$dim$panel_w
  ) 

# if (p$save == TRUE) {
#   saveplot(filename = paste(f_pfx,'_',"Trauma_ShockReact", p$savetype, sep = ''))
# }

```


### Trauma Recall
```{r, fig.width=4.5, fig.height=4.5}

#Create base layer from group level data
P.TR = ggplot(
  data = traumatest.stats,
  aes(x = Virus,
      y = freezing_mean,
      fill = Virus))

#Define plot specific parameters
p.ymin = 0
p.ymax = 100
p.yint = 25
p.breaks = seq(p.ymin,p.ymax,p.yint)

#Create plot
P.TR  +

  geom_crossbar(
    stat = "identity",
    width = p$width$bar,
    linewidth = p$linesize$error,
    colour = "black",
    aes(ymin = freezing_mean - freezing_sem,
        ymax = freezing_mean + freezing_sem,
        ),
    show.legend = FALSE
    ) +

  geom_jitter(
      data = traumatest,
      aes(x = Virus,
          y = freezing),
      width = p$width$bar * p$width$jitter,
      height = 0,
      shape = p$points$ind$shape,
      size = p$points$ind$size,
      colour = 'black',
      fill = NA,
      alpha = p$points$ind$alpha,
      show.legend=FALSE
  ) +
  
  scale_fill_manual(values=p$cols$fill) +
  coord_cartesian(ylim=c(p.ymin,p.ymax)) +
  scale_y_continuous(breaks=p.breaks, labels = p.breaks, expand = c(0, 0)) +
  
  labs(
    x = '',
    y = "% Freezing"
  ) +
  
  p$theme$base +
  p$theme$specs +
  theme(
    legend.position = c(1.5,.2),
    axis.text.x = element_text(angle=-45, hjust=0)
  ) +

  force_panelsizes(
      rows = p$dim$panel_h,
      cols = p$dim$panel_w
  ) 

if (p$save == TRUE) {
  saveplot(filename = paste(f_pfx,'_',"TraumaRecall1", p$savetype, sep = ''))
}


```



### Novel Stressor
#### Pre-Post
```{r, fig.width=4.5, fig.height=4.5}

#Create base layer from group level data
P.NovSts.prepost = ggplot(
  data = novstressor.prepost.stats,
  aes(x = component,
      y = freezing_mean,
      fill = Virus,
      group = Virus))

#Define plot specific parameters
p.ymin = 0
p.ymax = 50
p.yint = 10
p.breaks = seq(p.ymin,p.ymax,p.yint)

#Create plot
P.NovSts.prepost  +
  
  geom_errorbar(
    stat = "identity",
    linewidth = p$linesize$error,
    colour = "black",
    aes(ymin = freezing_mean - freezing_sem,
        ymax = freezing_mean + freezing_sem,
        width = p$width$error
        )) +
  
  geom_line(
    stat="identity",
    linewidth = p$linesize$data_grp
  ) +
  
  geom_point(
    stat = "identity",
    size = p$points$grp$size,
    shape = p$points$grp$shape,
    stroke = p$points$grp$stroke,
    show.legend = TRUE
  ) +
  
  scale_fill_manual(values=p$cols$fill) +
  coord_cartesian(ylim=c(p.ymin,p.ymax)) +
  scale_y_continuous(breaks=p.breaks, labels = p.breaks, expand = c(0, 0)) +
  scale_x_discrete(labels=c('pre','post')) +
  
  labs(
    x = '',
    y = "% Freezing"
  ) +
  
  p$theme$base +
  p$theme$specs +
  
  theme(
    legend.title = element_blank(),
    legend.position = 'right',
    axis.text.x = element_text(angle=-45, hjust=0, vjust=1),
  ) +
  
  force_panelsizes(
      rows = p$dim$panel_h,
      cols = p$dim$panel_w
  ) 

if (p$save == TRUE) {
  saveplot(filename = paste(f_pfx,'_',"NovSts_PrePost", p$savetype, sep = ''))
}

```


#### Startle
```{r, fig.width=4.5, fig.height=4.5}

#Create base layer from group level data
P.NovSts.startle = ggplot(
  data = novstressor.startle.stats,
  aes(x = component,
      y = motion_mean,
      fill = Virus,
      group = Virus))

#Define plot specific parameters
p.ymin = 0
p.ymax = 2500
p.yint = 500
p.breaks = seq(p.ymin,p.ymax,p.yint)

#Create plot
P.NovSts.startle  +
  
  geom_errorbar(
    stat = "identity",
    linewidth = p$linesize$error,
    colour = "black",
    aes(ymin = motion_mean - motion_sem,
        ymax = motion_mean + motion_sem,
        width = p$width$error
        )) +
  
  geom_line(
    stat="identity",
    linewidth = p$linesize$data_grp
  ) +
  
  geom_point(
    stat = "identity",
    size = p$points$grp$size,
    shape = p$points$grp$shape,
    stroke = p$points$grp$stroke,
    show.legend = FALSE
  ) +
  
  scale_fill_manual(values=p$cols$fill) +
  coord_cartesian(ylim=c(p.ymin,p.ymax)) +
  scale_x_discrete(labels=c('Pre','Noise','Post')) +
  scale_y_continuous(breaks=p.breaks, labels = p.breaks, expand = c(0, 0)) +
  
  labs(
    x = '',
    y = "Motion (au)"
  ) +
  
  p$theme$base +
  p$theme$specs +
  
  theme(
    axis.text.x= element_text(angle=-90, hjust=0),
  ) +
  
  force_panelsizes(
      rows = p$dim$panel_h,
      cols = p$dim$panel_w
  ) 

# if (p$save == TRUE) {
#   saveplot(filename = paste(f_pfx,'_',"NovSts_Star

```



### SEFL Test
```{r, fig.width=4.5, fig.height=4.5}

#Create base layer from group level data
P.sefltest = ggplot(
  data = sefltest.stats,
  aes(x = Virus,
      y = freezing_mean,
      fill = Virus)) 

#Define plot specific parameters
p.ymin = 0
p.ymax = 50
p.yint = 10
p.breaks = seq(p.ymin,p.ymax,p.yint)

#Create plot
P.sefltest  +
  
  geom_crossbar(
    stat = "identity",
    position = position_dodge(),
    linewidth = p$linesize$error,
    colour = "black",
    aes(ymin = freezing_mean - freezing_sem,
        ymax = freezing_mean + freezing_sem),
    width = p$width$bar,
    show.legend = FALSE) +
  
  geom_jitter(
      data = sefltest,
      aes(x = Virus,
          y = freezing),
      width = p$width$bar * p$width$jitter,
      height = 0,
      shape = p$points$ind$shape,
      size = p$points$ind$size,
      colour = 'black',
      fill = NA,
      alpha = p$points$ind$alpha,
      show.legend=FALSE
  ) +
  
  scale_fill_manual(values=p$cols$fill) +
  coord_cartesian(ylim=c(p.ymin,p.ymax)) +
  scale_y_continuous(breaks=p.breaks, labels = p.breaks, expand = c(0, 0)) +
  
  labs(
    x = "",
    y = "% Freezing"
  ) +
  
  p$theme$base +
  p$theme$specs +
  
  theme(
    legend.position = c(1,1),
    axis.text.x = element_text(angle = -45, hjust=0)
  ) +
  
  force_panelsizes(
      rows = p$dim$panel_h,
      cols = p$dim$panel_w, #unit(1.75, 'in')
  ) 

if (p$save == TRUE) {
  saveplot(filename = paste(f_pfx,'_',"SEFLtest", p$savetype, sep = ''))
}


```




## Stats
### trauma
```{r}

trauma.anova = ezANOVA(
  data = trauma.freezing[trauma.freezing$subphase!='bl',],
  wid = Animal,
  dv = freezing,
  within = component,
  between = Virus,
  type = 3,
  white.adjust = TRUE
)
trauma.anova

trauma.bl.anova = ezANOVA(
  data = trauma.freezing[trauma.freezing$component=='bl',],
  wid = Animal,
  dv = freezing,
  between = Virus,
  type = 3,
  white.adjust = TRUE
)
trauma.bl.anova

```

### trauma test
```{r}

# fit using anova
traumatest.anova = ezANOVA(
  data = traumatest,
  dv = freezing,
  wid = Animal,
  white.adjust = TRUE,
  between = Virus,
  type=3
)
traumatest.anova

```

### novel stressor
```{r}

novstressor.anova = ezANOVA(
  data = novstressor.prepost,
  dv = freezing,
  wid = Animal,
  white.adjust = TRUE,
  within = component,
  between = Virus,
  type=3
)
novstressor.anova

novstressor.bl.anova = ezANOVA(
  data = novstressor.prepost[novstressor.prepost$component=='bl',],
  dv = freezing,
  wid = Animal,
  white.adjust = TRUE,
  between = Virus,
  type=3
)
novstressor.bl.anova

novstressor.post.anova = ezANOVA(
  data = novstressor.prepost[novstressor.prepost$component=='post',],
  dv = freezing,
  wid = Animal,
  white.adjust = TRUE,
  between = Virus,
  type=3
)
novstressor.post.anova

```

### sefl test
```{r}

sefltest.anova = ezANOVA(
  data = sefltest,
  dv = freezing,
  wid = Animal,
  between = .(Virus),
  type = 3,
  white.adjust = TRUE
)
sefltest.anova

```

### light-dark test
```{r}

ld.pre.anova = ezANOVA(
  data = ld.collapsed,
  dv = dark,
  wid = Animal,
  between = .(Virus),
  within = laser,
  type = 3,
  white.adjust = TRUE
)
ld.pre.anova

```

# Export figure data for publication
```{r}

export = FALSE

#export S1 data
S1_export = trauma.freezing[,c('Animal','Sex','Virus','component','freezing')]
S1_export = S1_export[S1_export$component!='laser',]

#export S1 recall data
S1recall_export = traumatest[,c('Animal','Sex','Virus','freezing')]

#export S2 data
S2_export =  novstressor.prepost[,c('Animal','Sex','Virus','component','freezing')]
S2_export$component = recode_factor(S2_export$component, 'bl'='pre')

#export S2 recall data
S2recall_export = sefltest[,c('Animal','Sex','Virus','freezing')]

#export light-dark data
ld_export = ld.collapsed[,c('Animal','Sex','Virus','laser','dark','Distance_in')]
ld_export$dark = ld_export$dark*100

if (export){
  write.csv(S1_export, 'Fig3c.csv', row.names=FALSE)
  write.csv(S1recall_export, 'Fig3d.csv', row.names=FALSE)
  write.csv(S2_export, 'Fig3e.csv', row.names=FALSE)
  write.csv(S2recall_export, 'Fig3f.csv', row.names=FALSE)
  write.csv(ld_export, 'ExtFig4i.csv', row.names=FALSE)
}

```
















