---
title: "SEFL25c_combined_CaAnalysis"
author: "Zach"
output:
  html_document: 
    fig_height: 4.5
    fig_width: 4.5
    highlight: kate
    number_sections: yes
    theme: journal
    toc: yes
    toc_depth: 6
editor_options: 
  chunk_output_type: console
---


# Set Global Options
```{r global_options}

knitr::opts_chunk$set(fig.width=4.5, fig.height=4.5, echo=TRUE, warning=FALSE, message=FALSE)

```



# Load Packages 
```{r}

library(ggplot2)
library(svglite)
library(ez)
library(knitr)
library(tidyr)
library(dplyr)
library(purrr)
library(ggh4x)
library(stringr)
library(arrow)


```


# Import Data 
Load  files and factorize as necessary.
```{r}

#load animal data
groups = read.csv('SEFL_25c_25c2_25c5_AllAnimals.csv')

#load unit classifications and unit responses
rew_class = list()
shk_class = list()
sig = list()
for (f in dir('UnitClassification/Reward/')){
  rew_class[[f]] = read_feather(paste('UnitClassification/Reward/',f,sep=''))
}
for (f in dir('UnitClassification/Shock/')){
  shk_class[[f]] = read_feather(paste('UnitClassification/Shock/',f,sep=''))
}
for (f in dir('BinnedUnitResponses/')){
  sig[[f]] = read_feather(paste('BinnedUnitResponses/',f,sep=''))
}
rew_class = do.call('rbind', rew_class)
shk_class = do.call('rbind', shk_class)
sig = do.call('rbind', sig)

#code reward locations with correponding locations
rew_class = mutate(
  rew_class, reward = case_when(
    location == "4" ~ 'sucrose',
    location == "3" ~ 'water',
    location == "5" ~ 'quinine'
  )
)
rew_class$reward = factor(rew_class$reward, levels=c('sucrose','water','quinine'))

sig$event = recode_factor(
  sig$event,
  "4" = "sucrose",
  "3" = "water",
  "5" = "quinine",
  "low" = "low shock",
  "high" = "high shock"
)

```


# Make exclusions
```{r}

#For 25c5, mouse 1, sucrose port was delivering half volume
rew_class = rew_class[!(rew_class$Experiment=='25c5' & rew_class$Animal=='01'),]
shk_class = shk_class[!(shk_class$Experiment=='25c5' & shk_class$Animal=='01'),]
sig = sig[!(sig$Experiment=='25c5' & sig$Animal=='01'),]

#For 25c5, mouse 8, units 0 and 1 are lens artifact
rew_class = rew_class[
  !(rew_class$Experiment=='25c5' & rew_class$Animal=='08' & (rew_class$unit_id==0|rew_class$unit_id==1)),
]
shk_class = shk_class[
  !(shk_class$Experiment=='25c5' & shk_class$Animal=='08' & (shk_class$unit_id==0|shk_class$unit_id==1)),
]
sig = sig[
  !(sig$Experiment=='25c5' & sig$Animal=='08' & (sig$unit_id==0|sig$unit_id==1)),
]


```

# Pre-process data
## Define classifications and add to signal data
```{r}

thresh = 95
rew_class$rew_responsive = rew_class$loc_pct_shuff>=thresh
shk_class$shk_responsive = shk_class$shock_pct_shuff>=thresh

#add shock classification to reward classification
rew_class = left_join(
  rew_class,
  shk_class[,c('Animal','Experiment','unit_id','shk_responsive')],
  by = c('Animal','Experiment','unit_id')
)

#add shock classification to activity data
sig = left_join(
  sig, shk_class[,c('Animal','Experiment','unit_id','shk_responsive')],
  by = c('Animal','Experiment','unit_id')
)

```


## Give each unit unique id
```{r}

sig = mutate(sig, uniqID = paste(Animal,Experiment,unit_id))
rew_class = mutate(rew_class, uniqID = paste(Animal,Experiment,unit_id))
shk_class = mutate(shk_class, uniqID = paste(Animal,Experiment,unit_id))

```


## Get Average Unit Response for each event (water, sucrose, quinine, low shock, high shock)
```{r}

sig = sig[!is.na(sig$sig),]
sig_avg = summarise_at(
  .tbl = group_by(sig, uniqID, shk_responsive, event, bin),
  .funs = list(sig=mean),
  .vars = c('sig')
)

#Define average post-event response for each unit
period = c(0, 5000) 
sig_avg_post = summarise_at(
  .tbl = group_by(
      sig[sig$bin>period[1] & sig$bin<=period[2],], 
      uniqID, shk_responsive, event
    ),
  .funs = list(sig=mean),
  .vars = c('sig')
)


```


# Get summary statistics for plotting
Calculate relevant means and std. errors.
```{r}

sem = function(x) {sd(x)/sqrt(length(x))}

rew_stats = summarise_at(
  .tbl = group_by(rew_class, shk_responsive, reward),
  .funs = list(mean=mean, sem=sem),
  .vars = c('pre','post','diff')
)
rew_stats_all = summarise_at(
  .tbl = group_by(rew_class, reward),
  .funs = list(mean=mean, sem=sem),
  .vars = c('pre','post','diff')
)

shk_stats = summarise_at(
  .tbl = group_by(shk_class, shk_responsive),
  .funs = list(mean=mean, sem=sem),
  .vars = c('pre','post','diff')
)
shk_stats_all = summarise_at(
  .tbl = group_by(shk_class),
  .funs = list(mean=mean, sem=sem),
  .vars = c('pre','post','diff')
)

sig_stats = summarise_at(
  .tbl = group_by(sig_avg, shk_responsive, event, bin),
  .funs = list(mean=mean, sem=sem),
  .vars = c('sig')
)
sig_stats_all = summarise_at(
  .tbl = group_by(sig_avg, event, bin),
  .funs = list(mean=mean, sem=sem),
  .vars = c('sig')
)

sig_post_stats = summarise_at(
  .tbl = group_by(sig_avg_post, shk_responsive, event),
  .funs = list(mean=mean, sem=sem),
  .vars = c('sig')
)

```



# Plots
## Global Plot Parameters
```{r}

#set figure file prefix
f_pfx = '25c_collapsed_Ca2'

#list of plotting parameters
p = list()

#ggsave
p$save = FALSE
p$savetype = '.svg'

##Plot dimensions
p$dim = list(
  width = 4.5, #Plot width (in)
  height = 4.5, #Plot height (in)
  panel_w = unit(1.5, 'in'),
  panel_h = unit(2, 'in'),
  dpi = 500 #Plot dpi
)

#Plot text
p$textsize = list(
  title = 20,
  legend = 14,
  axis_title = 15,
  axis_text = 14
)

#Plot line dimensions
p$linesize = list(
  data_grp = 1,
  data_ind = .6,
  axis = .5,
  bar = .7,
  error =.75
)
p$linealpha =list(
  data_ind = .3
)

#Plot Widths
p$width = list(
  bar = .8,
  jitter = .2,
  error = .5
)

#Point/dot sizes
p$points = list(
  grp = list(
    shape = 21,
    size = 4,
    stroke = 1,
    alpha = 1
  ),
  ind = list(
    shape =  21,
    size = 2.5,
    alpha = .4
  )
)

#Plot colors
p$cols = list(
  color = c("forestgreen","gray100","yellow2","orange","red2"),
  fill = c("forestgreen","gray100","yellow2","orange","red2"),
  rec = 'slategray1'
)

#theme
p$theme = list(
  base = theme_bw(),
  specs = theme(
    panel.grid = element_blank(), 
    panel.border = element_rect(colour="white"),
    axis.line = element_line(
      linewidth=p$linesize$axis,
      color="black"),
    plot.title = element_text(
      size=p$textsize$title, 
      hjust=.5, 
      margin=margin(b=15,t=10), 
      face="bold"),
    axis.title = element_text(
      size=p$textsize$axis_title, 
      face="plain"),
    axis.text=element_text(
      size=p$textsize$axis_text, 
      colour="black",
      face = 'plain'),
    strip.background = element_blank(),
    strip.text = element_text(
      size=p$textsize$axis_title, 
      colour="black",
      face = 'plain'),
    legend.position = c(.1,.9),
    legend.title = element_blank(),
    legend.text = element_text(size=p$textsize$legend, face="plain"),
    legend.key = element_rect(colour=NA, fill=NA),
    legend.background = element_rect(fill=NA),
    plot.background = element_rect(fill = "transparent", colour = NA),
  ) 
)

saveplot = partial(
  ggsave,
  width = p$dim$width,
  height = p$dim$height,
  units = "in",
  dpi = p$dim$dpi,
  bg = 'transparent'
)





```


## Make Plots
### Response Magnitudes
#### Shock Responsive neurons
##### Avg Post Response for Shock Responsive neurons
```{r, fig.width=4.5, fig.height=4.5}

sig_post_stats$shock = sig_post_stats$event!='low shock' & sig_post_stats$event!='high shock'

#Create base layer from group level data
P.PostResp = ggplot(
  data = sig_post_stats[sig_post_stats$shk_responsive==TRUE,],
  aes(x=event,
      y=mean,
      fill=event,
      group=1))

#Define plot specific parameters
p.ymin = -.1
p.ymax = 1.5#1.5
p.yint = .5
p.breaks = seq(p.ymin,p.ymax,p.yint)

#Create plot
P.PostResp  +
  
  geom_hline(yintercept = 0, linetype='dashed') +
  
  geom_errorbar(
    stat = "identity",
    linewidth = p$linesize$error,
    aes(ymin = mean - sem,
        ymax = mean + sem,
        width = p$width$error*.3
        )) +
  
  
  geom_point(
    stat = "identity",
    size = p$points$grp$size,
    stroke = p$points$grp$stroke,
    shape = p$points$grp$shape,
    show.legend = FALSE
  ) +
  
  scale_fill_manual(values = p$cols$fill) +
  scale_x_discrete(expand = c(.3,.3)) +
  
  facet_grid(rows = vars(shock), scales = 'free_y') +
  facetted_pos_scales(
    y = list(
      scale_y_continuous(limits=c(.2, 2.5), breaks=seq(.5,2.5,.5)),
      scale_y_continuous(limits=c(-.11, .05), breaks=seq(-.2,.1,.1))
    )
  ) +
  
  force_panelsizes(
    rows = c(.7, .3),
    cols = unit(1.3, "in"),
    total_height = p$dim$panel_h
  ) +
  
  labs(
    x = "",
    y = "calcium (z)\n[0-5 sec post]"
  ) +
  
  p$theme +
  theme(
    strip.text = element_blank(),
    axis.text.x= element_text(
      face = 'plain',
      size = p$textsize$axis_title-2,
      angle = -90, 
      hjust = 0, 
      vjust = .5,
      margin = margin(b=2, t=3)
    )
  ) 

if (p$save == TRUE) {saveplot(filename = paste(f_pfx,'_',"AvgPostResp_ShkNeurons", p$savetype, sep = ''))}

```



##### All Responses (faceting)
```{r, fig.width=4.5, fig.height=4.5}

pdata = sig_stats[sig_stats$shk_responsive==TRUE,]
pdata = droplevels(rbind(
  mutate(pdata, facet = 1),
  mutate(pdata, facet = 2)
))


#Create base layer from group level data
P.AllResp.ShkNeurons = ggplot(
  data = pdata,
  aes(x = bin,
      y = mean,
      fill = event,
      group = event))

#Define plot specific parameters
p.ymin = -.2
p.ymax = 2.5
p.yint = .5
p.ybreaks = seq(p.ymin,p.ymax,p.yint)
p.xmin = -2000
p.xmax = 10000
p.xint = 5000
p.xbreaks = seq(0, p.xmax, p.xint)

#Create plot
P.AllResp.ShkNeurons  +
  
  geom_vline(xintercept=0, linetype='dashed') +
  geom_hline(yintercept=0, linetype='dashed') +
  
  geom_ribbon(
    stat = "identity",
    linewidth = p$linesize$error,
    color = 'black',
    aes(ymin = mean - sem,
        ymax = mean + sem
        )) +
  
  scale_fill_manual(values = p$cols$fill) +
  coord_cartesian(xlim = c(p.xmin,p.xmax)) +
  scale_x_continuous(breaks=p.xbreaks, labels=p.xbreaks/1000) +
  
  facet_grid(rows=vars(facet), scales = 'free_y') +
  facetted_pos_scales(
    y = list(
      scale_y_continuous(limits=c(.2, 2.5), breaks=seq(.5,2.5,.5)),
      scale_y_continuous(limits=c(-.11, .05), breaks=seq(-.2,.1,.1))
    )
  ) +
  
  force_panelsizes(
    rows = c(.7, .3),
    cols = unit(1, "in"),
    total_height = p$dim$panel_h
  ) +
  
  labs(
    x = "sec",
    y = "calcium (z)"
  ) +
  
  p$theme +
  theme(
    strip.text = element_blank(),
    axis.text.x= element_text(face = 'plain',size = p$textsize$axis_text),
    legend.title = element_blank(),
    legend.position = 'right'
  ) 

if (p$save == TRUE) {saveplot(filename = paste(f_pfx,'_',"AllResp_ShkNeurons_facet", p$savetype, sep = ''))}

```



### Response proportions
#### calculate proportions
```{r}

#bind reward and shock response classifications
resp_class = bind_rows(
  rew_class[,c('uniqID','reward','rew_responsive')], 
  shk_class[,c('uniqID','shk_responsive')]
)

#define stimulus column
names(resp_class)[names(resp_class)=='reward']='stimulus'
resp_class$stimulus = factor(
  resp_class$stimulus, 
  levels=c('sucrose','water','quinine','shock')
)
resp_class$stimulus[is.na(resp_class$stimulus)]='shock'

#define response classification column
resp_class = mutate(
  resp_class,
  responsive = case_when(
    !is.na(rew_responsive) ~ rew_responsive,
    !is.na(shk_responsive) ~ shk_responsive
  )
)

#create table
resp_table = table(resp_class[,c('responsive','stimulus')])
resp_prop = as.data.frame(prop.table(resp_table,margin=2))

```



#### plot percentages
```{r}

P.percent = ggplot(
  data = resp_prop[resp_prop$responsive==TRUE,],
  aes(x = stimulus,
      y = Freq,
      fill = stimulus,
      group = stimulus))

#Define plot specific parameters
p.ymin = 0
p.ymax = 1
p.yint = .2
p.ybreaks = seq(p.ymin, p.ymax, p.yint)

P.percent +
  
  geom_bar(
    stat = 'identity',
    position = position_dodge(),
    linewidth = p$linesize$bar,
    color = 'black',
    show.legend = FALSE
  ) +
  
  scale_fill_manual(
    values = append(
      p$cols$fill[1:(length(p$cols$fill)-2)], 'red'
    )
  ) +
  
  scale_y_continuous(
    limits = c(p.ymin,p.ymax),
    breaks = p.ybreaks, 
    labels = p.ybreaks*100, 
    expand = c(0,0)
  ) +
  
  labs(
    x = "",
    y = "% Responsive Neurons"
  ) +
  
  p$theme +
  theme(
    axis.text.x= element_text(
      angle = -90,
      hjust = 0,
      vjust = .5
    )
  ) +
  
  force_panelsizes(
    cols = unit(1, "in"),
    total_height = p$dim$panel_h
  ) 

if (p$save == TRUE) {saveplot(filename = paste(f_pfx,'_',"PrcntResponsive", p$savetype, sep = ''))}
  
  

```



# Stats
## Post Response
```{r}

post.anova = ezANOVA(
  data = sig_avg_post[sig_avg_post$shk_responsive,],
  dv = sig,
  wid = uniqID,
  within = event,
  type = 3,
  white.adjust = TRUE
)
post.anova

```

### posthoc
```{r}

comparisons = list(
  c('high shock','low shock'),
  c('low shock','quinine'),
  c('quinine', 'water'),
  c('water', 'sucrose'),
  c('quinine', 'sucrose')
)

posthoc = list()
for (comparison in comparisons) {
  print (comparison)
  posthoc[[paste(comparison[1],comparison[2],sep=' VS ')]] = ezANOVA(
    data = sig_avg_post[
      sig_avg_post$shk_responsive & (
        sig_avg_post$event==comparison[1] | sig_avg_post$event==comparison[2]
      ),
    ],
    dv = sig,
    wid = uniqID,
    within = event,
    type = 3,
    white.adjust = TRUE
  )
}

posthoc[['sucrose vs 0']] = t.test(
  sig_avg_post$sig[
    sig_avg_post$shk_responsive & sig_avg_post$event=='sucrose'
  ],
  mu = 0
)

posthoc[['water vs 0']] = t.test(
  sig_avg_post$sig[
    sig_avg_post$shk_responsive & sig_avg_post$event=='water'
  ],
  mu = 0
)

posthoc[['quinine vs 0']] = t.test(
  sig_avg_post$sig[
    sig_avg_post$shk_responsive & sig_avg_post$event=='quinine'
  ],
  mu = 0
)

posthoc

```


# Export figure data for publication
```{r}

export = FALSE

# export unit classification info
unit_export = resp_class
unit_export = unit_export[,c('uniqID','stimulus','responsive')]

#extract prob. relative to shuffled distribition
pshuff = pivot_wider(
  rew_class[,c('uniqID','loc_pct_shuff', 'reward')],
  names_from = 'reward',
  values_from = 'loc_pct_shuff',
  id_cols = 'uniqID'
)
pshuff = left_join(pshuff, shk_class[,c('uniqID','shock_pct_shuff')], by='uniqID')
names(pshuff) = c('uniqID','water_pct','sucrose_pct','quinine_pct','shock_pct')
unit_export = left_join(unit_export, pshuff, by='uniqID')

unit_export$Animal = as.integer(sapply(unit_export$uniqID, function(x){str_split(x," ")[[1]][1]}))
unit_export$Experiment = sapply(unit_export$uniqID, function(x){str_split(x," ")[[1]][2]})
names(unit_export)[names(unit_export)=='uniqID'] = 'unit'
unit_export = left_join(
  unit_export,
  groups[,c('Animal','Experiment','Sex',"Exclude")],
  by = c('Animal','Experiment')
)
unit_export = unit_export[,c('unit','Animal','Sex','stimulus','responsive','water_pct','sucrose_pct','quinine_pct','shock_pct')]
unit_export$water_pct[unit_export$stimulus!='water']=NA
unit_export$sucrose_pct[unit_export$stimulus!='sucrose']=NA
unit_export$quinine_pct[unit_export$stimulus!='quinine']=NA
unit_export$shock_pct[unit_export$stimulus!='shock']=NA

# export unit response info
unitresponse_export = sig_avg
unitresponse_export$bin = unitresponse_export$bin/1000
unitresponse_export$Animal = as.integer(sapply(unitresponse_export$uniqID, function(x){str_split(x," ")[[1]][1]}))
unitresponse_export$Experiment = sapply(unitresponse_export$uniqID, function(x){str_split(x," ")[[1]][2]})
names(unitresponse_export)[names(unitresponse_export)=='uniqID'] = 'unit'
unitresponse_export = left_join(
  unitresponse_export,
  groups[,c('Animal','Experiment','Sex')],
  by = c('Animal','Experiment')
)
unitresponse_export = unitresponse_export[,
  c('unit','Experiment','Animal','Sex','shk_responsive','event','bin','sig')
]
names(unitresponse_export)[names(unitresponse_export)=='event'] = 'stimulus'
names(unitresponse_export)[names(unitresponse_export)=='bin'] = 'second'
names(unitresponse_export)[names(unitresponse_export)=='sig'] = 'signal'

if (export){
  write.csv(unit_export, 'Fig2m_classification.csv', row.names=FALSE)
  write.csv(unitresponse_export, 'Fig2n_response.csv', row.names=FALSE)
}

```











