---
title: "SEFL25b_StimSensitivity"
author: "Zach"
output:
  html_document: 
    fig_height: 4.5
    fig_width: 4.5
    highlight: kate
    number_sections: yes
    theme: journal
    toc: yes
    toc_depth: 6
editor_options: 
  chunk_output_type: console
---


# Set Global Options
```{r global_options}
knitr::opts_chunk$set(fig.width=4.5, fig.height=4.5, echo=TRUE, warning=FALSE,message=FALSE)
```



# Load Packages 
```{r}

library(arrow)
library(ggplot2)
library(ggrepel)
library(ez)
library(knitr)
library(dplyr)
library(purrr)
library(scales)
library(tidyr)
library(ggh4x)
library(factoextra)

```



# Import Data 
Load  files and factorize as necessary.
```{r}

#load group info
groups = read.csv('../SEFL25b_groups.csv',header=TRUE)

#stim info
stiminfo = read.csv('stiminfo.csv')

#load average cell response
dfiles = dir('UnitClassification//')
data = list()
for (f in dfiles) {data[[f]] = read_feather(paste('UnitClassification/',f,sep=''))}
units = do.call('rbind', data)

#load average cell response
dfiles = dir('BinnedUnitResponses//')
data = list()
for (f in dfiles) {data[[f]] = read_feather(paste('BinnedUnitResponses/',f,sep=''))}
data = do.call('rbind', data)

#factorize as necessary
groups$Animal = factor(groups$Animal)
units$Animal = factor(as.numeric(units$Animal))
data$Animal = factor(as.numeric(data$Animal))
groups$Group = recode_factor(groups$Group, 'NT'='NS', 'T'='S')

#give each unit a unique id
units$uniqID = factor(paste(units$Animal, units$unit_id, sep="_"))
data$uniqID = factor(paste(data$Animal, data$unit_id, sep="_"))

#add group info to data files
units = droplevels(inner_join(x=groups, y=units, by="Animal"))
data = droplevels(inner_join(x=groups, y=data, by="Animal"))

#factor as necessary
data$amp = factor(data$amp, levels=c('high','low'))
data$type = factor(data$type, levels=c('shock','sound'))

```


# Drop subjects
```{r}

data = data[data$StimulusSensitivity==1,]
units = units[units$StimulusSensitivity==1,]

data = droplevels(data[data$Animal!=39,]) #this mouse was run twice. remove if comparining S/NS
units = droplevels(units[units$Animal!=39,]) #this mouse was run twice. remove if comparining S/NS

```


# Add unit categories and prepost
```{r}

unitClass = pivot_wider(
  units[,c('uniqID','Group','type','pct_shuff')],
  id_cols = c('uniqID','Group'),
  names_from = 'type',
  values_from = 'pct_shuff'
)

#here cells are classified only if they display an increase at p=.05, one-tailed
unitClass$ShockResp = unitClass$shock>=95
unitClass$SoundResp = unitClass$sound>=95
unitClass = mutate(
  unitClass, 
  class = case_when(
    (ShockResp & !SoundResp) ~ 'shock',
    (!ShockResp & SoundResp) ~ 'sound',
    (ShockResp & SoundResp) ~ 'both',
    (!ShockResp & !SoundResp) ~ 'neither'
  )
)
unitClass$class = factor(unitClass$class, levels=c('neither','sound','shock','both'))

#add classification to data
data = inner_join(data, unitClass[,c('uniqID','ShockResp','SoundResp','class')], by='uniqID')

```



# Add unit categories and tabulate
```{r}

#tabulate
# all cells
UnitClassTbl = table(unitClass[,c('class')])
UnitClassTblProp = as.data.frame(prop.table(UnitClassTbl))
# by group
UnitClassTbl.bygrp = table(unitClass[,c('Group','class')])
UnitClassTblProp.bygrp = as.data.frame(prop.table(UnitClassTbl.bygrp, margin=1))

fisher.test(
  UnitClassTbl.bygrp
)

```




# Get mean activity response, separated by group and amplitude (cells are 'subjects' here)
```{r}

sem = function(x){sd(x)/sqrt(length(x[!is.na(x)]))}

#get average of each cell across trials
data.unit.collapsed = summarise_at(
  .tbl = group_by(
    .data = data,
    Animal, uniqID, ShockResp, SoundResp, class, Group, type, amp, bin),
  .funs = list(sig=mean),
  .vars = 'sig'
)

#get statistics, separated by classificataion (responsive/nonresponsive) but collapsed across groups (NS/S)
#shock
data.shock.stats = summarise_at(
  .tbl = group_by(
    .data = data[data$type=='shock',],
    type, ShockResp, amp, bin),
  .funs = list(mean=mean, sem=sem),
  .vars = c('sig'))
#sound
data.sound.stats = summarise_at(
  .tbl = group_by(
    .data = data[data$type=='sound',],
    type, SoundResp, amp, bin),
  .funs = list(mean=mean, sem=sem),
  .vars = c('sig'))

#get statistics, separated by classificataion (responsive/nonresponsive) and group (NS/S)
#shock
data.shock.bygroupandclass.stats = summarise_at(
  .tbl = group_by(
    .data = data[data$type=='shock',],
    Group, type, ShockResp, amp, bin),
  .funs = list(mean=mean, sem=sem),
  .vars = c('sig'))

data.sound.bygroupandclass.stats = summarise_at(
  .tbl = group_by(
    .data = data[data$type=='sound',],
    Group, type, SoundResp, amp, bin),
  .funs = list(mean=mean, sem=sem),
  .vars = c('sig'))

#get statistics, separated by group (NS/S) but collapsed across classificataion (responsive/nonresponsive) 
#shock
data.bygroup.stats = summarise_at(
  .tbl = group_by(
    .data = data,
    Group, type, amp, bin),
  .funs = list(mean=mean, sem=sem),
  .vars = c('sig'))

```


# Define average post activity for each unit
```{r}

data.unit.post = summarise_at(
  .tbl = group_by(
    .data = data.unit.collapsed[
      data.unit.collapsed$bin>0 & data.unit.collapsed$bin<=5000
      ,],
    uniqID, Group, ShockResp, SoundResp, class, type, amp),
  .funs = list(sig = mean),
  .vars = c('sig'))

```


# Plots
## Global Plot Parameters
```{r}

#set figure file prefix
f_pfx = '25b'

#list of plotting parameters
p = list()

#ggsave
p$save = FALSE
p$savetype = '.svg'

##Plot dimensions
p$dim = list(
  width = 4.5, #Plot width (in)
  height = 4.5, #Plot height (in)
  panel_w = unit(1.5, 'in'),
  panel_h = unit(2, 'in'),
  dpi = 500 #Plot dpi
)

#Plot text
p$textsize = list(
  title = 20,
  legend = 16,
  axis_title = 16,
  axis_text = 14
)

#Plot line dimensions
p$linesize = list(
  data_grp = 1,
  data_ind = .6,
  axis = .5,
  bar = .7,
  error =.75
)
p$linealpha =list(
  data_ind = .2
)

#Plot Widths
p$width = list(
  bar = .8,
  jitter = .3,
  error = .5
)

#Point/dot sizes
p$points = list(
  grp = list(
    shape = 21,
    size = 4,
    stroke = 1,
    alpha = 1
  ),
  ind = list(
    shape =  21,
    size = 2.5,
    alpha = .4
  )
)

#Plot colors
p$cols = list(
  fill_shk = c('orangered3', 'orange2'),
  fill_snd = c('blue3','cornflowerblue'),
  fill_grp = c('gray80','red4')
)

#theme
p$theme = list(
  base = theme_bw(),
  specs = theme(
    panel.grid = element_blank(), 
    panel.border = element_rect(colour="white"),
    axis.line = element_line(
      linewidth=p$linesize$axis,
      color="black"),
    plot.title = element_text(
      size=p$textsize$title, 
      hjust=.5, 
      margin=margin(b=15,t=10), 
      face="bold"),
    axis.title = element_text(
      size=p$textsize$axis_title, 
      face="plain"),
    axis.text=element_text(
      size=p$textsize$axis_text, 
      colour="black",
      face = 'plain'),
    strip.background = element_blank(),
    strip.text = element_text(
      size=p$textsize$axis_title, 
      colour="black",
      face = 'plain'),
    legend.position = c(.1,.9),
    legend.title = element_blank(),
    legend.text = element_text(size=p$textsize$legend, face="plain"),
    legend.key = element_rect(colour=NA, fill=NA),
    legend.background = element_rect(fill=NA),
    plot.background = element_rect(fill = "transparent", colour = NA),
  ) 
)

saveplot = partial(
  ggsave,
  width = p$dim$width,
  height = p$dim$height,
  units = "in",
  dpi = p$dim$dpi,
  bg = 'transparent'
)


```


# Make Plots

## Groups collapsed
### Shock responsive
```{r, fig.width=4.5, fig.height=4.5}

#Create base layer from group level data
P.unit.SHKresponsive.LOWvsHIGH = ggplot(
  data = data.shock.stats[data.shock.stats$ShockResp,],
  aes(x = bin,
      y = mean,
      fill = amp,
      group = amp))


#Define plot specific parameters
p.xmin = -2000
p.xmax = 10000
p.xint = 5000
p.xbreaks = seq(0, p.xmax, p.xint)
p.ymin = -.1
p.ymax = 3
p.yint = 1
p.ybreaks = seq(0,p.ymax,p.yint)

#Create plot
P.unit.SHKresponsive.LOWvsHIGH  +
  
  #geom_rect(aes(xmin = 0, xmax = 2000, ymin = -Inf, ymax = Inf), fill = "lightsteelblue1") +
  #geom_hline(yintercept = 0, linetype = 'dashed') +
  geom_vline(xintercept = c(0,2000), linetype='dashed', color='gray50') +
  
  geom_ribbon(
    stat = "identity",
    linewidth = p$linesize$data_ind*.9,
    colour = 'black',
    show.legend = TRUE,
    aes(ymin = mean - sem,
        ymax = mean + sem)) +
  
  scale_fill_manual(values=p$cols$fill_shk) +
  scale_y_continuous(breaks=p.ybreaks, labels = p.ybreaks, expand = c(0, 0)) +
  scale_x_continuous(breaks=p.xbreaks, labels=p.xbreaks/1000) +
  coord_cartesian(ylim=c(p.ymin,p.ymax), xlim=c(-1000,p.xmax)) +
  
  labs(
    x = 't (sec)',
    y = "calcium (z)"
  ) +
  
  p$theme$base +
  p$theme$specs +
  theme(
    legend.position = c(.8, .9),
    legend.key.size = unit(.5, 'cm'),
    panel.spacing.x = unit(.25,'cm'),
  ) +

  force_panelsizes(
      rows = p$dim$panel_h,
      cols = unit(1., "in")
    ) 
  
  if (p$save == TRUE) {saveplot(filename = paste(f_pfx,'_',"UnitsAsSubjs_ShkResponsive_LOWvsHIGH", p$savetype, sep = ''))}

```

### Sound responsive
```{r, fig.width=4.5, fig.height=4.5}

#Create base layer from group level data
P.unit.SNDresponsive.LOWvsHIGH = ggplot(
  data = data.sound.stats[data.sound.stats$SoundResp,],
  aes(x = bin,
      y = mean,
      fill = amp,
      group = amp))


#Define plot specific parameters
p.xmin = -2000
p.xmax = 10000
p.xint = 5000
p.xbreaks = seq(0,p.xmax,p.xint)
p.ymin = -.1
p.ymax = 3
p.yint = 1
p.ybreaks = seq(0,p.ymax,p.yint)

#Create plot
P.unit.SNDresponsive.LOWvsHIGH  +
  
  # geom_rect(aes(xmin = 0, xmax = 2000, ymin = -Inf, ymax = Inf), fill = "lightsteelblue1") +
  # 
  # geom_hline(yintercept = 0, linetype = 'dashed') +
  
  geom_vline(xintercept = c(0,2000), linetype='dashed', color='gray50') +
  
  geom_ribbon(
    stat = "identity",
    linewidth = p$linesize$data_ind*.9,
    colour = 'black',
    show.legend = TRUE,
    aes(ymin = mean - sem,
        ymax = mean + sem)) +
  
  scale_fill_manual(values=p$cols$fill_snd) +
  scale_y_continuous(breaks=p.ybreaks, labels = p.ybreaks, expand = c(0, 0)) +
  scale_x_continuous(breaks=p.xbreaks, labels=p.xbreaks/1000) +
  coord_cartesian(ylim=c(p.ymin,p.ymax), xlim=c(-1000,p.xmax)) +
  
  labs(
    x = 't (sec)',
    y = "calcium (z)"
  ) +
  
  p$theme$base +
  p$theme$specs +
  theme(
    legend.position = c(.8, .9),
    legend.key.size = unit(.5, 'cm'),
    panel.spacing.x = unit(.25,'cm'),
  ) +

  force_panelsizes(
      rows = p$dim$panel_h,
      cols = unit(1., "in")
    ) 
  
if (p$save == TRUE) {saveplot(filename = paste(f_pfx,'_',"UnitsAsSubjs_SndResponsive_LOWvsHIGH", p$savetype, sep = ''))}

```


## Groups separated
### All cells
#### Shock
```{r, fig.width=4.5, fig.height=4.5}

#Create base layer from group level data
P.unit.all.bygroup = ggplot(
  data = data.bygroup.stats[data.bygroup.stats$type=='shock',],
  aes(x = bin,
      y = mean,
      fill = Group,
      group = Group))


#Define plot specific parameters
p.xmin = -2000
p.xmax = 10000
p.xint = 5000
p.xbreaks = seq(0, p.xmax, p.xint)
p.ymin = -.2
p.ymax = 3
p.yint = 1
p.ybreaks = seq(0,p.ymax,p.yint)

#Create plot
P.unit.all.bygroup  +
  
  #geom_rect(aes(xmin = 0, xmax = 2000, ymin = -Inf, ymax = Inf), fill = "lightsteelblue1") +
  #geom_hline(yintercept = 0, linetype = 'dashed') +
  geom_vline(xintercept = c(0,2000), linetype='dashed', color='gray50') +
  
  geom_ribbon(
    stat = "identity",
    linewidth = p$linesize$data_ind*.9,
    colour = 'black',
    show.legend = TRUE,
    aes(ymin = mean - sem,
        ymax = mean + sem)) +

  facet_grid(~factor(amp, levels=c('low','high'))) +
  
  scale_fill_manual(values=p$cols$fill_grp) +
  scale_y_continuous(breaks=p.ybreaks, labels = p.ybreaks, expand = c(0, 0)) +
  scale_x_continuous(breaks=p.xbreaks, labels=p.xbreaks/1000) +
  coord_cartesian(ylim=c(p.ymin,p.ymax), xlim=c(-1000,p.xmax)) +
  
  labs(
    x = 't (sec)',
    y = "calcium (z)"
  ) +
  
  p$theme$base +
  p$theme$specs +
  theme(
    legend.position = c(.175, .9),
    legend.key.size = unit(.5, 'cm'),
    panel.spacing.x = unit(.25,'cm'),
  ) +

  force_panelsizes(
      rows = p$dim$panel_h,
      cols = unit(1, "in")
    ) 
  
  if (p$save == TRUE) {saveplot(filename = paste(f_pfx,'_',"UnitsAsSubjs_All_Shk_SvsNS", p$savetype, sep = ''))}

```

#### Sound
```{r, fig.width=4.5, fig.height=4.5}

#Create base layer from group level data
P.unit.all.bygroup = ggplot(
  data = data.bygroup.stats[data.bygroup.stats$type=='sound',],
  aes(x = bin,
      y = mean,
      fill = Group,
      group = Group))

#Define plot specific parameters
p.xmin = -2000
p.xmax = 10000
p.xint = 5000
p.xbreaks = seq(0, p.xmax, p.xint)
p.ymin = -.2
p.ymax = 2
p.yint = 1
p.ybreaks = seq(0,p.ymax,p.yint)

#Create plot
P.unit.all.bygroup  +
  
  #geom_rect(aes(xmin = 0, xmax = 2000, ymin = -Inf, ymax = Inf), fill = "lightsteelblue1") +
  #geom_hline(yintercept = 0, linetype = 'dashed') +
  geom_vline(xintercept = c(0,2000), linetype='dashed', color='gray50') +
  
  geom_ribbon(
    stat = "identity",
    linewidth = p$linesize$data_ind*.9,
    colour = 'black',
    show.legend = TRUE,
    aes(ymin = mean - sem,
        ymax = mean + sem)) +
  
  facet_grid(~factor(amp, levels=c('low','high'))) +
  
  scale_fill_manual(values=p$cols$fill_grp) +
  scale_y_continuous(breaks=p.ybreaks, labels = p.ybreaks, expand = c(0, 0)) +
  scale_x_continuous(breaks=p.xbreaks, labels=p.xbreaks/1000) +
  coord_cartesian(ylim=c(p.ymin,p.ymax), xlim=c(-1000,p.xmax)) +
  
  labs(
    x = 't (sec)',
    y = "calcium (z)"
  ) +
  
  p$theme$base +
  p$theme$specs +
  theme(
    legend.position = c(.175, .9),
    legend.key.size = unit(.5, 'cm'),
    panel.spacing.x = unit(.25,'cm'),
  ) +

  force_panelsizes(
      rows = p$dim$panel_h,
      cols = unit(1., "in")
    ) 

  force_panelsizes(
      rows = p$dim$panel_h,
      cols = unit(1.5, "in")
    ) 
  
  if (p$save == TRUE) {saveplot(filename = paste(f_pfx,'_',"UnitsAsSubjs_All_Snd_SvsNS", p$savetype, sep = ''))}

```

### Shock responsive
```{r, fig.width=4.5, fig.height=4.5}

#Create base layer from group level data
P.unit.SHKresponsive.bygroup = ggplot(
  data = data.shock.bygroupandclass.stats[
    data.shock.bygroupandclass.stats$ShockResp,
  ],
  aes(x = bin,
      y = mean,
      fill = Group,
      group = Group))


#Define plot specific parameters
p.xmin = -2000
p.xmax = 10000
p.xint = 5000
p.xbreaks = seq(0, p.xmax, p.xint)
p.ymin = -.2
p.ymax = 3
p.yint = 1
p.ybreaks = seq(0,p.ymax,p.yint)

#Create plot
P.unit.SHKresponsive.bygroup  +
  
  #geom_rect(aes(xmin = 0, xmax = 2000, ymin = -Inf, ymax = Inf), fill = "lightsteelblue1") +
  #geom_hline(yintercept = 0, linetype = 'dashed') +
  geom_vline(xintercept = c(0,2000), linetype='dashed', color='gray50') +
  
  geom_ribbon(
    stat = "identity",
    linewidth = p$linesize$data_ind*.9,
    colour = 'black',
    show.legend = TRUE,
    aes(ymin = mean - sem,
        ymax = mean + sem)) +
  
  facet_grid(~factor(amp, levels=c('low','high'))) +
  
  scale_fill_manual(values=p$cols$fill_grp) +
  scale_y_continuous(breaks=p.ybreaks, labels = p.ybreaks, expand = c(0, 0)) +
  scale_x_continuous(breaks=p.xbreaks, labels=p.xbreaks/1000) +
  coord_cartesian(ylim=c(p.ymin,p.ymax), xlim=c(-1000,p.xmax)) +
  
  labs(
    x = 't (sec)',
    y = "calcium (z)"
  ) +
  
  p$theme$base +
  p$theme$specs +
  theme(
    legend.position = c(.175, .9),
    legend.key.size = unit(.5, 'cm'),
    panel.spacing.x = unit(.25,'cm'),
  ) +

  force_panelsizes(
      rows = p$dim$panel_h,
      cols = unit(1., "in")
    ) 
  
if (p$save == TRUE) {saveplot(filename = paste(f_pfx,'_',"UnitsAsSubjs_ShkResponsive_SvsNS", p$savetype, sep = ''))}

```


### Sound responsive
```{r, fig.width=4.5, fig.height=4.5}

#Create base layer from group level data
P.unit.SNDresponsive.bygroup = ggplot(
  data = data.sound.bygroupandclass.stats[
    data.sound.bygroupandclass.stats$SoundResp,
  ],
  aes(x = bin,
      y = mean,
      fill = Group,
      group = Group))


#Define plot specific parameters
p.xmin = -2000
p.xmax = 10000
p.xint = 5000
p.xbreaks = seq(0, p.xmax, p.xint)
p.ymin = -.2
p.ymax = 3
p.yint = 1
p.ybreaks = seq(0,p.ymax,p.yint)

#Create plot
P.unit.SNDresponsive.bygroup  +
  
  #geom_rect(aes(xmin = 0, xmax = 2000, ymin = -Inf, ymax = Inf), fill = "lightsteelblue1") +
  #geom_hline(yintercept = 0, linetype = 'dashed') +
  geom_vline(xintercept = c(0,2000), linetype='dashed', color='gray50') +
  
  geom_ribbon(
    stat = "identity",
    linewidth = p$linesize$data_ind*.9,
    colour = 'black',
    show.legend = TRUE,
    aes(ymin = mean - sem,
        ymax = mean + sem)) +
  
  facet_grid(~factor(amp, levels=c('low','high'))) +
  
  scale_fill_manual(values=p$cols$fill_grp) +
  scale_y_continuous(breaks=p.ybreaks, labels = p.ybreaks, expand = c(0, 0)) +
  scale_x_continuous(breaks=p.xbreaks, labels=p.xbreaks/1000) +
  coord_cartesian(ylim=c(p.ymin,p.ymax), xlim=c(-1000,p.xmax)) +
  
  labs(
    x = 't (sec)',
    y = "calcium (z)"
  ) +
  
  p$theme$base +
  p$theme$specs +
  theme(
    legend.position = c(.175, .9),
    legend.key.size = unit(.5, 'cm'),
    panel.spacing.x = unit(.25,'cm'),
  ) +

  force_panelsizes(
      rows = p$dim$panel_h,
      cols = unit(1., "in")
    ) 

  
if (p$save == TRUE) {saveplot(filename = paste(f_pfx,'_',"UnitsAsSubjs_SndResponsive_SvsNS", p$savetype, sep = ''))}

```


## Bar Chart
### All cells
```{r}


P.bar = ggplot(
  UnitClassTblProp, 
  aes(x = 1, y = Freq, fill = class, label=class)
) 

p.ymin = 0
p.ymax = 1
p.yint = .25
p.ybreaks = seq(p.ymin,p.ymax,p.yint)

P.bar +
  
  geom_col(
    width = 2,
    colour = 'gray10',
    #linewidth = ,
    show.legend = FALSE
  ) + 
  
  force_panelsizes(
      rows = p$dim$panel_h,
      cols = unit(1, "in")
  ) +
  
  geom_text(
    aes(
      x=2.3
    ),
    size = 5,
    fontface = 'plain',
    angle = 35,
    position = position_stack(vjust = 0.5),
    vjust = .5,
    hjust = 0
  ) +
  
  labs(
    x = '',
    y = "% Responsive Neurons"
  ) +
  
  scale_y_continuous(
    breaks = p.ybreaks, labels = p.ybreaks*100, expand = c(.005,.005)
  ) +
  scale_x_continuous(breaks=c(), limits=c(0,5)) +
  scale_fill_manual(values=c('gray90','blue2','orangered3','purple2')) +
  
  p$theme$base +
  p$theme$specs +
  theme(
    axis.line.x = element_blank()
  ) 

if (p$save == TRUE) {saveplot(filename = paste(f_pfx,'_',"StimSensitivity_BarClass_ALL", p$savetype, sep = ''))}

```

### By Group
```{r}


P.bar = ggplot(
  UnitClassTblProp.bygrp, 
  aes(x = Group, y = Freq, fill = class, label=class)
) 

p.ymin = 0
p.ymax = 1
p.yint = .25
p.ybreaks = seq(p.ymin,p.ymax,p.yint)

P.bar +
  
  geom_col(
    #width = 2,
    colour = 'gray10',
    #linewidth = ,
    show.legend = TRUE
  ) + 
  
  force_panelsizes(
      rows = p$dim$panel_h,
      cols = unit(1, "in")
  ) +
  
  labs(
    x = '',
    y = "% Responsive Neurons"
  ) +
  
  scale_y_continuous(
    breaks = p.ybreaks, labels = p.ybreaks*100, expand = c(.005,.005)
  ) +

  scale_fill_manual(values=c('gray90','blue2','orangered3','purple2')) +
  
  p$theme$base +
  p$theme$specs +
  theme(
    axis.text.x = element_text(size=p$textsize$axis_title),
    axis.line.x = element_blank(),
    legend.title = element_blank(),
    legend.position = 'right'
  ) 

if (p$save == TRUE) {saveplot(filename = paste(f_pfx,'_',"StimSensitivity_BarClass_byGrp", p$savetype, sep = ''))}

```



## Raster
### Shock resp
```{r}

pdata = droplevels(data.unit.collapsed[data.unit.collapsed$type=='shock' & data.unit.collapsed$ShockResp,])
pdata$amp = factor(pdata$amp, levels = c('low','high'))


#Define plot specific parameters
p.xmin = -5000
p.xmax = 10000
p.xint = 5000
p.xbreaks = seq(p.xmin,p.xmax,p.xint)
p.ymin = -.1
p.ymax = 3
p.yint = 1
p.ybreaks = seq(0,p.ymax,p.yint)

P.raster.shock = ggplot(
  data = pdata, 
  aes(x = bin, y = uniqID, fill = sig)
) 

P.raster.shock +
  
  geom_raster() +
  
  geom_vline(xintercept = 0, linetype='dashed', color='red') +
  geom_vline(xintercept = 2000, linetype='dashed', color='red') +
  
  facet_grid(cols=vars(amp)) +
  
  scale_fill_gradient(
    low = 'gray93',
    high = 'black',
    limits=c(0,3),
    oob=scales::squish,
    name = 'z'
  ) +
  
  scale_x_continuous(breaks=p.xbreaks, labels=p.xbreaks/1000) +
  coord_cartesian(xlim=c(-2000,p.xmax)) +
  
  labs(
    x = 't (sec)',
    y = "neuron",
  ) +
  
  p$theme$base +
  p$theme$specs +
  theme(
    legend.title = element_text(
      size=p$textsize$axis_text, 
      vjust = 1
    ),
    legend.position = 'bottom',
    legend.key.size = unit(.4, 'cm'),
    legend.text = element_text(face='plain', size=p$textsize$axis_text-2),
    panel.spacing.x = unit(.5, 'cm'),
    strip.background = element_blank(),
    strip.text = element_text(
      size=p$textsize$axis_title, 
      hjust=.5
    ),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.text.x = element_text(size=p$textsize$axis_text-2),
    axis.title.x = element_text(margin = margin(t=0, b=0))
  ) +

  force_panelsizes(
      rows = p$dim$panel_h,
      cols = unit(.45, "in")
    ) 
  
if (p$save == TRUE) {saveplot(filename = paste(f_pfx,'_',"Raster_ShkResponsive", p$savetype, sep = ''))}

```


### Sound resp
```{r}

pdata = droplevels(data.unit.collapsed[data.unit.collapsed$type=='sound' & data.unit.collapsed$SoundResp,])
pdata$amp = factor(pdata$amp, levels = c('low','high'))


#Define plot specific parameters
p.xmin = -5000
p.xmax = 10000
p.xint = 5000
p.xbreaks = seq(p.xmin,p.xmax,p.xint)
p.ymin = -.1
p.ymax = 3
p.yint = 1
p.ybreaks = seq(0,p.ymax,p.yint)

P.raster.sound = ggplot(
  data = pdata, 
  aes(x = bin, y = uniqID, fill = sig)
) 

P.raster.sound +
  
  geom_raster() +
  
  geom_vline(xintercept = 0, linetype='dashed', color='red') +
  geom_vline(xintercept = 2000, linetype='dashed', color='red') +
  
  facet_grid(cols=vars(amp)) +
  
  scale_fill_gradient(
    low = 'gray93',
    high = 'black',
    limits=c(0,3),
    oob=scales::squish,
    name = 'z'
  ) +
  
  scale_x_continuous(breaks=p.xbreaks, labels=p.xbreaks/1000) +
  coord_cartesian(xlim=c(-2000,p.xmax)) +
  
  labs(
    x = 't (sec)',
    y = "neuron",
  ) +
  
  p$theme$base +
  p$theme$specs +
  theme(
    legend.title = element_text(
      size=p$textsize$axis_text, 
      vjust = 1
    ),
    legend.position = 'bottom',
    legend.key.size = unit(.4, 'cm'),
    legend.text = element_text(face='plain', size=p$textsize$axis_text-2),
    panel.spacing.x = unit(.5, 'cm'),
    strip.background = element_blank(),
    strip.text = element_text(
      size=p$textsize$axis_title, 
      hjust=.5
    ),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.text.x = element_text(size=p$textsize$axis_text-2),
    axis.title.x = element_text(margin = margin(t=0, b=0))
  ) +

  force_panelsizes(
      rows = p$dim$panel_h,
      cols = unit(.45, "in")
  ) 
  
if (p$save == TRUE) {saveplot(filename = paste(f_pfx,'_',"Raster_SndResponsive", p$savetype, sep = ''))}

```


## PCA
### Compute
```{r}

pcdata = pivot_wider(
  data = data[,c('uniqID','trial','type','amp','bin','sig')],
  values_from = 'sig',
  names_from = "uniqID"
)
pcdata = pcdata[complete.cases(pcdata),]

pca = prcomp(pcdata[,-(1:4)])

pcdataX = cbind(pcdata, as.data.frame(pca$x))
pcdataX = mutate(
  pcdataX,
  case = case_when(
    bin <= 0 ~ 'pre-stimulus',
    bin > 0 & type=='sound' & amp=='low' ~ 'low sound',
    bin > 0 & type=='sound' & amp=='high' ~ 'high sound',
    bin > 0 & type=='shock' & amp=='low' ~ 'low shock',
    bin > 0 & type=='shock' & amp=='high' ~ 'high shock'
  )
)
pcdataX$case = factor(
  pcdataX$case, 
  levels = c(
    'pre-stimulus','low sound','high sound','low shock','high shock'
  )
)

pcdataXmean = summarise_at(
  .tbl = group_by(
    .data = pcdataX,
    case),
  .funs = list(mean),
  .vars = c('PC1','PC2','PC3','PC4','PC5')
)


```


### scree plot
```{r}

pca_summary = as.data.frame(t(summary(pca)$importance))
names(pca_summary) = c('stdev','proportion','cumprop')

p.eigen = ggplot(
  data = pca_summary,
  aes(y = proportion, x=1:nrow(pca_summary)))

p.eigen +
  geom_line() +
  scale_x_continuous(limits = c(1,10), breaks=1:10)

```

### plot
```{r}


p.pca_data = ggplot(
  data = pcdataXmean,
  aes(x=PC1, y=PC2, color=case)
)

#Define plot specific parameters
p.xmin = -5
p.xmax = 18
p.xint = 5
p.xbreaks = seq(p.xmin,p.xmax,p.xint)
p.ymin = -5
p.ymax = 18
p.yint = 5
p.ybreaks = seq(p.ymin,p.ymax,p.yint)

p.pca_data +
  
  
  geom_hline(yintercept=0, linetype='dashed', color='gray80') +
  geom_vline(xintercept=0, linetype='dashed', color='gray80') +
  
  geom_point(size=4, alpha=1) +
  
  scale_x_continuous(
    breaks=p.xbreaks, labels=p.xbreaks, 
    limits = c(p.xmin,p.xmax), expand = c(0, 0)) +
  scale_y_continuous(
    breaks=p.ybreaks, labels=p.ybreaks, 
    limits = c(p.ymin,p.ymax), expand = c(0, 0)) +
  scale_color_manual(
    values=c('gray50','cornflowerblue','blue3','orange2','orangered3')
  ) +
  
  p$theme$base +
  p$theme$specs +
  theme(
    legend.position = 'right',
    legend.title = element_blank(),
    legend.text = element_text(
      size=p$textsize$axis_text-4,
      face='plain'
    )
  ) +
  
  force_panelsizes(
      rows = unit(1.5, "in"),
      cols = unit(1.5, "in")
  ) 

if (p$save == TRUE) {saveplot(filename = paste(f_pfx,'_',"PCAplot", p$savetype, sep = ''))}

```



# Stats
## S/NS collapsed
### Shock response (shock responsive)
```{r}

anova.shock.shockresp = ezANOVA(
  data = data.unit.post[
    data.unit.post$ShockResp &data.unit.post$type == 'shock',
  ],
  wid = uniqID,
  dv = sig,
  within = c(amp),
  white.adjust = TRUE,
  type = 3
)
anova.shock.shockresp

```
### Sound response (sound responsive )
```{r}

anova.sound.soundresp = ezANOVA(
  data = data.unit.post[
    data.unit.post$SoundResp & data.unit.post$type == 'sound',
  ],
  wid = uniqID,
  dv = sig,
  within = c(amp),
  white.adjust = TRUE,
  type = 3
)
anova.sound.soundresp

```

## Group (S/NS) comparison
### All neurons
#### Shock
```{r}

anova.shock.all.bygroup = ezANOVA(
  data = data.unit.post[
    data.unit.post$type == 'shock',
  ],
  wid = uniqID,
  dv = sig,
  between = Group,
  within = c(amp),
  white.adjust = TRUE,
  type = 3
)
anova.shock.all.bygroup

anova.shock.low.all.bygroup = ezANOVA(
  data = data.unit.post[
    data.unit.post$type == 'shock' &
      data.unit.post$amp == 'low',
  ],
  wid = uniqID,
  dv = sig,
  between = Group,
  white.adjust = TRUE,
  type = 3
)
anova.shock.low.all.bygroup

anova.shock.high.all.bygroup = ezANOVA(
  data = data.unit.post[
    data.unit.post$type == 'shock' &
      data.unit.post$amp == 'high',
  ],
  wid = uniqID,
  dv = sig,
  between = Group,
  white.adjust = TRUE,
  type = 3
)
anova.shock.high.all.bygroup

```

#### Sound
```{r}

anova.sound.all.bygroup = ezANOVA(
  data = data.unit.post[
    data.unit.post$type == 'sound',
  ],
  wid = uniqID,
  dv = sig,
  between = Group,
  within = c(amp),
  white.adjust = TRUE,
  type = 3
)
anova.sound.all.bygroup

anova.sound.low.all.bygroup = ezANOVA(
  data = data.unit.post[
    data.unit.post$type == 'sound' &
      data.unit.post$amp == 'low',
  ],
  wid = uniqID,
  dv = sig,
  between = Group,
  white.adjust = TRUE,
  type = 3
)
anova.sound.low.all.bygroup

anova.sound.high.all.bygroup = ezANOVA(
  data = data.unit.post[
    data.unit.post$type == 'sound' &
      data.unit.post$amp == 'high',
  ],
  wid = uniqID,
  dv = sig,
  between = Group,
  white.adjust = TRUE,
  type = 3
)
anova.sound.high.all.bygroup

```


### Responsive neurons
#### Shock
```{r}

anova.shock.shockresp.bygroup = ezANOVA(
  data = data.unit.post[
    data.unit.post$ShockResp &
      data.unit.post$type == 'shock',
  ],
  wid = uniqID,
  dv = sig,
  between = Group,
  within = c(amp),
  white.adjust = TRUE,
  type = 3
)
anova.shock.shockresp.bygroup

```

#### Sound
```{r}

anova.sound.soundresp.bygroup = ezANOVA(
  data = data.unit.post[
    data.unit.post$SoundResp &
      data.unit.post$type == 'sound',
  ],
  wid = uniqID,
  dv = sig,
  between = Group,
  within = c(amp),
  white.adjust = TRUE,
  type = 3
)
anova.sound.soundresp.bygroup

```


# Export figure data for publication
```{r}

export = FALSE

# export unit classification info
unit_export = unitClass[,
  c('Group','uniqID','shock','sound','class')
]
unit_export = left_join(
  unique(data[,c('Animal','Sex','uniqID')]),
  unit_export,
  by = 'uniqID'
)
unit_export = unit_export[,c('Group','uniqID','Animal','Sex','shock','sound','class')]

names(unit_export)[names(unit_export)=='uniqID'] = 'unit'
names(unit_export)[names(unit_export)=='shock'] = 'shock_pct'
names(unit_export)[names(unit_export)=='sound'] = 'sound_pct'


# export mean unit responses
unitresponse_export = data.unit.collapsed
names(unitresponse_export)[names(unitresponse_export)=='uniqID'] = 'unit'
names(unitresponse_export)[names(unitresponse_export)=='sig'] = 'signal'
names(unitresponse_export)[names(unitresponse_export)=='type'] = 'stimulus'
names(unitresponse_export)[names(unitresponse_export)=='amp'] = 'amplitude'
unitresponse_export$second = unitresponse_export$bin/1000
unitresponse_export = left_join(
  unitresponse_export,
  unit_export[,c('unit','Sex')],
  by = 'unit'
)
unitresponse_export = unitresponse_export[
  ,
  c('unit','Animal','Sex','Group','class','stimulus','amplitude','second','signal')
]

pcdata_export = pcdata
names(pcdata_export)[names(pcdata_export)=='bin'] = 'second'
pcdata_export$second = pcdata_export$second/1000

if (export){
  write.csv(unit_export, 'Fig2g_units.csv', row.names=FALSE)
  write.csv(unitresponse_export, 'Fig2hi_unitresponse.csv', row.names=FALSE)
  write.csv(pcdata_export, 'Fig2j_unitresponse.csv', row.names=FALSE)
}

```











