---
title: "SEFL10N6d_S2_Analysis"
author: "Zach"
output:
  html_document: 
    fig_height: 4.5
    fig_width: 4.5
    highlight: kate
    number_sections: yes
    theme: journal
    toc: yes
    toc_depth: 6
editor_options: 
  chunk_output_type: console
---


# Set Global Options
```{r global_options}
knitr::opts_chunk$set(fig.width=4.5, fig.height=4.5, echo=TRUE, warning=FALSE,message=FALSE)
```



# Load Packages 
```{r}

library(ggplot2)
library(ez)
library(knitr)
library(tidyr)
library(car) 
library(dplyr)
library(purrr)
library(lme4)
library(lmerTest)
library(ggh4x)

```



# Import Data 
Load  files and factorize as necessary.
```{r}

#load  files
data_fz = read.csv('SEFL10N6d_S2_fzdata.csv',header=TRUE)
groups = read.csv('../SEFL10N6d_groups.csv',header=TRUE)

#factorize as necessary
groups$Animal = factor(groups$Animal)
groups$NS = factor(groups$NS, levels=c('NL','L'))
data_fz$Animal = factor(data_fz$Animal)

# #add group info to data files
data_fz = inner_join(x=groups,y=data_fz,by="Animal")

```



# Drop subjects
See `groups` dataframe for exclusion rationale.
```{r}

data_fz = data_fz[data_fz$Exclude!=1,]


```


# Subset data
Separate phases (trauma, trauma recall, etc) are placed in unique dataframes for ease of analysis.
```{r}


#trauma
#freezing
trauma.freezing = data_fz[data_fz$phase=='trauma' & data_fz$subphase!="shock",]

#trauma shock reactivity
trauma.shockreact = summarise_at(
  .tbl = group_by(
    .data = data_fz[data_fz$phase=='trauma' & data_fz$subphase!="postshock",],
    Animal, subphase, NS),
  .funs = list(motion=mean),
  .vars = c('motion')
)

#trauma recall
traumatest = data_fz[data_fz$phase=='traumatest' & data_fz$subphase=='avg',]


#nov stressor data

##prepost startle stimulus
novstressor.prepost = data_fz[data_fz$phase=='novstressor' & data_fz$subphase=='prepost',]

##startle response
novstressor.startle = data_fz[data_fz$phase=='novstressor' & data_fz$subphase=='startle',]
novstressor.startle$component = factor(
  novstressor.startle$component,
  levels=c('startle_pre','startle_startle','startle_post'))


#novel stressor (sefl) test data
sefltest = data_fz[data_fz$phase=='sefltest' & data_fz$subphase=='avg',]


# trauma recall 2/3
# drop mice 27 & 28, who were given No Light twice, by mistake
traumatest_light = data_fz[data_fz$phase=='traumatest2' | data_fz$phase=='traumatest3',]
traumatest_light$subphase = factor(traumatest_light$subphase, levels=c('NL','L'))
traumatest_light = mutate(
  traumatest_light,
  light = case_when(
    phase == 'traumatest2' ~ Trec2,
    phase == 'traumatest3' ~ Trec3
  ))
traumatest_light = traumatest_light[
  traumatest_light$Animal != 17 & traumatest_light$Animal != 19,
]
#collapse data
traumatest_light.collapsed = summarise_at(
  .tbl = group_by(
    .data = traumatest_light,
    Animal, light, subphase),
  .funs = list(mean),
  .vars = c('freezing','motion'))


```


# Get summary statistics for plotting
Calculate relevant means and std. errors.
```{r}


#trauma session
#freezing
trauma.freezing.stats = summarise_at(
  .tbl = group_by(
    .data = trauma.freezing,
    component, NS),
  .funs = list(mean=~mean(.), sem=~sd(.)/sqrt(n())),
  .vars = c('motion','freezing')
)
trauma.freezing.stats$subphase = 'postshock'
trauma.freezing.stats$subphase[trauma.freezing.stats$component=='baseline'] = 'baseline'

#shock reactivity
trauma.shockreact.stats = summarise_at(
  .tbl = group_by(
    .data = trauma.shockreact,
    subphase, NS),
  .funs = list(mean=~mean(.), sem=~sd(.)/sqrt(n())),
  .vars = c('motion')
)

#trauma recall
traumatest.stats = summarise_at(
  .tbl = group_by(
    .data = traumatest,
    NS),
  .funs = list(mean=~mean(.), sem=~sd(.)/sqrt(n())),
  .vars = c('freezing','motion')
)


#novel stressor
#novel stressor pre/post
novstressor.prepost.stats = summarise_at(
  .tbl = group_by(
    .data = novstressor.prepost,
    component, NS),
  .funs = list(mean=~mean(.), sem=~sd(.)/sqrt(n())),
  .vars = c('freezing','motion')
)

#novel stressor startle
novstressor.startle.stats = summarise_at(
  .tbl = group_by(
    .data = novstressor.startle,
    component, NS),
  .funs = list(mean=~mean(.), sem=~sd(.)/sqrt(n())),
  .vars = c('freezing','motion')
)


#sefl test
sefltest.stats = summarise_at(
  .tbl = group_by(
    .data = sefltest,
    NS),
  .funs = list(mean=~mean(.), sem=~sd(.)/sqrt(n())),
  .vars = c('freezing','motion')
)

#trauma recall with light 
traumatest_light.stats = summarise_at(
  .tbl = group_by(
    .data = traumatest_light.collapsed,
    light, subphase),
  .funs = list(mean=~mean(.), sem=~sd(.)/sqrt(n())),
  .vars = c('freezing','motion')
)


```



# Plots
## Global Plot Parameters
```{r}

#set figure file prefix
f_pfx = '10N6d_EGFP'

p = list()

#ggsave
p$save = FALSE
p$savetype = '.svg'

##Plot dimensions
p$dim = list(
  width = 4.5, #Plot width (in)
  height = 4.5, #Plot height (in)
  panel_w = unit(1, 'in'),
  panel_h = unit(2, 'in'),
  dpi = 500 #Plot dpi
)

#Plot text
p$textsize = list(
  title = 20,
  legend = 16,
  axis_title = 16,
  axis_text = 14
)

#Plot line dimensions
p$linesize = list(
  data_grp = 1,
  data_ind = .6,
  axis = .5,
  bar = .7,
  error =.75
)
p$linealpha =list(
  data_ind = .2
)

#Plot Widths
p$width = list(
  bar = .8,
  jitter = .3,
  error = .5
)

#Point/dot sizes
p$points = list(
  grp = list(
    shape = 21,
    size = 4,
    stroke = 1,
    alpha = 1
  ),
  ind = list(
    shape =  21,
    size = 2.5,
    alpha = .4
  )
)

#theme
p$theme = list(
  base = theme_bw(),
  specs = theme(
    panel.grid = element_blank(), 
    panel.border = element_rect(colour="white"),
    axis.line = element_line(
      linewidth=p$linesize$axis,
      color="black"
    ),
    plot.title = element_text(
      size=p$textsize$title, 
      hjust=.5, 
      margin=margin(b=15,t=10), 
      face="bold"
    ),
    axis.title = element_text(
      size=p$textsize$axis_title, 
      face="plain"
    ),
    axis.text=element_text(
      size=p$textsize$axis_text, 
      colour="black"
    ),
    legend.position = c(.3,.85),
    legend.title = element_blank(),
    legend.text = element_text(
      size=p$textsize$legend, 
      face="plain"
    ),
    legend.key = element_rect(colour=NA, fill=NA),
    legend.background = element_rect(fill=NA),
    plot.background = element_rect(fill = "transparent", colour = NA),
  ) 
)

#Plot colors
p$cols = list(
  color = c("white","blue4"),
  fill = c("white","blue4"),
  rec = 'slategray1'
)

#predefine partial function for saving plots
saveplot = partial(
  ggsave,
  width = p$dim$width,
  height = p$dim$height,
  units = "in", 
  dpi = p$dim$dpi,
  bg = 'transparent'
)

```


## Make Plots

### Trauma
#### Freezing
```{r, fig.width=4.5, fig.height=4.5}


#Create base layer from group level data
P.Trauma.Freezing = ggplot(
  data = trauma.freezing.stats,
  aes(x = component,
      y = freezing_mean,
      fill = NS,
      group = interaction(subphase, NS)
      )
  )

#Define plot specific parameters
p.ymin = 0
p.ymax = 100
p.yint = 25
p.breaks = seq(p.ymin,p.ymax,p.yint)

#Create plot
P.Trauma.Freezing  +
  
  geom_errorbar(
    stat = "identity",
    linewidth = p$linesize$error,
    colour = "black",
    aes(ymin = freezing_mean - freezing_sem,
        ymax = freezing_mean + freezing_sem,
        width = p$width$error
        )) +
  
  geom_line(
    stat = "identity",
    linewidth = p$linesize$data_grp
  ) +
  
  geom_point(
    stat = "identity",
    size = p$points$grp$size,
    stroke = p$points$grp$stroke,
    shape = p$points$grp$shape,
    show.legend = TRUE
  ) +
  
  scale_fill_manual(values=p$cols$fill) +
  coord_cartesian(ylim=c(p.ymin,p.ymax)) +
  scale_y_continuous(breaks=p.breaks, expand = c(0, 0)) +
  scale_x_discrete(
    labels=c('P','1','2','3','4','5','6','7','8','9','10'),
    expand = c(.1,.1)
  ) +
  
  labs(
    x = "Shock",
    y = "% Freezing"
  ) +
  guides(fill=guide_legend(title="Stressor 2\nLaser")) +
  
  p$theme$base +
  p$theme$specs +
  
  theme(
    axis.text.x = element_text(size=p$textsize$axis_text-4),
    #axis.title.x = element_text(margin=margin(t=-5)),
    legend.title = element_text(
      hjust=0, 
      size=p$textsize$axis_title, 
      margin = margin(b=-5)
    ),
    legend.position = 'right', #c(.375,.8),
    legend.box.margin = margin(t=-0)
  ) +
  
  force_panelsizes(
      rows = p$dim$panel_h,
      cols = unit(1.35, "in")
  ) 

if (p$save == TRUE) {
  saveplot(filename = paste(f_pfx,'_',"Trauma", p$savetype, sep = ''))
}

```



#### Shock Reactivity
```{r, fig.width=4.5, fig.height=4.5}

#Create base layer from group level data
P.Trauma.Shock = ggplot(
  data = trauma.shockreact.stats,
  aes(x = subphase,
      y = mean,
      fill = NS,
      group = NS)) 

#Define plot specific parameters
p.ymin = 0
p.ymax = 1600
p.yint = 400
p.breaks = seq(p.ymin,p.ymax,p.yint)

#Create plot
P.Trauma.Shock  +
  
  geom_errorbar(
    stat = "identity",
    linewidth = p$linesize$error,
    colour = "black",
    aes(ymin = mean - sem,
        ymax = mean + sem,
        width = p$width$error)) +
  
  geom_line(
    stat="identity",
    linewidth = p$linesize$data_grp
  ) +
  
  geom_point(
    stat = "identity",
    size = p$points$grp$size,
    shape = p$points$grp$shape,
    stroke = p$points$grp$stroke,
    show.legend = FALSE
  ) +
  
  scale_fill_manual(values=p$cols$fill) +
  coord_cartesian(ylim=c(p.ymin,p.ymax)) +
  scale_y_continuous(breaks=p.breaks, expand = c(0, 0)) +
  scale_x_discrete(labels=c('bl','shock')) +
  
  labs(
    x = "",
    y = "Motion (au)"
  ) +
  
  p$theme$base +
  p$theme$specs +

  force_panelsizes(
      rows = p$dim$panel_h,
      cols = p$dim$panel_w
  ) 

# if (p$save == TRUE) {
#   saveplot(filename = paste(f_pfx,'_',"Trauma_ShockReact", p$savetype, sep = ''))
# }

```


### Trauma Recall 1 (no laser)
```{r, fig.width=4.5, fig.height=4.5}

#Create base layer from group level data
P.TR = ggplot(
  data = traumatest.stats,
  aes(x = NS,
      y = freezing_mean,
      fill = NS))

#Define plot specific parameters
p.ymin = 0
p.ymax = 100
p.yint = 25
p.breaks = seq(p.ymin,p.ymax,p.yint)

#Create plot
P.TR  +

  geom_crossbar(
    stat = "identity",
    width = p$width$bar,
    linewidth = p$linesize$error,
    colour = "black",
    aes(ymin = freezing_mean - freezing_sem,
        ymax = freezing_mean + freezing_sem,
        ),
    show.legend = FALSE
    ) +

  geom_jitter(
      data = traumatest,
      aes(x = NS,
          y = freezing),
      width = p$width$bar * p$width$jitter,
      height = 0,
      shape = p$points$ind$shape,
      size = p$points$ind$size,
      colour = 'black',
      fill = NA,
      alpha = p$points$ind$alpha,
      show.legend=FALSE
  ) +
  
  scale_fill_manual(values=p$cols$fill) +
  coord_cartesian(ylim=c(p.ymin,p.ymax)) +
  scale_y_continuous(breaks=p.breaks, labels = p.breaks, expand = c(0, 0)) +
  
  labs(
    x = 'Stressor 2\nLaser',
    y = "% Freezing"
  ) +
  
  p$theme$base +
  p$theme$specs +
  theme(
    legend.position = c(1.5,.2),
    axis.text.x = element_text(angle=-0, hjust=0.5)
  ) +

  force_panelsizes(
      rows = p$dim$panel_h,
      cols = p$dim$panel_w
  ) 

if (p$save == TRUE) {
  saveplot(filename = paste(f_pfx,'_',"TraumaRecall1", p$savetype, sep = ''))
}

```



### Novel Stressor
#### Pre-Post
```{r, fig.width=4.5, fig.height=4.5}


#Create base layer from group level data
P.NovSts.prepost = ggplot(
  data = novstressor.prepost.stats,
  aes(x = component,
      y = freezing_mean,
      fill = NS,
      group = NS))

#Define plot specific parameters
p.ymin = 0
p.ymax = 50
p.yint = 10
p.breaks = seq(p.ymin,p.ymax,p.yint)

#Create plot
P.NovSts.prepost  +
  
  geom_tile(
    aes(x='post',y=0),
    width = 1,
    height = Inf,
    show.legend = FALSE,
    fill = p$cols$rec
  ) +
  
  geom_errorbar(
    stat = "identity",
    linewidth = p$linesize$error,
    colour = "black",
    aes(ymin = freezing_mean - freezing_sem,
        ymax = freezing_mean + freezing_sem,
        width = p$width$error
        )) +
  
  geom_line(
    stat="identity",
    linewidth = p$linesize$data_grp
  ) +
  
  geom_point(
    stat = "identity",
    size = p$points$grp$size,
    shape = p$points$grp$shape,
    stroke = p$points$grp$stroke,
    show.legend = TRUE
  ) +
  
  scale_fill_manual(values=p$cols$fill) +
  coord_cartesian(ylim=c(p.ymin,p.ymax)) +
  scale_y_continuous(breaks=p.breaks, labels = p.breaks, expand = c(0, 0)) +
  
  labs(
    x = '',
    y = "% Freezing"
  ) +
  guides(fill=guide_legend(title="Stressor 2\nLaser")) +
  
  p$theme$base +
  p$theme$specs +
  
  theme(
    legend.title = element_text(
      hjust=0, 
      size=p$textsize$axis_title, 
      margin = margin(b=-5)
    ),
    legend.position = c(1.55, .15),
  ) +
  
  force_panelsizes(
      rows = p$dim$panel_h,
      cols = p$dim$panel_w
  ) 

if (p$save == TRUE) {
  saveplot(filename = paste(f_pfx,'_',"NovSts_PrePost", p$savetype, sep = ''))
}

```



#### Startle
```{r, fig.width=4.5, fig.height=4.5}


#Create base layer from group level data
P.NovSts.startle = ggplot(
  data = novstressor.startle.stats,
  aes(x = component,
      y = motion_mean,
      fill = NS,
      group = NS))

#Define plot specific parameters
p.ymin = 0
p.ymax = 2500
p.yint = 500
p.breaks = seq(p.ymin,p.ymax,p.yint)

#Create plot
P.NovSts.startle  +
  
  geom_errorbar(
    stat = "identity",
    linewidth = p$linesize$error,
    colour = "black",
    aes(ymin = motion_mean - motion_sem,
        ymax = motion_mean + motion_sem,
        width = p$width$error
        )) +
  
  geom_line(
    stat="identity",
    linewidth = p$linesize$data_grp
  ) +
  
  geom_point(
    stat = "identity",
    size = p$points$grp$size,
    shape = p$points$grp$shape,
    stroke = p$points$grp$stroke,
    show.legend = FALSE
  ) +
  
  scale_fill_manual(values=p$cols$fill) +
  coord_cartesian(ylim=c(p.ymin,p.ymax)) +
  scale_x_discrete(labels=c('Pre','Noise','Post')) +
  scale_y_continuous(breaks=p.breaks, labels = p.breaks, expand = c(0, 0)) +
  
  labs(
    x = '',
    y = "Motion (au)"
  ) +
  
  p$theme$base +
  p$theme$specs +
  
  theme(
    axis.text.x= element_text(angle=-90, hjust=0),
  ) +
  
  force_panelsizes(
      rows = p$dim$panel_h,
      cols = p$dim$panel_w
  ) 

# if (p$save == TRUE) {
#   saveplot(filename = paste(f_pfx,'_',"NovSts_Startle", p$savetype, sep = ''))
# }

```



### SEFL Test
```{r, fig.width=4.5, fig.height=4.5}

#Create base layer from group level data
P.sefltest = ggplot(
  data = sefltest.stats,
  aes(x = NS,
      y = freezing_mean,
      fill = NS)) 

#Define plot specific parameters
p.ymin = 0
p.ymax = 50
p.yint = 10
p.breaks = seq(p.ymin,p.ymax,p.yint)

#Create plot
P.sefltest  +
  
  geom_crossbar(
    stat = "identity",
    position = position_dodge(),
    linewidth = p$linesize$error,
    colour = "black",
    aes(ymin = freezing_mean - freezing_sem,
        ymax = freezing_mean + freezing_sem),
    width = p$width$bar,
    show.legend = FALSE) +
  
  geom_jitter(
      data = sefltest,
      aes(x = NS,
          y = freezing),
      width = p$width$bar * p$width$jitter,
      height = 0,
      shape = p$points$ind$shape,
      size = p$points$ind$size,
      colour = 'black',
      fill = NA,
      alpha = p$points$ind$alpha,
      show.legend=FALSE
  ) +
  
  scale_fill_manual(values=p$cols$fill) +
  coord_cartesian(ylim=c(p.ymin,p.ymax)) +
  scale_y_continuous(breaks=p.breaks, labels = p.breaks, expand = c(0, 0)) +
  
  labs(
    x = "",
    y = "% Freezing"
  ) +
  
  p$theme$base +
  p$theme$specs +
  
  theme(
    legend.position = c(1,1)
  ) +
  
  force_panelsizes(
      rows = p$dim$panel_h,
      cols = p$dim$panel_w, #unit(1.75, 'in')
  ) 

# if (p$save == TRUE) {
#   saveplot(filename = paste(f_pfx,'_',"SEFLtest", p$savetype, sep = ''))
# }

```


# Stats

## Trauma
```{r}

trauma.anova = ezANOVA(
  data = trauma.freezing[trauma.freezing$component!='baseline',],
  wid = Animal,
  dv = freezing,
  within = component,
  between = NS,
  white.adjust = TRUE,
  type = 3
)
trauma.anova

trauma.posthoc = list()
for (lvl in unique(trauma.freezing$component)) {
  trauma.posthoc[[lvl]] = ezANOVA(
    data = trauma.freezing[trauma.freezing$component==lvl,],
    wid = Animal,
    dv = freezing,
    between = NS,
    white.adjust = TRUE,
    type = 3
  )
}
trauma.posthoc

```

## Trauma Test (no light)
```{r}

traumatest.anova = ezANOVA(
  data = traumatest,
  wid = Animal,
  dv = freezing,
  between = NS,
  white.adjust = TRUE,
  type = 3
)
traumatest.anova

```

## Novel Stressor
```{r}

novstressor.anova = ezANOVA(
  data = novstressor.prepost,
  wid = Animal,
  dv = freezing,
  within = component,
  between = NS,
  white.adjust = TRUE,
  type = 3
)
novstressor.anova


```


# Export figure data for publication
```{r}

export = FALSE

#export S1 data
S1_export = trauma.freezing[,c('Animal','Sex','NS','component','freezing')]
names(S1_export)[names(S1_export)=='NS'] = 'Group'

#export S2 data
S1recall_export = traumatest[,c('Animal','Sex','NS','freezing')]
names(S1recall_export)[names(S1recall_export)=='NS'] = 'Group'

#export S2 data
S2_export =  novstressor.prepost[,c('Animal','Sex','NS','component','freezing')]
S2_export$component[S2_export$component=='bl'] = 'pre'
names(S2_export)[names(S2_export)=='NS'] = 'Group'


if (export){
  write.csv(S1_export, 'ExtFig5h_Stressor1.csv', row.names=FALSE)
  write.csv(S1recall_export, 'ExtFig5i_Stressor1recall.csv', row.names=FALSE)
  write.csv(S2_export, 'ExtFig5j_Stressor2.csv', row.names=FALSE)
}

```













