---
title: "SEFL21g3_Analysis"
author: "Zach"
output:
  html_document: 
    fig_height: 4.5
    fig_width: 4.5
    highlight: kate
    number_sections: yes
    theme: journal
    toc: yes
    toc_depth: 6
editor_options: 
  chunk_output_type: console
---


# Set Global Options
```{r global_options}
knitr::opts_chunk$set(
  fig.width=4.5, 
  fig.height=4.5, 
  echo=TRUE, 
  warning=FALSE,
  message=FALSE
  )
```



# Load Packages 
```{r}
library(ggplot2)
library(ez)
library(knitr)
library(tidyr)
library(lmtest)
library(car)
library(dplyr)
library(purrr)
library(ggh4x)
```



# Import Data 
Load  files and factorize as necessary.
```{r}

#load data
groups = read.csv('SEFL21g3_groups.csv', header=TRUE)
data_vf = read.csv('SEFL21g3_vfdata.csv', header=TRUE)
data_ld = read.csv('SEFL21g3_lddata.csv', header=TRUE)

#factorize animal to allow joing
groups$Virus = factor(groups$Virus, levels=c('mCherry','HM3DQ'))
groups$Virus = recode_factor(groups$Virus, 'mCherry' = 'mCherry', 'HM3DQ' = 'HM3Dq')
groups$Animal = factor(groups$Animal)
data_vf$Animal = factor(data_vf$Animal)
data_ld$Animal = factor(data_ld$Animal)

#add group info
data_vf = right_join(x=groups,y=data_vf,by="Animal")
data_ld = right_join(x=groups,y=data_ld,by="Animal")

```



# Drop excluded subjects
See `groups` dataframe for exclusion rationale.
```{r}

data_vf = data_vf[data_vf$Exclude!=1,]
data_ld = data_ld[data_ld$Exclude!=1,]

```



# Subset data
Separate phases (trauma, trauma recall, etc) are placed in unique dataframes for ease of analysis.
```{r}

#Collapse ld data
data_ld = data_ld[data_ld$bin<5,]
data_ld$drug = data_ld$LD1
data_ld$drug[data_ld$testnum==2] = data_ld$LD2[data_ld$testnum==2]
data_ld$drug = factor(data_ld$drug, levels = c("veh","cno"))
ld.collapsed = summarize_at(
  .tbl = group_by(data_ld, Animal, Virus, drug, testnum, Sex),
  .funs = list(mean),
  .vars = c('dark','Distance_in'))


#mild stressor data
##prepost startle stimulus
mildstressor.prepost = data_vf[data_vf$phase=='mildstressor' & data_vf$subphase=='prepost',]
mildstressor.prepost$component = factor(mildstressor.prepost$component, levels=c('bl','post'))

##startle response
mildstressor.startle = data_vf[data_vf$phase=='mildstressor' & data_vf$subphase=='startle',]
mildstressor.startle$component = factor(
  mildstressor.startle$component,
  levels=c('startle_pre','startle_startle','startle_post'))


#mild stressor test data
mildstressortest = data_vf[data_vf$phase=='mildstressortest' & data_vf$subphase=='avg',]

#trauma data
trauma.freezing = data_vf[data_vf$phase=='trauma' & data_vf$subphase!="shock",]

```



# Get summary statistics for plotting
Calculate relevant means and std. errors.
```{r}


#ld stats
ld.stats = summarise_at(
  .tbl = group_by(
    .data = ld.collapsed,
    Virus, drug),
  .funs = list(mean=~mean(.), sem=~sd(.)/sqrt(n())),
  .vars = c('dark','Distance_in'))


#mild stressor
#mild stressor pre/post
mildstressor.prepost.stats = summarise_at(
  .tbl = group_by(
    .data = mildstressor.prepost,
    component,Virus),
  .funs = list(mean=~mean(.), sem=~sd(.)/sqrt(n())),
  .vars = c('freezing','motion')
)


#mildstressor test
mildstressortest.stats = summarise_at(
  .tbl = group_by(
    .data = mildstressortest,
    Virus),
  .funs = list(mean=~mean(.), sem=~sd(.)/sqrt(n())),
  .vars = c('freezing','motion')
)


#trauma 
#trauma session freezing
trauma.freezing.stats = summarise_at(
  .tbl = group_by(
    .data = trauma.freezing,
    component, Virus),
  .funs = list(mean=~mean(.), sem=~sd(.)/sqrt(n())),
  .vars = c('motion','freezing')
)
trauma.freezing.stats$subphase = 'postshock'
trauma.freezing.stats$subphase[trauma.freezing.stats$component=='baseline'] = 'baseline'


```



# Plots
## Global Plot Parameters
```{r}

#set figure file prefix
f_pfx = '21g3'

p = list()

#ggsave
p$save = FALSE
p$savetype = '.svg'

##Plot dimensions
p$dim = list(
  width = 4.5, #Plot width (in)
  height = 4.5, #Plot height (in)
  panel_w = unit(1, 'in'),
  panel_h = unit(2, 'in'),
  dpi = 500 #Plot dpi
)

#Plot text
p$textsize = list(
  title = 20,
  legend = 14,
  axis_title = 16,
  axis_text = 14
)

#Plot line dimensions
p$linesize = list(
  data_grp = 1,
  data_ind = .6,
  axis = .5,
  bar = .7,
  error =.75
)
p$linealpha =list(
  data_ind = .2
)

#Plot Widths
p$width = list(
  bar = .8,
  jitter = .3,
  error = .3
)

#Point/dot sizes
p$points = list(
  grp = list(
    shape = 21,
    size = 4,
    stroke = 1,
    alpha = 1
  ),
  ind = list(
    shape =  21,
    size = 2.5,
    alpha = .4
  )
)

#theme
p$theme = list(
  base = theme_bw(),
  specs = theme(
    panel.grid = element_blank(), 
    panel.border = element_rect(colour="white"),
    axis.line = element_line(
      linewidth=p$linesize$axis,
      color="black"
    ),
    plot.title = element_text(
      size=p$textsize$title, 
      hjust=.5, 
      margin=margin(b=15,t=10), 
      face="bold"
    ),
    axis.title = element_text(
      size=p$textsize$axis_title, 
      face="plain"
    ),
    axis.text=element_text(
      size=p$textsize$axis_text, 
      colour="black"
    ),
    legend.position = c(.3,.85),
    legend.title = element_blank(),
    legend.text = element_text(
      size=p$textsize$legend, 
      face="plain"
    ),
    legend.key = element_rect(colour=NA, fill=NA),
    legend.background = element_rect(fill=NA),
    plot.background = element_rect(fill = "transparent", colour = NA),
  ) 
)

#Plot colors
p$cols = list(
  color = c("white","red4"),
  fill = c("white","red4"),
  rec = 'slategray1'
)

#predefine partial function for saving plots
saveplot = partial(
  ggsave,
  width = p$dim$width,
  height = p$dim$height,
  units = "in", 
  dpi = p$dim$dpi,
  bg = 'transparent'
)

```



### Light Dark, Pre-trauma
#### Dark
```{r, fig.width=4.5, fig.height=4.5}


#Create base layer from group level data
P.LD.dark = ggplot(
  data = ld.stats,
  aes(x = drug,
      y = dark_mean,
      fill = Virus,
      group = Virus))

#Define plot specific parameters
p.ymin = .4
p.ymax = 1
p.yint = .2
p.breaks = seq(p.ymin,p.ymax,p.yint)

#Create plot
P.LD.dark  +

  geom_hline(yintercept=.5, linetype='dashed') +
  
  geom_errorbar(
    stat = "identity",
    linewidth = p$linesize$error,
    colour = "black",
    aes(ymin = dark_mean - dark_sem,
        ymax = dark_mean + dark_sem,
        width = p$width$error
        )) +
  
  geom_line(
    stat = "identity",
    linewidth = p$linesize$data_grp
  ) +
  
  geom_point(
    stat = "identity",
    size = p$points$grp$size,
    shape = p$points$grp$shape,
    stroke = p$points$grp$stroke,
    show.legend = TRUE
  ) +
  
  scale_fill_manual(values=p$cols$fill) +
  coord_cartesian(ylim=c(p.ymin,p.ymax)) +
  scale_y_continuous(breaks=p.breaks, labels = p.breaks*100, expand = c(0, 0)) +
  
  labs(
    x = '',
    y = "% Time in Dark"
  ) +
  
  p$theme$base +
  p$theme$specs +
  theme(
    axis.text.x = element_text(angle=-45, hjust=0, vjust=1),
    legend.position = c(.5, -.4),
  ) +
  
  force_panelsizes(
      rows = p$dim$panel_h,
      cols = p$dim$panel_w
  ) 

if (p$save == TRUE) {
  saveplot(filename = paste(f_pfx,'_',"LD_Dark", p$savetype, sep = ''))
}

```


#### Distance 
```{r, fig.width=4.5, fig.height=4.5}


#Create base layer from group level data
P.LD.dist = ggplot(
  data = ld.stats,
  aes(x = drug,
      y = Distance_in_mean,
      fill = Virus,
      group = Virus))

#Define plot specific parameters
p.ymin = 0
p.ymax = 200
p.yint = 50
p.breaks = seq(p.ymin,p.ymax,p.yint)

#Create plot
P.LD.dist  +
  
  geom_errorbar(
    stat = "identity",
    linewidth = p$linesize$error,
    colour = "black",
    aes(ymin = Distance_in_mean - Distance_in_sem,
        ymax = Distance_in_mean + Distance_in_sem,
        width = p$width$error
        )) +
  
  geom_line(
    stat = "identity",
    linewidth = p$linesize$data_grp
  ) +
  
  geom_point(
    stat = "identity",
    size = p$points$grp$size,
    shape = p$points$grp$shape,
    stroke = p$points$grp$stroke,
    show.legend = FALSE
  ) +
  
  scale_fill_manual(values=p$cols$fill) +
  coord_cartesian(ylim=c(p.ymin,p.ymax)) +
  scale_y_continuous(breaks=p.breaks, labels = p.breaks, expand = c(0, 0)) +
  
  labs(
    x = '',
    y = "Distance (in/min)"
  ) +
  
  p$theme$base +
  p$theme$specs +
  
  theme(
    axis.text.x = element_text(angle=-45, hjust=0, vjust=1),
    legend.position = c(.5, -.4),
  ) +
  
  force_panelsizes(
      rows = p$dim$panel_h,
      cols = p$dim$panel_w
  ) 

if (p$save == TRUE) {
  saveplot(filename = paste(f_pfx,'_',"LD_Dist", p$savetype, sep = ''))
}

```


### Mild Stressor
#### Pre-Post
```{r, fig.width=4.5, fig.height=4.5}



#Create base layer from group level data
P.MildSts.prepost = ggplot(
  data = mildstressor.prepost.stats,
  aes(x = component,
      y = freezing_mean,
      fill = Virus,
      group = Virus))

#Define plot specific parameters
p.ymin = 0
p.ymax = 75
p.yint = 25
p.breaks = seq(p.ymin,p.ymax,p.yint)

#Create plot
P.MildSts.prepost  +
  
  geom_errorbar(
    stat = "identity",
    linewidth = p$linesize$error,
    colour = "black",
    aes(ymin = freezing_mean - freezing_sem,
        ymax = freezing_mean + freezing_sem,
        width = p$width$error
        )) +
  
  geom_line(
    stat="identity",
    linewidth = p$linesize$data_grp
  ) +
  
  geom_point(
    stat = "identity",
    size = p$points$grp$size,
    shape = p$points$grp$shape,
    stroke = p$points$grp$stroke,
    show.legend = TRUE   
  ) +
  
  scale_fill_manual(values=p$cols$fill) +
  coord_cartesian(ylim=c(p.ymin,p.ymax)) +
  scale_x_discrete(labels=c('pre','post')) +
  scale_y_continuous(breaks=p.breaks, labels = p.breaks, expand = c(0, 0)) +
  
  labs(
    x = '',
    y = "% Freezing"
  ) +
  
  p$theme$base +
  p$theme$specs +
  
  theme(
    axis.text.x = element_text(angle=-45, hjust=0, vjust=1),
    legend.position = c(.5, -.4),
  ) +
  
  force_panelsizes(
      rows = p$dim$panel_h,
      cols = p$dim$panel_w
  ) 

if (p$save == TRUE) {
  saveplot(filename = paste(f_pfx,'_',"S2_PrePost", p$savetype, sep = ''))
}

```




### Mild Stressor Test
```{r, fig.width=4.5, fig.height=4.5}

#Create base layer from group level data
P.MSTest = ggplot(
  data = mildstressortest.stats,
  aes(x = Virus,
      y = freezing_mean,
      fill = Virus,
      group = Virus)) 

#Define plot specific parameters
p.ymin = 0
p.ymax = 40
p.yint = 10
p.breaks = seq(p.ymin,p.ymax,p.yint)

#Create plot
P.MSTest  +
  
  geom_crossbar(
    stat = "identity",
    linewidth = p$linesize$error,
    colour = "black",
    aes(ymin = freezing_mean - freezing_sem,
        ymax = freezing_mean + freezing_sem),
    width = p$width$bar,
    show.legend = FALSE) +
  
  
  geom_jitter(
    data = mildstressortest,
    aes(x = Virus,
        y = freezing),
    width = p$width$bar * p$width$jitter,
    height = 0,
    shape = p$points$ind$shape,
    size = p$points$ind$size,
    colour = 'black',
    fill = NA,
    alpha = p$points$ind$alpha,
    show.legend=FALSE
  ) +
  
  scale_fill_manual(values=p$cols$fill) +
  coord_cartesian(ylim=c(p.ymin,p.ymax)) +
  scale_y_continuous(breaks=p.breaks, labels = p.breaks, expand = c(0, 0)) +
  
  labs(
    x = "",
    y = "% Freezing"
  ) +
  
  p$theme$base +
  p$theme$specs +
  
  theme(
    axis.text.x = element_text(angle=-45, hjust=0, vjust=1),
    legend.position = c(.5, -.4),
  ) +
  
  force_panelsizes(
      rows = p$dim$panel_h,
      cols = p$dim$panel_w
  ) 

# if (p$save == TRUE) {
#   saveplot(filename = paste(f_pfx,'_',"S2_Stl", p$savetype, sep = ''))
# }


```


### Trauma
#### Freezing
```{r, fig.width=4.5, fig.height=4.5}


#Create base layer from group level data
P.Trauma.Freezing = ggplot(
  data = trauma.freezing.stats,
  aes(x = component,
      y = freezing_mean,
      fill = Virus,
      shape = Virus,
      group = interaction(subphase,Virus)))

#Define plot specific parameters
p.ymin = 0
p.ymax = 100
p.yint = 25
p.breaks = seq(p.ymin,p.ymax,p.yint)

#Create plot
P.Trauma.Freezing  +
  
  geom_errorbar(
    stat = "identity",
    linewidth = p$linesize$error,
    colour = "black",
    aes(ymin = freezing_mean - freezing_sem,
        ymax = freezing_mean + freezing_sem,
        width = p$width$error
        )) +
  
  geom_line(
    stat = "identity",
    linewidth = p$linesize$data_grp
  ) +
  
  geom_point(
    stat = "identity",
    size = p$points$grp$size,
    shape = p$points$grp$shape,
    stroke = p$points$grp$stroke,
    show.legend = TRUE
  ) +
  
  scale_fill_manual(values=p$cols$fill) +
  coord_cartesian(ylim=c(p.ymin,p.ymax)) +
  scale_y_continuous(breaks=p.breaks, expand = c(0, 0)) +
  scale_x_discrete(labels=c('bl','1','2','3','4','5','6','7','8','9','10')) +
  
  labs(
    x = "Shock",
    y = "% Freezing"
  ) +
  
  p$theme$base +
  p$theme$specs +
  
  theme(
    axis.text.x = element_text(size=p$textsize$axis_text-4),
    #axis.title.x = element_text(margin=margin(t=-5)),
    legend.title = element_blank(),
    legend.position = c(.45, 1.02),
    legend.box.margin = margin(t=-0)
  ) +
  
  force_panelsizes(
      rows = p$dim$panel_h,
      cols = unit(1.35, "in")
  ) 

# if (p$save == TRUE) {
#   saveplot(filename = paste(f_pfx,'_',"Trauma", p$savetype, sep = ''))
# }

```


# Stats

## Light Dark Stats
### Dark side time
```{r}


#comparison of time in dark side
ld.dark.anova = ezANOVA(
  data = ld.collapsed,
  dv = dark,
  wid = Animal,
  between = Virus,
  within = drug,
  white.adjust = TRUE,
  type = 3,
  detailed = TRUE
)
ld.dark.anova

```


### Distance travelled
```{r}

#comparison of time in dark side
ld.dist.anova = ezANOVA(
  data = ld.collapsed,
  dv = Distance_in,
  wid = Animal,
  between = Virus,
  within = drug,
  white.adjust = TRUE,
  type = 3,
  detailed = TRUE
)
ld.dist.anova

```


## Mild Stressor 
```{r}

ms.anova = ezANOVA(
  data = mildstressor.prepost,
  dv = freezing,
  wid = Animal,
  within = component,
  between = Virus,
  white.adjust = TRUE,
  type = 3,
  detailed = TRUE
)
ms.anova

ms.bl.anova = ezANOVA(
  data = mildstressor.prepost[mildstressor.prepost$component=='bl',],
  dv = freezing,
  wid = Animal,
  between = Virus,
  white.adjust = TRUE,
  type = 3,
  detailed = TRUE
)
ms.bl.anova

ms.post.anova = ezANOVA(
  data = mildstressor.prepost[mildstressor.prepost$component=='post',],
  dv = freezing,
  wid = Animal,
  between = Virus,
  white.adjust = TRUE,
  type = 3,
  detailed = TRUE
)
ms.post.anova



```


## Mild Stressor Test 
```{r}

mstest.anova = ezANOVA(
  data = mildstressortest,
  dv = freezing,
  wid = Animal,
  between = Virus,
  white.adjust = TRUE,
  type = 3,
  detailed = TRUE
)
mstest.anova

```


## Trauma
### full session
```{r}

trauma.fz.anova = ezANOVA(
  data = trauma.freezing,
  dv = freezing,
  wid = Animal,
  within = component,
  white.adjust = TRUE,
  between = Virus,
  type = 3,
  detailed = TRUE
)
trauma.fz.anova

```
### post shock 1
```{r}

trauma.s1.fz.comparison = t.test(freezing ~ Virus, data=trauma.freezing[trauma.freezing$component=='postshock01',], var.equal = TRUE)

```


# Export figure data for publication
```{r}

export = FALSE

#export S1 data
S_export = mildstressor.prepost[,c('Animal','Sex','Virus','component','freezing')]

#export light-dark data
ld_export = ld.collapsed[,c('Animal','Sex','Virus','drug','dark','Distance_in')]
ld_export$dark = ld_export$dark*100

if (export){
  write.csv(S_export, 'Fig3o.csv', row.names=FALSE)
  write.csv(ld_export, 'ExtFig4l.csv', row.names=FALSE)
}

```









