---
title: "SEFL25d_CaAnalysis"
author: "Zach"
output:
  html_document: 
    fig_height: 4.5
    fig_width: 4.5
    highlight: kate
    number_sections: yes
    theme: journal
    toc: yes
    toc_depth: 6
editor_options: 
  chunk_output_type: console
---


# Set Global Options
```{r global_options}

knitr::opts_chunk$set(fig.width=4.5, fig.height=4.5, echo=TRUE, warning=FALSE, message=FALSE)

```



# Load Packages 
```{r}

library(ggplot2)
library(svglite)
library(ez)
library(knitr)
library(tidyr)
library(dplyr)
library(purrr)
library(ggh4x)
library(stringr)
library(arrow)


```



# Import Data 
Load  files and factorize as necessary.
```{r}

#load unit classifications and unit responses
unit_class = list()
sig = list()
for (f in dir('UnitClassification/')){
  unit_class[[f]] = read_feather(paste('UnitClassification/',f,sep=''))
}
for (f in dir('BinnedUnitResponses/')){
  sig[[f]] = read_feather(paste('BinnedUnitResponses/',f,sep=''))
}
unit_class = do.call('rbind', unit_class)
sig = do.call('rbind', sig)

sig$amp = factor(sig$amp, levels=c('high','low'))
sig$amp = recode_factor(sig$amp, 'high'='High','low'='Low')


```



# Pre-process data
## Define classifications and add to signal data
```{r}

thresh = 95

#pivot wider so each cell has one column
unit_class_wd = pivot_wider(
  data = unit_class,
  id_cols = c('Animal','Experiment','unit_id'),
  names_from = c('type'),
  names_prefix = 'pctshuff_',
  values_from = c('pct_shuff')
)

#here cells are classified only if they display an increase at p=.05, one-tailed
unit_class_wd = mutate(unit_class_wd, ShockResp=ifelse(pctshuff_shock>=thresh,TRUE,FALSE))
unit_class_wd = mutate(unit_class_wd, SoundResp=ifelse(pctshuff_sound>=thresh,TRUE,FALSE))
unit_class_wd = mutate(
  unit_class_wd, 
  Classification=case_when(
    (ShockResp & !SoundResp) ~ 'shock',
    (!ShockResp & SoundResp) ~ 'sound',
    (ShockResp & SoundResp) ~ 'both',
    (!ShockResp & !SoundResp) ~ 'neither'
  )
)
unit_class_wd$Classification = factor(
  unit_class_wd$Classification, 
  levels=c('neither','sound','shock','both')
)

```


## Give each unit unique id and add classification to signal data
```{r}

sig = mutate(sig, uniqID = paste(Animal,Experiment,unit_id))
unit_class_wd = mutate(unit_class_wd, uniqID = paste(Animal,Experiment,unit_id))
sig = left_join(sig, unit_class_wd, by=c('Animal','Experiment','unit_id','uniqID'))

```


## Get Average Unit Response for each event (sounds and shocks of both amps)
```{r}

sig = sig[!is.na(sig$sig),]
sig_avg = summarise_at(
  .tbl = group_by(sig,uniqID,ShockResp,SoundResp,Classification,type,amp,bin),
  .funs = list(sig=mean),
  .vars = c('sig')
)

#Define average post-event response for each unit
period = c(0, 5000) 
sig_avg_post = summarise_at(
  .tbl = group_by(
      sig[sig$bin>period[1] & sig$bin<=period[2],], 
      uniqID,ShockResp,SoundResp,Classification,type,amp
    ),
  .funs = list(sig=mean),
  .vars = c('sig')
)


```


# Get summary statistics for plotting
Calculate relevant means and std. errors.
```{r}

sem = function(x) {sd(x)/sqrt(length(x))}

sig_avg.shock.stats = summarise_at(
  .tbl = group_by(
    .data = sig_avg[sig_avg$type=='shock',],
    ShockResp, amp, bin),
  .funs = list(mean=mean, sem=sem),
  .vars = c('sig'))

sig_avg.sound.stats = summarise_at(
  .tbl = group_by(
    .data = sig_avg[sig_avg$type=='sound',],
    SoundResp, amp, bin),
  .funs = list(mean=mean, sem=sem),
  .vars = c('sig'))

```



# Plots
## Global Plot Parameters
```{r}

#set figure file prefix
f_pfx = '25d_Ca2'

#list of plotting parameters
p = list()

#ggsave
p$save = FALSE
p$savetype = '.svg'

p$dim = list(
  width = 4.5, #Plot width (in)
  height = 4.5, #Plot height (in)
  panel_w = unit(1.5, 'in'),
  panel_h = unit(2, 'in'),
  dpi = 500 #Plot dpi
)

#Plot text
p$textsize = list(
  title = 20,
  legend = 16,
  axis_title = 16,
  axis_text = 14
)

#Plot line dimensions
p$linesize = list(
  data_grp = 1,
  data_ind = .6,
  axis = .5,
  bar = .7,
  error =.75
)
p$linealpha =list(
  data_ind = .3
)

#Plot Widths
p$width = list(
  bar = .8,
  jitter = .3,
  error = .5
)

#Point/dot sizes
p$points = list(
  grp = list(
    shape = 21,
    size = 4,
    stroke = 1,
    alpha = 1
  ),
  ind = list(
    shape =  21,
    size = 2.5,
    alpha = .4
  )
)

#Plot colors
p$cols = list(
  fill_shk = c('orangered3','orange2'),
  fill_snd = c('blue3','cornflowerblue'),
  rec = 'slategray1'
)

#theme
p$theme = list(
  base = theme_bw(),
  specs = theme(
    panel.grid = element_blank(), 
    panel.border = element_rect(colour="white"),
    axis.line = element_line(
      linewidth=p$linesize$axis,
      color="black"),
    plot.title = element_text(
      size=p$textsize$title, 
      hjust=.5, 
      margin=margin(b=15,t=10), 
      face="bold"),
    axis.title = element_text(
      size=p$textsize$axis_title, 
      face="plain"),
    axis.text=element_text(
      size=p$textsize$axis_text, 
      colour="black",
      face = 'plain'),
    strip.background = element_blank(),
    strip.text = element_text(
      size=p$textsize$axis_title, 
      colour="black",
      face = 'plain'),
    legend.position = c(.1,.9),
    legend.title = element_blank(),
    legend.text = element_text(size=p$textsize$legend, face="plain"),
    legend.key = element_rect(colour=NA, fill=NA),
    legend.background = element_rect(fill=NA),
    plot.background = element_rect(fill = "transparent", colour = NA),
  ) 
)

saveplot = partial(
  ggsave,
  width = p$dim$width,
  height = p$dim$height,
  units = "in",
  dpi = p$dim$dpi,
  bg = 'transparent'
)

#predefine partial function for saving plots
saveplot = partial(
  ggsave,
  width = p$dim$width,
  height = p$dim$height,
  units = "in",
  dpi = p$dim$dpi,
  bg = 'transparent'
)

```


## Make Plots

##### Shock Responses 
```{r, fig.width=4.5, fig.height=4.5}

#Create base layer from group level data
P.ShkResp.ShkNeurons = ggplot(
  data = sig_avg.shock.stats[sig_avg.shock.stats$ShockResp,],
  aes(x = bin,
      y = mean,
      fill = amp,
      group = amp))

#Define plot specific parameters
p.xmin = -2000
p.xmax = 10000
p.xint = 5000
p.xbreaks = seq(0, p.xmax, p.xint)
p.ymin = -.1
p.ymax = 6
p.yint = 1
p.ybreaks = seq(0,p.ymax,p.yint)

#Create plot
P.ShkResp.ShkNeurons  +
  
  geom_vline(xintercept = c(0,2000), linetype='dashed', color='gray50') +
  
  geom_ribbon(
    stat = "identity",
    linewidth = p$linesize$error,
    color = 'black',
    aes(ymin = mean - sem,
        ymax = mean + sem
        )) +
  
  scale_fill_manual(values = p$cols$fill_shk) +
  scale_y_continuous(breaks=p.ybreaks, labels = p.ybreaks, expand = c(0, 0)) +
  scale_x_continuous(breaks=p.xbreaks, labels=p.xbreaks/1000) +
  coord_cartesian(ylim=c(p.ymin,p.ymax), xlim=c(-1000,p.xmax)) +
  
  labs(
    x = 'T (sec)',
    y = "Calcium (z)"
  ) +
  
  p$theme$base +
  p$theme$specs +
  theme(
    strip.text = element_blank(),
    axis.text.x= element_text(face = 'plain',size = p$textsize$axis_text),
    legend.title = element_blank(),
    legend.position = 'right'
  ) +
  
  force_panelsizes(
      rows = p$dim$panel_h,
      cols = unit(1, "in")
    ) 

if (p$save == TRUE) {saveplot(filename = paste(f_pfx,'_',"ShkResp_ShkNeurons", p$savetype, sep = ''))}

```


##### Sound Responses 
```{r, fig.width=4.5, fig.height=4.5}

#Create base layer from group level data
P.SndResp.SndNeurons = ggplot(
  data = sig_avg.sound.stats[sig_avg.sound.stats$SoundResp,],
  aes(x = bin,
      y = mean,
      fill = amp,
      group = amp))

#Define plot specific parameters
p.xmin = -2000
p.xmax = 10000
p.xint = 5000
p.xbreaks = seq(0, p.xmax, p.xint)
p.ymin = -.1
p.ymax = 6
p.yint = 1
p.ybreaks = seq(0,p.ymax,p.yint)

#Create plot
P.SndResp.SndNeurons  +
  
  geom_vline(xintercept = c(0,2000), linetype='dashed', color='gray50') +
  
  geom_ribbon(
    stat = "identity",
    linewidth = p$linesize$error,
    color = 'black',
    aes(ymin = mean - sem,
        ymax = mean + sem
        )) +
  
  scale_fill_manual(values = p$cols$fill_snd) +
  scale_y_continuous(breaks=p.ybreaks, labels = p.ybreaks, expand = c(0, 0)) +
  scale_x_continuous(breaks=p.xbreaks, labels=p.xbreaks/1000) +
  coord_cartesian(ylim=c(p.ymin,p.ymax), xlim=c(-1000,p.xmax)) +
  
  labs(
    x = 'T (sec)',
    y = "Calcium (z)"
  ) +
  
  p$theme$base +
  p$theme$specs +
  theme(
    strip.text = element_blank(),
    axis.text.x= element_text(face = 'plain',size = p$textsize$axis_text),
    legend.title = element_blank(),
    legend.position = 'right'
  ) +
  
  force_panelsizes(
      rows = p$dim$panel_h,
      cols = unit(1., "in")
  ) 

if (p$save == TRUE) {saveplot(filename = paste(f_pfx,'_',"SndResp_SndNeurons", p$savetype, sep = ''))}

```


### Response proportions

#### Calculate proporptions
```{r}

#tabulate
UnitClassTbl = as.data.frame(prop.table(table(unit_class_wd$Classification)))
names(UnitClassTbl)[1] = 'Class'

```


## Bar Chart
### All cells
```{r}


P.bar = ggplot(
  UnitClassTbl, 
  aes(x = 1, y = Freq, fill = Class, label=Class)
) 

p.ymin = 0
p.ymax = 1
p.yint = .25
p.ybreaks = seq(p.ymin,p.ymax,p.yint)

P.bar +
  
  geom_col(
    width = 2,
    colour = 'gray10',
    #linewidth = ,
    show.legend = FALSE
  ) + 
  
  force_panelsizes(
      rows = p$dim$panel_h,
      cols = unit(1, "in")
  ) +
  
  geom_text(
    aes(
      x=2.3
    ),
    size = 5,
    fontface = 'plain',
    angle = 35,
    position = position_stack(vjust = 0.5),
    vjust = .5,
    hjust = 0
  ) +
  
  labs(
    x = '',
    y = "% Responsive"
  ) +
  
  scale_y_continuous(
    breaks = p.ybreaks, labels = p.ybreaks*100, expand = c(.005,.005)
  ) +
  scale_x_continuous(breaks=c(), limits=c(0,5)) +
  scale_fill_manual(values=c('gray90','blue2','orangered3','purple2')) +
  
  p$theme$base +
  p$theme$specs +
  theme(
    axis.line.x = element_blank()
  ) 

if (p$save == TRUE) {saveplot(filename = paste(f_pfx,'_',"StimSensitivity_BarClass_ALL", p$savetype, sep = ''))}

```


## Raster
### Shock resp
```{r}

pdata = droplevels(sig_avg[sig_avg$type=='shock' & sig_avg$ShockResp,])
pdata$amp = factor(pdata$amp, levels = c('Low','High'))

#Define plot specific parameters
p.xmin = -5000
p.xmax = 10000
p.xint = 5000
p.xbreaks = seq(p.xmin,p.xmax,p.xint)
p.ymin = -.1
p.ymax = 3
p.yint = 1
p.ybreaks = seq(0,p.ymax,p.yint)

P.raster.shock = ggplot(
  data = pdata, 
  aes(x = bin, y = uniqID, fill = sig)
) 

P.raster.shock +
  
  geom_raster() +
  
  geom_vline(xintercept = 0, linetype='dashed', color='red') +
  geom_vline(xintercept = 2000, linetype='dashed', color='red') +
  
  facet_grid(cols=vars(amp)) +
  
  scale_fill_gradient(
    low = 'gray93',
    high = 'black',
    limits=c(0,3),
    oob=scales::squish,
    name = 'z'
  ) +
  
  scale_x_continuous(breaks=p.xbreaks, labels=p.xbreaks/1000) +
  coord_cartesian(xlim=c(-2000,p.xmax)) +
  
  labs(
    x = 'T (sec)',
    y = "Neuron",
  ) +
  
  p$theme$base +
  p$theme$specs +
  theme(
    legend.title = element_text(
      size=p$textsize$axis_text, 
      vjust = 1
    ),
    legend.position = 'bottom',
    legend.key.size = unit(.4, 'cm'),
    legend.text = element_text(face='plain', size=p$textsize$axis_text-2),
    panel.spacing.x = unit(.5, 'cm'),
    strip.background = element_blank(),
    strip.text = element_text(
      size=p$textsize$axis_title, 
      hjust=.5
    ),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.text.x = element_text(size=p$textsize$axis_text-2),
    axis.title.x = element_text(margin = margin(t=0, b=0))
  ) +

  force_panelsizes(
      rows = p$dim$panel_h,
      cols = unit(.45, "in")
    ) 
  
if (p$save == TRUE) {saveplot(filename = paste(f_pfx,'_',"Raster_ShkResponsive", p$savetype, sep = ''))}

```


### Sound resp
```{r}

pdata = droplevels(sig_avg[sig_avg$type=='sound' & sig_avg$SoundResp,])
pdata$amp = factor(pdata$amp, levels = c('Low','High'))


#Define plot specific parameters
p.xmin = -5000
p.xmax = 10000
p.xint = 5000
p.xbreaks = seq(p.xmin,p.xmax,p.xint)
p.ymin = -.1
p.ymax = 3
p.yint = 1
p.ybreaks = seq(0,p.ymax,p.yint)

P.raster.sound = ggplot(
  data = pdata, 
  aes(x = bin, y = uniqID, fill = sig)
) 

P.raster.sound +
  
  geom_raster() +
  
  geom_vline(xintercept = 0, linetype='dashed', color='red') +
  geom_vline(xintercept = 2000, linetype='dashed', color='red') +
  
  facet_grid(cols=vars(amp)) +
  
  scale_fill_gradient(
    low = 'gray93',
    high = 'black',
    limits=c(0,3),
    breaks = 0:3,
    oob=scales::squish,
    name = 'z'
  ) +
  
  scale_x_continuous(breaks=p.xbreaks, labels=p.xbreaks/1000) +
  coord_cartesian(xlim=c(-2000,p.xmax)) +
  
  labs(
    x = 'T (sec)',
    y = "Neuron",
  ) +
  
  p$theme$base +
  p$theme$specs +
  theme(
    legend.title = element_text(
      size=p$textsize$axis_text, 
      vjust = 1
    ),
    legend.position = 'bottom',
    legend.key.size = unit(.4, 'cm'),
    legend.text = element_text(face='plain', size=p$textsize$axis_text-2),
    panel.spacing.x = unit(.5, 'cm'),
    strip.background = element_blank(),
    strip.text = element_text(
      size=p$textsize$axis_title, 
      hjust=.5
    ),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.text.x = element_text(size=p$textsize$axis_text-2),
    axis.title.x = element_text(margin = margin(t=0, b=0))
  ) +

  force_panelsizes(
      rows = p$dim$panel_h,
      cols = unit(.45, "in")
  ) 
  
if (p$save == TRUE) {saveplot(filename = paste(f_pfx,'_',"Raster_SndResponsive", p$savetype, sep = ''))}

```


## PCA
### Compute
```{r}


pcdata = pivot_wider(
  data = sig[,c('uniqID','trial','type','amp','bin','sig')],
  values_from = 'sig',
  names_from = "uniqID"
)
pcdata = pcdata[complete.cases(pcdata),]

pca = prcomp(pcdata[,-(1:4)])

pcdataX = cbind(pcdata, as.data.frame(pca$x))
pcdataX = mutate(
  pcdataX,
  case = case_when(
    bin <= 0 ~ 'pre-stimulus',
    bin > 0 & type=='sound' & amp=='Low' ~ 'low sound',
    bin > 0 & type=='sound' & amp=='High' ~ 'high sound',
    bin > 0 & type=='shock' & amp=='Low' ~ 'low shock',
    bin > 0 & type=='shock' & amp=='High' ~ 'high shock'
  )
)
pcdataX$case = factor(
  pcdataX$case, 
  levels = c(
    'pre-stimulus','low sound','high sound','low shock','high shock'
  )
)

pcdataXmean = summarise_at(
  .tbl = group_by(
    .data = pcdataX,
    case),
  .funs = list(mean),
  .vars = c('PC1','PC2','PC3','PC4','PC5')
)


```


### scree plot
```{r}

pca_summary = as.data.frame(t(summary(pca)$importance))
names(pca_summary) = c('stdev','proportion','cumprop')

p.eigen = ggplot(
  data = pca_summary,
  aes(y = proportion, x=1:nrow(pca_summary)))

p.eigen +
  geom_line() +
  scale_x_continuous(limits = c(1,10), breaks=1:10)

```

### plot
```{r}


p.pca_data = ggplot(
  data = pcdataXmean,
  aes(x=PC1, y=PC2, color=case)
)

#Define plot specific parameters
p.xmin = -5
p.xmax = 18
p.xint = 5
p.xbreaks = seq(p.xmin,p.xmax,p.xint)
p.ymin = -5
p.ymax = 18
p.yint = 5
p.ybreaks = seq(p.ymin,p.ymax,p.yint)

p.pca_data +
  
  
  geom_hline(yintercept=0, linetype='dashed', color='gray80') +
  geom_vline(xintercept=0, linetype='dashed', color='gray80') +
  
  geom_point(size=4, alpha=1) +
  
  scale_x_continuous(
    breaks=p.xbreaks, labels=p.xbreaks, 
    limits = c(p.xmin,p.xmax), expand = c(0, 0)) +
  scale_y_continuous(
    breaks=p.ybreaks, labels=p.ybreaks, 
    limits = c(p.ymin,p.ymax), expand = c(0, 0)) +
  scale_color_manual(
    values=c('gray50','cornflowerblue','blue3','orange2','orangered3')
  ) +
  
  p$theme$base +
  p$theme$specs +
  theme(
    legend.position = 'right',
    legend.title = element_blank(),
    legend.text = element_text(
      size=p$textsize$axis_text-4,
      face='plain'
    )
  ) +
  
  force_panelsizes(
      rows = unit(1.5, "in"),
      cols = unit(1.5, "in")
  ) 

if (p$save == TRUE) {saveplot(filename = paste(f_pfx,'_',"PCAplot", p$savetype, sep = ''))}

```




# Stats
## Post Rresponse
### Shock
```{r}

post.shock.anova = ezANOVA(
  data = sig_avg_post[sig_avg_post$ShockResp & sig_avg_post$type=='shock',],
  dv = sig,
  wid = uniqID,
  within = amp,
  type = 3,
  white.adjust = TRUE
)
post.shock.anova

```
### Sound
```{r}

post.sound.anova = ezANOVA(
  data = sig_avg_post[sig_avg_post$SoundResp & sig_avg_post$type=='sound',],
  dv = sig,
  wid = uniqID,
  within = amp,
  type = 3,
  white.adjust = TRUE
)
post.sound.anova

```


# Export figure data for publication
```{r}

export = FALSE

# export unit classification info
unit_export = unit_class_wd[,
  c('uniqID','Animal','pctshuff_shock','pctshuff_sound','Classification')
]
unit_export$Sex = 'M'
unit_export = unit_export[,c('uniqID','Animal','Sex','pctshuff_shock','pctshuff_sound','Classification')]

names(unit_export)[names(unit_export)=='uniqID'] = 'unit'
names(unit_export)[names(unit_export)=='pctshuff_shock'] = 'shock_pct'
names(unit_export)[names(unit_export)=='pctshuff_sound'] = 'sound_pct'
names(unit_export)[names(unit_export)=='Classification'] = 'class'


# export mean unit responses
unitresponse_export = sig_avg
names(unitresponse_export)[names(unitresponse_export)=='uniqID'] = 'unit'
names(unitresponse_export)[names(unitresponse_export)=='sig'] = 'signal'
names(unitresponse_export)[names(unitresponse_export)=='type'] = 'stimulus'
names(unitresponse_export)[names(unitresponse_export)=='amp'] = 'amplitude'
unitresponse_export$second = unitresponse_export$bin/1000
unitresponse_export = left_join(
  unitresponse_export,
  unit_export[,c('unit','Animal','Sex','class')],
  by = 'unit'
)
unitresponse_export = unitresponse_export[
  ,
  c('unit','Animal','Sex','class','stimulus','amplitude','second','signal')
]

pcdata_export = pcdata
names(pcdata_export)[names(pcdata_export)=='bin'] = 'second'
pcdata_export$second = pcdata_export$second/1000

if (export){
  write.csv(unit_export, 'Fig4e_units.csv', row.names=FALSE)
  write.csv(unitresponse_export, 'Fig4fg_unitresponse.csv', row.names=FALSE)
  write.csv(pcdata_export, 'Fig4h_unitresponse.csv', row.names=FALSE)
}

```










